---
title: "Prelim_data_exploration"
author: "Nicole Hill"
date: "15/09/2021"
output: html_document
---

```{r setup, message=FALSE, warning=FALSE,echo=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(raster)
library(rasterVis)
library(tidyverse)
library(PerformanceAnalytics) #correlation chart
library(RColorBrewer)


sci.dir <-      "C:/Users/hillna/OneDrive - University of Tasmania/UTAS_work/Projects/Benthic Diversity ARC/"
env.derived <-  paste0(sci.dir,"data_environmental/derived/")
tools.dir <-    paste0(sci.dir,"Analysis/Useful_Functions_Tools/")
ARC_Data.dir <- paste0(sci.dir,"Analysis/ARC_Data/")

load(paste0(env.derived,"Circumpolar_Coastline.Rdata"))

#source(paste0(tools.dir,"SOmap_functions_JJ.R")) 

```

## Environmental Data


## Plots of environmental layers for the shelf (defined as <3000m here) with location of samples. {.tabset}

Haven't used fancy SOmap as it takes way too long with 500m rasters. Revisit if needed for publication figures.

```{r plot env layers, echo=FALSE}

## get file names of all environmental rasters and bricks and load into one big stack

#all files with "gri" extension
env_list<-list.files(path = env.derived, pattern="gri$",  full.names=TRUE) 
#subset to  "shelf" files
env_list<-env_list[grep(".500m_shelf", env_list)]
#for the single raters layer names are missing. Extract from file name.
env_names<-gsub(".*_|\\..*","",env_list)


#stack all environmental layers and make sure they have appropriate names (currently manual and a bit messy!)
env_stack<-stack(env_list)
#names(env_stack)
names(env_stack)[1:5]<-env_names[1:5]
names(env_stack)[6:14]<-paste(rep(c("CARS_NO3", "CARS_O2", "CARS_PO4"),each=3),c("mean", "seas_range", "std_dev"), sep="_")
names(env_stack)[25]<-"NPP_su_mean"
names(env_stack)[38:39]<-env_names[13:14]


#add environmental data with non-conformant names
env_stack<-stack( env_stack,
                  raster(paste0(env.derived, "Circumpolar_EnvData_geomorphology")))
names(env_stack)[40]<-"geomorph"

geomorph_cat<-levels(env_stack[[40]])[[1]]



# load annotation data- need locations of cells with scored images
load(paste0(ARC_Data.dir, "annotation/Circumpolar_Annotation_Env_Data.RData"))


#plot subsets of environmental data. Using plain old raster as SOmap took way too long to plot with 500 m resolution rasters. Investigate later for publication plots.
# create function for adding locations of image cells
im_loc<-SpatialPoints(coords = cell_metadata_env[,c("proj_coord_x","proj_coord_y")], proj4string = crs(env_stack))
fun <- function() {
    points(im_loc, col="black", pch=21, cex=0.7)
}
```


### Bathymetry derivatives
```{r bathy plot, fig.height=9, fig.width=11}
raster::plot(env_stack, 1:5, nc=3, addfun=fun)
```


### Ice
```{r ice, fig.height=12, fig.width=10}
raster::plot(env_stack, grep("ice", names(env_stack)), nc=3, addfun=fun)
```

### SST
```{r sst,  fig.height=10, fig.width=8}
raster::plot(env_stack, grep("sst", names(env_stack)), nc=2, addfun=fun) 
```

### NPP
```{r NPP}
raster::plot(env_stack, "NPP_su_mean", addfun=fun)
```

### Seafloor parameters
``` {r seafloor,fig.height=6, fig.width=10}
raster::plot(env_stack, c("seafloorcurrents", "seafloortemperature"),addfun=fun)
```

### CARS BGC
```{r CARS,  fig.height=12, fig.width=10}
raster::plot(env_stack, grep("CARS", names(env_stack)), nc=3, addfun=fun)
```


### Geomorphology
```{r geomorph}

#define paleete to help categories stand out abit more - especially canyons
st<-brewer.pal("Greys", n=3)
can<-"lawngreen"
fin1<-brewer.pal("Paired", n=12)
fin2<-brewer.pal("Set3", n=11)


levelplot(subset(env_stack, "geomorph"),par.settings=rasterTheme(region=c(st,can,fin1,fin2)), main="Circum Antarctic")

# Zoom in on geomorphology to see how many sites/cells are near canyons

#Ross Sea
levelplot(subset(env_stack, "geomorph"),par.settings=rasterTheme(region=c(st,can,fin1,fin2)),
xlim=c(-600000,600000), ylim=c(-2650000,-1250000), main="Ross Sea", axes=FALSE) +
latticeExtra::layer (sp.lines(coast.proj, col="black")) +
latticeExtra::layer(sp.points(im_loc, col="blue"))

#Sabrina Coast
levelplot(subset(env_stack, "geomorph"),par.settings=rasterTheme(region=c(st,can,fin1,fin2)),
 xlim=c(2150000,2300000), ylim=c(-1350000,-1250000), main="Sabrina Coast", axes=FALSE) +
latticeExtra::layer (sp.lines(coast.proj, col="black")) +
latticeExtra::layer(sp.points(im_loc, col="blue"))


#Antarctic Peninsula
levelplot(subset(env_stack, "geomorph"),par.settings=rasterTheme(region=c(st,can,fin1,fin2)),
xlim=c(-2750000,-2100000), ylim=c(750000,2410000), main="Antarctic Peninsula", axes=FALSE) +
latticeExtra::layer (sp.lines(coast.proj, col="black")) +
latticeExtra::layer(sp.points(im_loc, col="blue"))

# weddell Sea
levelplot(subset(env_stack, "geomorph"),par.settings=rasterTheme(region=c(st,can,fin1,fin2)),
xlim=c(-1400000,600000), ylim=c(800000,2300000),main="Weddell Sea", axes=FALSE) +
latticeExtra::layer (sp.lines(coast.proj, col="black")) +
latticeExtra::layer(sp.points(im_loc, col="blue"))



```


## Density/distribution plots of environmental variables 

```{r env_density, fig.width=12, fig.height=18, warnings=FALSE}
cell_metadata_env %>% 
  filter(!is.na(cover_N)) %>%
  pivot_longer(cols= depth:seafloortemperature, names_to= "Variable", values_to= "Value") %>%
  ggplot (aes(x= Value))+
 geom_density()+
    theme_bw()+
  facet_wrap(~Variable, scales="free", nrow=7)
```

```{r geomorph barplot, fig.height=5, fig.width=6}
cell_metadata_env %>% 
  ggplot(aes(x=geomorph_cat))+
  geom_bar() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))


```


## Biplots with correlation between environmental variables at image cells {.tabset}

### Bathy Derivatives

```{r env_corr,  fig.width= 10, fig.height=10}
cell_metadata_env<-as_tibble(cell_metadata_env)

#bathy derivatives
cell_metadata_env %>% 
  dplyr::select(depth, slope, tpi, tpi5, tpi11) %>%
chart.Correlation(., na.action=na.omit)
title(main="Bathy vars", line=3)
```


### SST
```{r sst_corr,  fig.width= 10, fig.height=10}
cell_metadata_env %>% 
  dplyr::select(starts_with("sst")) %>%
chart.Correlation(., na.action=na.omit)
title(main="SST vars", line=3)
```


### CARS

```{r CARS_cor,  fig.width= 10, fig.height=10}

#CARS BGC
cell_metadata_env %>% 
  dplyr::select(starts_with("CARS")) %>%
chart.Correlation(., na.action=na.omit)
title(main="Ice vars", line=3)
```


### Ice 
```{r ice_cor, fig.width= 10, fig.height=10}
cell_metadata_env %>% 
  dplyr::select(starts_with("ice")) %>%
chart.Correlation(., na.action=na.omit)
title(main="Ice vars", line=3)
```

### Others
```{r,  fig.width= 10, fig.height=10}
cell_metadata_env %>% 
  dplyr::select(NPP_su_mean, seafloorcurrents, seafloortemperature) %>%
chart.Correlation(., na.action=na.omit)
title(main="Other vars", line=3)
```

### Across var types
``` {r across types, fig.width= 10, fig.height=10}

cell_metadata_env %>% 
  dplyr::select(depth, slope, tpi, ice_sp_mean, ice_sp_sd, ice_su_mean, ice_su_max, ice_su_sd, NPP_su_mean, sst_su_mean, sst_su_sd, seafloorcurrents, seafloortemperature) %>%
chart.Correlation(., na.action=na.omit)
title(main="Across var types", line=3)

```




