---
title: "Data_explore_Annotations"
author: "Nicole Hill"
date: "04/11/2021"
output: html_document
---

```{r setup, message=FALSE, warning=FALSE,echo=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(raster)
library(rasterVis)
library(tidyverse)
library(classInt)
library(scales)
library(kableExtra)
library(htmlTable)
library(RColorBrewer)

user = "Jan"
#user = "charley"
#user="nicole"

if (user == "Jan") {
  
  sci.dir <-      "C:/Users/jjansen/Desktop/science/"
  env.derived <-  paste0(sci.dir,"data_environmental/derived/")
  tools.dir <-    paste0(sci.dir,"SouthernOceanBiodiversityMapping/Useful_Functions_Tools/")
  ARC_Data.dir <- paste0(sci.dir,"SouthernOceanBiodiversityMapping/ARC_Data/")
  
} 
if (user == "charley") {
  
  sci.dir <- "C:/Users/cgros/code/IMAS/"
  ARC_Data.dir <- paste0(sci.dir,"ARC_Data/")
  env.derived <-  "C:/Users/cgros/data/SO_env_layers/derived/"
  tools.dir <-    paste0(sci.dir,"Useful_Functions_Tools/")
  
}
if (user == "nicole") {
  
  sci.dir <-    "C:/Users/hillna/OneDrive - University of Tasmania/UTAS_work/Projects/Benthic Diversity ARC/"
  ARC_Data.dir <- paste0(sci.dir,"Analysis/ARC_Data/")
  env.derived <-  paste0(sci.dir,"data_environmental/derived/")
  tools.dir <-    paste0(sci.dir,"Analysis/Useful_Functions_Tools/")
  
}

load(paste0(env.derived,"Circumpolar_Coastline.Rdata"))
bathy<-raster(paste0(env.derived, "Circumpolar_EnvData_500m_shelf_bathy_ibcso2_depth.tif"))

load(paste0(ARC_Data.dir, "annotation/Circumpolar_Annotation_Data_2km.RData"))
load(paste0(ARC_Data.dir, "annotation/Circumpolar_Annotation_Env_Data_2km.RData"))

source(paste0(tools.dir,"bubbleplot.R"))
```


## Image Level Plots

### Temporal (and spatial) distribution of images by years

```{r im year, fig.width=9, fig.height=8}

cover_im<-image_metadata %>%
    filter(cover== "yes")


#knitr::kables( list(
  
  cover_im %>%
  count(., year) %>%
  kbl(caption="No. images by year") %>%
      kable_styling(full_width = F, position ="float_left")
  
cover_im %>%
  count(., year, survey) %>%
  kbl(caption="No. images by year") %>%
      kable_styling(full_width = F, position ="float_right")
#))

#change colour palette, keep and plot NAs
depth_pal<-brewer.pal(n=9, name="Blues")
col_yr<-palette(c(brewer.pal(n = 11, name = "RdYlGn"), "darkgray"))
cover_im$yr_fac<-as.character(cover_im$year)
cover_im$yr_fac[is.na(cover_im$yr_fac)]<-"missing"
cover_im$yr_fac<-as.factor(cover_im$yr_fac)

plot(bathy, col=depth_pal, main="Image by Year")
points(cover_im$proj_coord_x, cover_im$proj_coord_y, col=cover_im$yr_fac)
legend("topright", legend=levels(cover_im$yr_fac),  col=1:length(levels(cover_im$yr_fac)),pch=20, text.col="white", title="Year", bg="#08519C")
  
```

### Spatial distribution of images by Gear

```{r im gear, fig.width=9, fig.height=8}

#knitr:: kables( list(
  
  cover_im %>%
  count(., gear) %>%
  kbl(caption="Images by gear") %>%
      kable_styling(full_width = F, position="float_left")

cover_im %>%
  count(., gear, survey) %>%
  kbl(caption="Images by gear & survey") %>%
      kable_styling(full_width = F, position="float_right")
#))

col_gear<-palette(c("#A50026","#D73027","#F46D43" , "#FDAE61" , "#FEE08B" , "#FFFFBF" , "#D9EF8B" , "#A6D96A" , "#66BD63" , "#1A9850",  "#006837" , "darkgray"))

plot(bathy, col=depth_pal, main="Images by Gear")
points(cover_im$proj_coord_x, cover_im$proj_coord_y, col=as.factor(cover_im$gear))
legend("topright", legend=levels(as.factor(cover_im$gear)), col=1:length(levels(as.factor(cover_im$gear))), pch=20, text.col="white", bg="#08519C", title="Gear")


```

### Distribution of area and image quality by: survey,  platform  , year
```{r im qual, fig.width=7}

cover_im %>%
  ggplot (aes(x= gear, y=image_quality_score))+
  geom_boxplot()+
  theme_bw()+ 
  theme(axis.text.x =element_text(angle=90, hjust=1))+ 
  labs(title="Image Quality vs Gear", x="Gear", y="Quality Score")


cover_im %>%
  ggplot (aes(x= survey, y=image_quality_score))+
  geom_boxplot()+
  theme_bw()+
  theme(axis.text.x =element_text(angle=90, hjust=1))+ 
  labs(title="Image Quality vs Survey", x="Survey", y="Quality Score")

cover_im %>%
  ggplot (aes(x= as.factor(year), y=image_quality_score))+
  geom_boxplot()+
  theme_bw()+
  theme(axis.text.x =element_text(angle=90, hjust=1))+
   labs(title="Image Quality vs Year", x="Year", y="Quality Score")



```


### Check if there are cells with mulitple gear types

```{r cellcheck gear, echo=TRUE}
length(unique(cover_im$cellID))

cover_im %>%
  count(cellID, gear) %>%
  dim(.)

```

number of rows the same so no cells where there are 2 different gear types


### Check if there are cells sampled in different years

```{r cellcheck year, echo=TRUE}
length(unique(cover_im$cellID))

cover_im %>%
  count(cellID, year) %>%
  dim(.)

```
number of rows the same so no cells where there are 2 different sampling years


## Cell level plots

### Images per cell

```{r images_per_cell, fig.width=8}


## At cell level

#COVER
cell_metadata_env %>%
  ggplot(aes(x=cover_N)) +
  geom_histogram(bins=14, col="black")+
  ggtitle("COVER: Images per environmental cell")+
   labs(x="No. images", y="No. cells")


#COUNT

cell_metadata_env %>%
    filter(! is.na (counts_N)) %>%
  ggplot(aes(x=counts_N)) +
  geom_histogram(bins=8, col="black")+
  ggtitle("COUNT: Images per environmental cell")+
  labs(x="No. images", y="No. cells")

```


### Spatial distribution of number of images per cell for COVER data. {.tabset}

#### Ross Sea
```{r Ross_Sea_N, message=FALSE, warning=FALSE, fig.width=9, fig.height=8 }
#plot number of images per cell
#try bubbleplot
sp_im<-SpatialPointsDataFrame(cell_metadata_env[,c("proj_coord_x", "proj_coord_y")], cell_metadata_env,proj4string = crs(bathy))

raster::plot(bathy, xlim=c(-600000,600000), ylim=c(-2650000,-1250000))
plot(coast.proj, add=TRUE)
bubble.plot(sp_im, var = "cover_N", pch=1)

```

#### East Antarctica
```{r East_Ant_N, warning=FALSE, fig.width=8 }
raster::plot( bathy,  xlim=c(2150000,2300000), ylim=c(-1350000,-1250000), main="")
plot(coast.proj, add=TRUE)
bubble.plot(sp_im, var = "cover_N", pch=1)

#raster::plot( env_stack, "depth",  xlim=c(2150000,2300000), ylim=c(-1350000,-1250000), main="")
#plot(coast.proj, add=TRUE)
#  bubble.plot(na.omit(sp_im), var = "counts_N", pch=1, col="red", max.symbol.size=6)
```

#### Antarctic Peninsula
```{r ANt_Pen_N, fig.height=10, fig.width=8 }
raster::plot( bathy, xlim=c(-2750000,-2100000), ylim=c(750000,2410000), main="")
plot(coast.proj, add=TRUE)
bubble.plot(sp_im, var = "cover_N", pch=1)
```

#### Weddell Sea
  
```{r Weddell_N , fig.width=8, fig.height=6 }
raster::plot( bathy, xlim=c(-1400000,600000), ylim=c(800000,2300000), main="")
plot(coast.proj, add=TRUE)
bubble.plot(sp_im, var = "cover_N", pch=1)
```

### Spatial distribution of number of images per cell for COUNT data. {.tabset}

#### Ross Sea
```{r Ross_Sea_count,  message=FALSE, warning=FALSE, fig.width=9, fig.height=8  }
#plot number of images per cell
source(paste0(tools.dir,"bubbleplot.R"))
sp_im_count<-sp_im [!is.na(sp_im$counts_N),]


raster::plot(bathy, xlim=c(-600000,600000), ylim=c(-2650000,-1250000))
plot(coast.proj, add=TRUE)
bubble.plot(sp_im_count, var = "counts_N", pch=1, max.symbol.size=6)

```

#### East Antarctica
```{r East_Ant_count,  warning=FALSE, fig.width=8 }
raster::plot( bathy,  xlim=c(2150000,2300000), ylim=c(-1350000,-1250000), main="")
plot(coast.proj, add=TRUE)
bubble.plot(sp_im_count, var = "counts_N", pch=1, max.symbol.size=6)

#raster::plot( env_stack, "depth",  xlim=c(2150000,2300000), ylim=c(-1350000,-1250000), main="")
#plot(coast.proj, add=TRUE)
#  bubble.plot(na.omit(sp_im), var = "counts_N", pch=1, col="red", max.symbol.size=6)
```

#### Antarctic Peninsula
```{r ANt_Pen_count, fig.width=10, fig.height=8 }
raster::plot( bathy, xlim=c(-2750000,-2100000), ylim=c(750000,2410000), main="")
plot(coast.proj, add=TRUE)
bubble.plot(sp_im_count, var = "counts_N", pch=1, max.symbol.size=6)
```

#### Weddell Sea
  
```{r Weddell_count , fig.width=8, fig.height=6 }
raster::plot( bathy, xlim=c(-1400000,600000), ylim=c(800000,2300000), main="")
plot(coast.proj, add=TRUE)
bubble.plot(sp_im_count, var = "counts_N", pch=1, max.symbol.size=6)
```


## Biological Data
3) Any platform -related effects??  Platform is confounded with survey and, in many cases, region. Not sure how to look at this!
4) GAMS with env data- TO Do once narrowed down species to look at.


### Prevalence of each taxa/ OTU
```{r bio prevalence, fig.width=13}

cover_prev<-data.frame(count=colSums(cover_cells>0)) %>%
  mutate(., prev= count/950 ) %>%
  rownames_to_column(., var="Label") 


cover_prev %>%
  ggplot(aes(x=reorder(Label, -prev), y=prev))+
  geom_bar(stat="identity")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=5))+
   labs(title="All Taxa", x="Label")

```





### Prevalence of Common Taxa (Occurring in >= 2% or 19/950 cells)
```{r common, fig.width=13}
# cut prevalence of common species at 5%  of cells 
  
cover_common<-cover_prev %>%
  filter(prev >= 0.02)

cover_common %>%
  ggplot(aes(x=reorder(Label, -prev), y=prev))+
  geom_bar(stat="identity")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=8))+ 
  labs(title="Cover: Common (> 5%) Taxa", x="Label")  

temp<-cover_common %>%
  arrange(., desc(count)) %>%
  
  #kbl(caption="Cover: Prevalence (proportion of sites) of Common Taxa") %>%
  #    kable_styling(full_width = F)
  mutate(prev= round(prev, 3)) 

      #kbl(caption="Overall abundance") %>%
      #kable_styling(full_width = F)
htmlTable( cbind(temp[1:32,], temp[33:64,]),
          cgroup= c("",""), n.cgroup=c(3,3),
          rnames=FALSE, caption="No. occupied cells and prevalence of Common Taxa")
  
```




#### Prevalence of Rarer species (Occuring at <2% cells) 
```{r prev rare}
  cover_rare<-cover_prev %>%
      filter(prev < 0.02) %>%
      mutate(prev= round(prev, 3)) 

temp<-cover_rare %>%
  arrange(., desc(count))
 #kbl(caption="Cover: Prevalence (proportion of sites) of Rare Taxa") %>%
      #kable_styling(full_width = F)
      
htmlTable( cbind(temp[1:27,], temp[28:54,]),
          cgroup= c("",""), n.cgroup=c(3,3),
rnames=FALSE, caption="No. occupied cells and prevalence of Rarer Taxa")
  

```


### Abundance

Removing 'Unscorable' and calculating overall percentages (across all cells) for all taxa
```{r abund}

perc<-cover_cells_env %>%
  select(., -Unscorable) 

  perc_cell<- cbind(cover_cells_env %>% select (cellID:logslope),
        prop.table(as.matrix(perc %>% select ("1Sub_Fine":"Octocorals_Seapen")), margin=1)*100)
  
  perc_overall<-colSums(prop.table(as.matrix(perc %>% select ("1Sub_Fine":"Octocorals_Seapen")))*100)

perc_overall<- as.data.frame(perc_overall)

temp<-perc_overall%>%
  mutate(Label=rownames(perc_overall)) %>%

  arrange(., desc(perc_overall))%>%
  rownames_to_column(., var="Taxa") %>%
  mutate(perc_overall= round(perc_overall, 3)) 

      #kbl(caption="Overall abundance") %>%
      #kable_styling(full_width = F)
htmlTable( cbind(temp[1:62,], temp[63:124,]),
          cgroup= c("",""), n.cgroup=c(2,2),
rnames=FALSE)

  
```



#### Abundance of key groups and taxa (at individual cell level)

Would be good to colour-code by survey- TO DO

```{r taxa bubbles, fig.width=9, fig.height=8}

taxa_im<-SpatialPointsDataFrame(perc_cell[,c("proj_coord_x", "proj_coord_y")], perc_cell,proj4string = crs(bathy))

raster::plot(bathy, col=depth_pal)
        bubble.plot(taxa_im, var = "1Sub_Fine", pch=1)

raster::plot(bathy, col=depth_pal)
        bubble.plot(taxa_im, var = "2Sub_PbGrv", pch=1)

raster::plot(bathy, col=depth_pal)
        bubble.plot(taxa_im, var = "5Sub_BioRu", pch=1)

raster::plot(bathy, col=depth_pal)
      bubble.plot(taxa_im, var = "Bryozoan_Hard_BrAnt", pch=1)


raster::plot(bathy, col=depth_pal)
      bubble.plot(taxa_im, var = "Echinoderms_Ophiuroids_Other", pch=1)
      
raster::plot(bathy, col=depth_pal)
      bubble.plot(taxa_im, var = "Hydrozoa_Hydrocorals_Branching", pch=1)
      

```
## Raw pearsons correlations between substrate vars and taxa

```{r subs taxa cor, warning=FALSE}
 
subs<-c("1Sub_Fine", "2Sub_PbGrv", "4Sub_Bould", "3Sub_Cbble", "5Sub_BioRu")

sub_cors<-as.data.frame(matrix(ncol = length(subs), nrow = 124))
for(i in 1:length(subs)){
  temp<-round(cor(perc_cell[,subs[i]], perc_cell%>% select(Sub_Fine:S_CupIncmp)),2)
  sub_cors[,i]<-temp[1:124]
}
rownames(sub_cors)<-names(perc_cell%>% select(Sub_Fine:S_CupIncmp))
names(sub_cors)<-subs

sub_cors %>% 
 kbl(caption="Correlation with substrate") %>%
      kable_styling(full_width = F)


```


## Round maps for data-paper

```{r}
library(SOmap)
sci.dir <-      "C:/Users/jjansen/Desktop/science/"
tools.dir <-    paste0(sci.dir,"SouthernOceanBiodiversityMapping/Useful_Functions_Tools/")
source(paste0(tools.dir,"SOmap_functions_JJ.R"))

blues <- rev(blues9)
better.ramp2 <- grDevices::colorRampPalette(blues)
leg.steps <- c(1,2,2.75,3)
lcex <- 1

#### specify colour ramp for percentage cover
g.cols <- rev(heat.colors(10))
```


```{r, warning=FALSE, echo=FALSE, fig.height=10, fig.width=10}
morphotype.dat <- data.frame(taxa_im$`1Sub_Fine`)

## split range of diatom values into 10
val.10 <- cut(morphotype.dat[,1],breaks = 11)
val.range <- range(morphotype.dat[,1])
val.seq <- seq(val.range[1],val.range[2],length.out=11)
val.labels <- round(val.seq,3)
## colour by abundance
morphotype.dat$Col <- g.cols[as.numeric(val.10)]

#pdf(file="Circumpolar_overview_annotations_substrate.pdf", width=10, height=10)
par(oma=c(1,1,1,1), mar=c(3,3,3,3))
JJ_SOmap(ramp.col = better.ramp2, border=FALSE, graticules = TRUE, label = "depth\n(m)", leg.steps=leg.steps, lcex=lcex)

JJ_SOleg(col=g.cols, ticks=length(val.seq), border_width=0, label="%-cover", position="bottomright", tlabs=val.labels, leg.steps=leg.steps, type="continuous")
## add sampling points
points(taxa_im, pch=20, col=morphotype.dat$Col)

scalebar(1000000, type="bar", label=c("0","500","1000"), below="km")
#dev.off()
```

```{r, warning=FALSE, echo=FALSE, fig.height=10, fig.width=10}
toMatch <- c("Sponge","Octocoral","Ascidian","Anemone","Hydr","Bryozoa")
sel <- unique(grep(paste(toMatch,collapse="|"),names(taxa_im), value=TRUE))
morphotype.dat.raw <- data.frame(taxa_im[sel])[,-c(53:55)] ## take care when transforming from a spatialdataframe
morphotype.dat <- data.frame(log(rowSums(morphotype.dat.raw)+1))

## split range of diatom values into 10
val.10 <- cut(morphotype.dat[,1],breaks = 11)
val.range <- range(morphotype.dat[,1])
val.seq <- seq(val.range[1],val.range[2],length.out=11)
val.labels <- round(exp(val.seq),0)
## colour by abundance
morphotype.dat$Col <- g.cols[as.numeric(val.10)]

#pdf(file="Circumpolar_overview_annotations_SF.pdf", width=10, height=10)
par(oma=c(1,1,1,1), mar=c(3,3,3,3))
JJ_SOmap(ramp.col = better.ramp2, border=FALSE, graticules = TRUE, label = "depth\n(m)", leg.steps=leg.steps, lcex=lcex)

JJ_SOleg(col=g.cols, ticks=length(val.seq), border_width=0, label="%-cover", position="bottomright", tlabs=val.labels, leg.steps=leg.steps, type="continuous")
## add sampling points
points(taxa_im, pch=20, col=morphotype.dat$Col)

scalebar(1000000, type="bar", label=c("0","500","1000"), below="km")
#dev.off()
```


