---
title: "Data_explore_Annotations"
author: "Nicole Hill"
date: "04/11/2021"
output: html_document
---

```{r setup, message=FALSE, warning=FALSE,echo=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(raster)
library(rasterVis)
library(tidyverse)
library(classInt)
library(scales)
library(kableExtra)

sci.dir <-      "C:/Users/hillna/OneDrive - University of Tasmania/UTAS_work/Projects/Benthic Diversity ARC/"
env.derived <-  paste0(sci.dir,"data_environmental/derived/")
tools.dir <-    paste0(sci.dir,"Analysis/Useful_Functions_Tools/")
ARC_Data.dir <- paste0(sci.dir,"Analysis/ARC_Data/")

load(paste0(env.derived,"Circumpolar_Coastline.Rdata"))
bathy<-raster(paste0(env.derived, "Circumpolar_EnvData_500m_shelf_bathy_gebco_depth"))

```


## Image data- NOT RUN -CHECK MAP LABELS.

### Temporal (and spatial) distribution of images and gear type

```{r im year}

cover_im<-image_metadata %>%
    filter(cover== "yes")



cover_im %>%
  count(., year) %>%
  kbl(caption="No. images by year") %>%
      kable_styling(full_width = F)

plot(bathy)
points(cover_im$proj_coord_x, cover_im$proj_coord_y, col=as.factor(cover_im$year))
legend("topright", legend=unique(cover_im$year), col=1:length(unique(cover_im$year)), pch=21, title="year")
  
```



```{r im gear}

cover_im %>%
  count(., gear) %>%
  kbl(caption="No. images by gear") %>%
      kable_styling(full_width = F)

plot(bathy)
points(cover_im$proj_coord_x, cover_im$proj_coord_y, col=as.factor(cover_im$gear))
legend("topright", legend=unique(cover_im$gear), col=1:length(unique(cover_im$gear)), pch=21, title="year")


```

### Distribution of area and image quality by: survey,  platform  , year
```{r im qual}

cover_im %>%
  ggplot (aes(x= gear, y=image_quality_score))+
  geom_boxplot()+
  theme_bw()+ 
  theme(axis.text.x =element_text(angle=90, hjust=1))


cover_im %>%
  ggplot (aes(x= survey, y=image_quality_score))+
  geom_boxplot()+
  theme_bw()+
   theme(axis.text.x =element_text(angle=90, hjust=1))


cover_im %>%
  ggplot (aes(x= as.factor(year), y=image_quality_score))+
  geom_boxplot()+
  theme_bw()+
  theme(axis.text.x =element_text(angle=90, hjust=1))



```



####Check if there are cells with mulitple gear types

```{r cellcheck gear, echo=TRUE}
length(unique(cover_im$cellID))

cover_im %>%
  count(cellID, gear) %>%
  dim(.)

```

number of rows the same so no cells where there are 2 different gear types


####Check if there are cells sampled in different years

```{r cellcheck year, echo=TRUE}
length(unique(cover_im$cellID))

cover_im %>%
  count(cellID, year) %>%
  dim(.)

```
number of rows the same so no cells where there are 2 different sampling years


### Images per cell

```{r images_per_cell}
load(paste0(ARC_Data.dir, "annotation/Circumpolar_Annotation_Data.RData"))
load(paste0(ARC_Data.dir, "annotation/Circumpolar_Annotation_Env_Data.RData"))

## At cell level

#COVER
cell_metadata_env %>%
  ggplot(aes(x=cover_N)) +
  geom_histogram(bins=14, col="black")+
  ggtitle("COVER_ No. image per environmental cell")


#COUNT

cell_metadata_env %>%
    filter(! is.na (counts_N)) %>%
  ggplot(aes(x=counts_N)) +
  geom_histogram(bins=8, col="black")+
  ggtitle("COUNT_ No. image per environmental cell")

```


### Spatial distribution of number of images per cell for COVER data. {.tabset}

#### Ross Sea
```{r Ross_Sea_N, message=FALSE, warning=FALSE }
#plot number of images per cell
#try bubbleplot
source(paste0(tools.dir,"bubbleplot.R"))
sp_im<-SpatialPointsDataFrame(cell_metadata_env[,c("proj_coord_x", "proj_coord_y")], cell_metadata_env,proj4string = crs(bathy))

raster::plot(bathy, xlim=c(-600000,600000), ylim=c(-2650000,-1250000))
plot(coast.proj, add=TRUE)
bubble.plot(sp_im, var = "cover_N", pch=1)

```

#### East Antarctica
```{r East_Ant_N, warning=FALSE}
raster::plot( bathy,  xlim=c(2150000,2300000), ylim=c(-1350000,-1250000), main="")
plot(coast.proj, add=TRUE)
bubble.plot(sp_im, var = "cover_N", pch=1)

#raster::plot( env_stack, "depth",  xlim=c(2150000,2300000), ylim=c(-1350000,-1250000), main="")
#plot(coast.proj, add=TRUE)
#  bubble.plot(na.omit(sp_im), var = "counts_N", pch=1, col="red", max.symbol.size=6)
```

#### Antarctic Peninsula
```{r ANt_Pen_N}
raster::plot( bathy, xlim=c(-2750000,-2100000), ylim=c(750000,2410000), main="")
plot(coast.proj, add=TRUE)
bubble.plot(sp_im, var = "cover_N", pch=1)
```

#### Weddell Sea
  
```{r Weddell_N }
raster::plot( bathy, xlim=c(-1400000,600000), ylim=c(800000,2300000), main="")
plot(coast.proj, add=TRUE)
bubble.plot(sp_im, var = "cover_N", pch=1)
```

### Spatial distribution of number of images per cell for COUNT data. {.tabset}

#### Ross Sea
```{r Ross_Sea_count,  message=FALSE, warning=FALSE }
#plot number of images per cell
source(paste0(tools.dir,"bubbleplot.R"))
sp_im_count<-sp_im [!is.na(sp_im$counts_N),]


raster::plot(bathy, xlim=c(-600000,600000), ylim=c(-2650000,-1250000))
plot(coast.proj, add=TRUE)
bubble.plot(sp_im_count, var = "counts_N", pch=1, max.symbol.size=6)

```

#### East Antarctica
```{r East_Ant_count,  warning=FALSE}
raster::plot( bathy,  xlim=c(2150000,2300000), ylim=c(-1350000,-1250000), main="")
plot(coast.proj, add=TRUE)
bubble.plot(sp_im_count, var = "counts_N", pch=1, max.symbol.size=6)

#raster::plot( env_stack, "depth",  xlim=c(2150000,2300000), ylim=c(-1350000,-1250000), main="")
#plot(coast.proj, add=TRUE)
#  bubble.plot(na.omit(sp_im), var = "counts_N", pch=1, col="red", max.symbol.size=6)
```

#### Antarctic Peninsula
```{r ANt_Pen_count}
raster::plot( bathy, xlim=c(-2750000,-2100000), ylim=c(750000,2410000), main="")
plot(coast.proj, add=TRUE)
bubble.plot(sp_im_count, var = "counts_N", pch=1, max.symbol.size=6)
```

#### Weddell Sea
  
```{r Weddell_count }
raster::plot( bathy, xlim=c(-1400000,600000), ylim=c(800000,2300000), main="")
plot(coast.proj, add=TRUE)
bubble.plot(sp_im_count, var = "counts_N", pch=1, max.symbol.size=6)
```


## Biological Data

1) prevalence of each biological taxa (will help inform any aggregation). DONE  
2) key taxa spatial bubbleplots  
3) Any platform -related effects??  Need platform and image level data
4) GAMS with env data- TO Do once narrowed down species to look at.


### Prevalence 
```{r bio prevalence}

cover_prev<-data.frame(count=colSums(cover_cells>0)) %>%
  mutate(., prev= count/950 ) %>%
  rownames_to_column(., var="Label") 


cover_prev %>%
  ggplot(aes(x=reorder(Label, -prev), y=prev))+
  geom_bar(stat="identity")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=5))+
   labs(title="All Taxa", x="Label")

```


#### Common
```{r common}
# cut prevalence at less than 5%  
  
cover_common<-cover_prev %>%
  filter(prev >= 0.05)

cover_common %>%
  ggplot(aes(x=reorder(Label, -prev), y=prev))+
  geom_bar(stat="identity")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=8))+ 
  labs(title="Cover: Common (> 5%) Taxa", x="Label")  

cover_common %>%
  arrange(., desc(prev)) %>%
  kbl(caption="Cover: Common (> 5%) Taxa") %>%
      kable_styling(full_width = F)
  
```


#### Rare 
```{r prev rare}
  cover_rare<-cover_prev %>%
      filter(prev < 0.05) %>%
      mutate(prev= round(prev, 2)) 

cover_rare %>%
  arrange(., desc(prev))%>%
      kbl(caption="Cover: Rare (> 5%) Taxa") %>%
      kable_styling(full_width = F)


```


### Abundance

Removing 'Unscorable' and calculating overall percentages (across all cells)
```{r abund}

perc<-cover_cells_env %>%
  select(., -Unscorable) 

  perc_cell<-cbind(cover_cells_env[,1:58],
        prop.table(as.matrix(perc[,59:182]), margin=1)*100)
  
  perc_overall<-colSums(prop.table(as.matrix(perc[,59:182]))*100)

perc_overall<- as.data.frame(perc_overall)

perc_overall%>%
  mutate(Label=rownames(perc_overall)) %>%

  arrange(., desc(perc_overall))%>%
      kbl(caption="Overall abundance") %>%
      kable_styling(full_width = F)

```

#### abundance of key groups and taxa (at individual cell level)

Would be good to colour-code by survey- TO DO

```{r taxa bubbles}

taxa_im<-SpatialPointsDataFrame(perc_cell[,c("proj_coord_x", "proj_coord_y")], perc_cell,proj4string = crs(bathy))

raster::plot(bathy)
        bubble.plot(taxa_im, var = "Sub_Fine", pch=1)

raster::plot(bathy)
        bubble.plot(taxa_im, var = "Sub_PbGrv", pch=1)

raster::plot(bathy)
        bubble.plot(taxa_im, var = "Sub_BioRu", pch=1)

raster::plot(bathy)
      bubble.plot(taxa_im, var = "BH_BrAnt", pch=1)


raster::plot(bathy)
      bubble.plot(taxa_im, var = "E_O_BritO", pch=1)
      
raster::plot(bathy)
      bubble.plot(taxa_im, var = "HydrC_Br", pch=1)
      

```
##Raw pearsons correlations between substrate vars and taxa

```{r subs taxa cor}
 
subs<-c("Sub_Fine", "Sub_PbGrv", "Sub_Bould", "Sub_Cbble", "Sub_BioRu")

sub_cors<-as.data.frame(matrix(ncol = length(subs), nrow = 123))
for(i in 1:length(subs)){
  temp<-round(cor(perc_cell[,subs[i]], perc_cell[,51:173]),2)
  sub_cors[,i]<-temp[1:123]
}
rownames(sub_cors)<-names(perc_cell[,51:173])
names(sub_cors)<-subs

sub_cors %>% 
 kbl(caption="Correlation with substrate") %>%
      kable_styling(full_width = F)


```



