---
title: "DP190101858 Biologcial Data Overview"
author: "Jan Jansen"
date: "2019-2022"
output: html_document
---

This is an overview of the methods, scripts and code used to generate the biological data for the 2019-2022 ARC-DP  

Main contents:  
* Overview of image sampling locations  
* Overview of Diatom sampling locations  
* Overview of Diatom abundances  

```{r, warning=FALSE, message=FALSE}
library(SOmap)
library(raster)
library(viridis) ## for a nice continuous color palette (viridis or magma)
library(RColorBrewer)
library(graticule)
library(terra)
library(rgdal)
```

### Setting up the environment

```{r}
## Jan's local machine:
sci.dir <-      "C:/Users/jjansen/Desktop/science/"
env.derived <-  paste0(sci.dir,"data_environmental/derived/")
bio.dir <-      paste0(sci.dir,"data_biological/")

## remote repository (DOESN'T WORK YET):
# env.dir <- "https://data.imas.utas.edu.au/data_transfer/admin/files/EnvironmentalData/"

## common paths (after "sci.dir")
tools.dir <-    paste0(sci.dir,"SouthernOceanBiodiversityMapping/Useful_Functions_Tools/")
ARC_Data.dir <- paste0(sci.dir,"SouthernOceanBiodiversityMapping/ARC_Data/")
```

```{r}
source(paste0(tools.dir,"SOmap_functions_JJ.R"))
```

```{r}
##### load still and diatom sample locations and bathymetry:

## from "Readin_Circumpolar_DownwardImage_Data.Rmd"
load(paste0(ARC_Data.dir,"prep_image/Circumpolar_DownwardImages_metadata.Rdata"))

## from "Readin_Circumpolar_Diatom_Data.Rmd"
load(paste0(ARC_Data.dir,"FAM/Circumpolar_Diatom_metadata.Rdata"))

## from "ReadIn_Circumpolar_Environmental_Data.Rmd"
r2 <- raster(paste0(env.derived,"Circumpolar_EnvData_500m_shelf_bathy_ibcso2_depth.tif"))

##### load coastline
#stereo <- as.character(r2@crs)
load(paste0(env.derived,"Circumpolar_Coastline.Rdata"))
```

```{r, eval=FALSE}
##### load shelfline
# shelf_contour <- vect(readOGR(paste0(env.derived,"Antarctic_shelf_1500m_outline/Antarctic_shelf_1500m_outline.shp")))
# shelf_contour.proj <- terra::project(shelf_contour, "+proj=stere +lat_0=-90 +lat_ts=-71 +lon_0=0 +x_0=0 +y_0=0 +datum=WGS84 +units=m +no_defs")
# r2.t <- rast(r2)
# test1 <- terra::mask(r2.t, shelf_contour.proj)
#shelf_contour_proj <- readOGR(paste0(env.derived,"Antarctic_shelf_1500m_outline/Antarctic_shelf_1500m_outline_polygons.shp"))
#shelf_r2 <- mask(r2, shelf_contour_proj)

ra.stack <- rast(paste0(env.derived,"Circumpolar_EnvData_500m_shelf_mask_unscaled_variables.tif"))
shelf.area <- ra.stack$depth
shelf.area[shelf.area[]<(-1500)] <- NA

length(which(!is.na(shelf.area[])))#12808075
length(which(shelf.area[]<(-200)))#12134133
12134133/12808075
mean(shelf.area[], na.rm=TRUE) # something like 570m
```



```{r, echo=FALSE}
##### set colour parameters
## disrete colours:
leg.cols <- brewer.pal(12,"Paired")
wap_cols <- leg.cols[3:11]#leg.cols[c(4,8,11,6)]

## set continuous colors for raster:
# depth.pal <- colorRampPalette(c('black','grey10','grey20','navy','blue','dodgerblue','lightcyan'))
# range.pal <- colorRampPalette(c('black','grey10','grey20','navy','blue','dodgerblue','skyblue1','cadetblue1','aquamarine1','palegreen2','green2','OliveDrab1','yellow1','wheat1'))
#better.ramp2 <- grDevices::colorRampPalette(c("#54A3D1", "#54A3D1", "#54A3D1", "#54A3D1", "#54A3D1", "#54A3D1", "#60B3EB", "#60B3EB", "#78C8F0", "#98D1F5", "#B5DCFF", "#BDE1F0", "#CDEBFA", "#D6EFFF", "#EBFAFF")) #, "grey99", "grey90", "grey92", "grey94", "grey96", "white"
#blues <- c(blues9[9],blues9[9],rev(blues9))
blues <- rev(blues9)

## choose depending on the bathymetry file used (whether deeper than -3000 is NA or not)
better.ramp2 <- grDevices::colorRampPalette(blues)
cols <- better.ramp2(99)
breaks <- seq(-3000,0,length.out=100)
leg.args <- seq(-3000,0, by=500)

#full.ramp <- grDevices::colorRampPalette(c("black","grey15","grey25",blues))
#cols <- full.ramp(132)
#breaks <- c(seq(-7500,-3000,length.out=34),seq(-2500,0,length.out=100))
#leg.args <- seq(-7500,0, by=500)


West.cols.AP <- brewer.pal(9,"Set1")
West.cols.WS <- brewer.pal(5,"Set1")[-2]
East.cols <- brewer.pal(7,"Set1")[-2]

West.cols.leg.AP <- West.cols.AP
West.cols.leg.WS <- c(West.cols.WS,rep("white",5))
East.cols.leg <- c(rep("white",3),East.cols)

West.cruise.AP <- c("CRS","JR17001","LMG1311","PS81","PS61","JR17003","PS118","JR262","JR15005")
West.cruise.WS <- c("PS96","PS06","PS14","PS18","","","","","")
East.cruise <- c("","","","NBP1402","AA2011","NBP1502","TAN0802","TAN1802","TAN1901")

## points plotted in the sequence of objects in names(polar.dat.list.clean)
col.v <- c(West.cols.AP[1],West.cols.WS[2:4],West.cols.AP[c(5,4,4)],West.cols.WS[1],West.cols.AP[7],East.cols[c(4:6,1,3,2)],West.cols.AP[c(8:9,2,6,3)])

pch <- 20
cex <- 2

text.plot <- function(transects,dat,polar.dat,col="black", adj=1.5, cex=0.7){
  for(i in 1:length(transects)){
    pts <- polar.dat[dat$transectID==transects[i]]
    text(pts[1],labels=transects[i], adj=adj, cex=cex,col=col)
  }}
leg.steps <- c(1,2,2.75,3)

lcex <- 1
```

## Image Data

### Surveys, metadata and image subset

*	GER data: PS06, PS14, PS18, PS61, PS81, PS96, PS118 (Ant. Peninsula to Lazarev Sea)
*	US data:
    + WAP (Craig Smith)
    + LMG1311 (Weat Ant. Peninsula)
    + NBP1402 (Sabrina Shelf, East Antarctica)
    + NBP1502 (Ross Sea)
* UK data: JR262, JR15005, JR17001, JR17003 (Tip of Ant. Pensinsula & South Orkneys)
* NZ data: TAN0802, TAN1802, TAN1901 (Ross Sea)
* AUS data: AA2011 (Mertz region)
* ITA data: never received due to COVID19-restrictions preventing access

Each survey has their own script to read in raw data and create an image subset. The scripts are in the folder _ReadIn_Circumpolar_DownwardImage_Data_And_Subset_Surveys_ and are named:  
__Readin_Circumpolar_DownwardImages_...R__  

Summary file containing all survey data:  
__Readin_Circumpolar_DownwardImage_Data.Rmd__  

Overview script containing plots for each individual transect + locations of the image subset:  
__DP190101858_Circumpolar_Data_Overview_Transects.Rmd__

Image annotation on CoralNet (every image of the subset) and Biigle (the first 50% of the subset images)

Image-Annotation Quality Control and Final Library:  
__AnnotationLibrary_counts.R__  
__AnnotationQualityControl_CoralNet_SurveyBySurvey.R__  


```{r}
a <- length(levels(dat.list.clean[[1]]$transectID))
for(i in 2:length(dat.list.clean)){
  a <- a + length(levels(dat.list.clean[[i]]$transectID))
}
message(paste0(a," different sampling transects"))
```


#### Overview of sampling locations

```{r, warning=FALSE, echo=FALSE, fig.height=10, fig.width=10}
#### Image survey sites on bathymetry
#pdf(file="Circumpolar_overview_images.pdf", width=10, height=10)
par(oma=c(1,1,1,1), mar=c(3,3,3,3))
JJ_SOmap(ramp.col = better.ramp2, border=FALSE, graticules = TRUE, label = "depth\n(m)", leg.steps=leg.steps, lcex=lcex)
JJ_SOleg(col=West.cols.leg.AP, ticks=length(West.cols.leg.AP), border_width=0, lcex=lcex, leg.steps=leg.steps,
         label="Survey IDs\nAntarctic Pensinsula", position = "topleft",tlabs=West.cruise.AP)
JJ_SOleg(col=West.cols.leg.WS, ticks=length(West.cols.WS), border_width=0, lcex=lcex, leg.steps=leg.steps,
         label="Survey IDs\nWeddell and Lazarev Sea", position = "topright",tlabs=West.cruise.WS)
JJ_SOleg(col=East.cols.leg, ticks=length(East.cols), border_width=0, lcex=lcex, leg.steps=leg.steps,
         label="Survey IDs\nEast Antarctic", position = "bottomright",tlabs=East.cruise)
for(i in 1:length(polar.dat.list.clean)){
  plot(polar.dat.list.clean[[i]], add=TRUE, pch=pch, cex=cex, col=col.v[i])
}
scalebar(1000000, type="bar", label=c("0","500","1000"), below="km")
#dev.off()
```

#### Overview of SHALLOW (<300m) sampling locations

```{r}
xy.sel.list <- list(1:2,6:7,6:7,6:7,6:7,
                    1:2,1:2,1:2,1:2,1:2,
                    1:2,1:2,1:2,c(4,3),c(7,6),
                    5:6,5:6,5:6,5:6,5:6)
```


```{r, warning=FALSE, echo=FALSE, fig.height=10, fig.width=10}
#### Image survey sites on bathymetry
lcex=1.5
pdf(file="DECRA_Circumpolar_overview_images.pdf", width=10, height=10)
par(oma=c(1,1,1,1), mar=c(3,3,3,3))
JJ_SOmap(ramp.col = better.ramp2, border=FALSE, graticules = TRUE, label = "depth\n(m)", leg.steps=leg.steps, lcex=lcex)
for(i in 1:length(polar.dat.list.clean)){
  plot(polar.dat.list.clean[[i]], add=TRUE, pch=pch, cex=cex, col="black")
  
  xy.dat2 <- as.data.frame(proj4::project(dat.subset.list[[i]][,1:2], proj=crs(r2)))
  depth <- extract(r2,xy.dat2)
  sel <- which(depth>=-200)
  sel.t <- which(dat.list.clean[[i]]$transectID%in%as.character(unique(dat.subset.list[[i]]$transectID[sel])))
  plot(polar.dat.list.clean[[i]][sel.t,], add=TRUE, pch=pch, cex=cex, col="purple")
  
  sel2 <- which(depth>=-150)
  sel2.t <- which(dat.list.clean[[i]]$transectID%in%as.character(unique(dat.subset.list[[i]]$transectID[sel2])))
  plot(polar.dat.list.clean[[i]][sel2.t,], add=TRUE, pch=pch, cex=cex, col="red")
}
JJ_SOleg(col=c("white","white","white","red","purple","black"), ticks=2, border_width=0, lcex=lcex, leg.steps=leg.steps, label="AS-AID images", position = "bottomright",tlabs=c("","","","< 150m depth","150-200m depth","> 200m depth"))

scalebar(1000000, type="bar", label=c("0","500","1000"), below="km")
dev.off()
```

### Region specific overviews {.tabset}

#### Ross Sea - overview
```{r, warning=FALSE, echo=FALSE, fig.height=10, fig.width=9}
## add points for NBP1502:

### for adding graticules:
# ## specify plotting variables:
# lons <- seq(-180, 180, by = 10)
# lats <- seq(-79, -65, by = 2)
# xl <-  range(lons)# + c(-0.4, 0.4)
# yl <- range(lats) + c(-0.4, 0.4)
# grat <- graticule(lons, lats, proj = stereo, ylim = yl, xlim = xl)
# labs <- graticule_labels(lons, lats, xline = min(xl), yline = max(yl), proj = stereo) #
## the lines are a SpatialLinesDataFrame
#plot(grat, add = TRUE, lty = 5, col = rgb(0, 0, 0, 0.8))
## the labels are a SpatialPointsDataFrame, and islon tells us which kind
#text(subset(labs, labs$islon), lab = parse(text = labs$lab[labs$islon]), pos = 3)
#text(subset(labs, !labs$islon), lab = parse(text = labs$lab[!labs$islon]), pos = 2)

plot(r2, xlim=c(-600000,600000), ylim=c(-2650000,-1250000),col=cols, main="Ross Sea", axes=FALSE, breaks=breaks, legend=FALSE)
#plot(r2, xlim=c(-600000,600000), ylim=c(-2650000,-1250000),col=cols, main="Ross Sea", axes=FALSE)
plot(coast.proj, add=TRUE)
scalebar(200000, type="bar", label=c("0","100","200"), below="km")
points(polar.dat.start.list$TAN0802, pch=pch, col=East.cols[4], cex=cex)
points(polar.dat.start.list$TAN1802, pch=pch, col=East.cols[5], cex=cex)
points(polar.dat.start.list$TAN1901, pch=pch, col=East.cols[6], cex=cex)
points(polar.dat.NBP1502, pch=pch, col=East.cols[3], cex=cex)
legend("left", pch=pch, pt.cex=cex, col=East.cols[c(4:6,3)], legend=c("TAN0802","TAN1802","TAN1901","NBP1502"))
plot(r2,legend.only=TRUE,col=cols, breaks=breaks, axis.args=list(at=leg.args,labels=leg.args))
```

#### Ross Sea - zoom

```{r, warning=FALSE, echo=FALSE, fig.width=10, fig.height=8.9}
## shelf/slope only
plot(r2, xlim=c(-430000,480000), ylim=c(-2130000,-1330000),col=cols, main="Ross Sea - shelf & slope only", axes=FALSE, breaks=breaks, legend=FALSE)
plot(coast.proj, add=TRUE)
scalebar(200000, type="bar", label=c("0","100","200"), below="km")
points(polar.dat.start.list$TAN0802, pch=pch, col=East.cols[4], cex=cex)
points(polar.dat.start.list$TAN1802, pch=pch, col=East.cols[5], cex=cex)
points(polar.dat.start.list$TAN1901, pch=pch, col=East.cols[6], cex=cex)
points(polar.dat.NBP1502, pch=pch, col=East.cols[3], cex=cex)
legend("left", pch=pch, pt.cex=cex, col=East.cols[c(4:6,3)], legend=c("TAN0802","TAN1802","TAN1901","NBP1502"))
text.plot(levels(dat.list$TAN0802$transectID),dat.start.list$TAN0802,polar.dat.start.list$TAN0802, col=East.cols[4])
text.plot(levels(dat.list$TAN1802$transectID),dat.start.list$TAN1802,polar.dat.start.list$TAN1802,col=East.cols[5], adj=-0.7)
text.plot(levels(dat.list$TAN1901$transectID),dat.start.list$TAN1901,polar.dat.start.list$TAN1901,col=East.cols[6])
text(polar.dat.NBP1502,labels=dat.list$NBP1502$transectID, adj=1.5, cex=0.7,col=East.cols[3])
plot(r2,legend.only=TRUE,col=cols, breaks=breaks, axis.args=list(at=leg.args,labels=leg.args))
```


```{r, warning=FALSE, echo=FALSE, fig.width=10, fig.height=7}
## shelf/slope only
plot(r2, xlim=c(-430000,50000), ylim=c(-1630000,-1330000),col=cols, main="Ross Sea - shelf & slope only", axes=FALSE, breaks=breaks, legend=FALSE)
plot(coast.proj, add=TRUE)
scalebar(100000, type="bar", label=c("0","50","100"), below="km")
points(polar.dat.start.list$TAN0802, pch=pch, col=East.cols[4], cex=cex)
points(polar.dat.start.list$TAN1802, pch=pch, col=East.cols[5], cex=cex)
points(polar.dat.start.list$TAN1901, pch=pch, col=East.cols[6], cex=cex)
points(polar.dat.NBP1502, pch=pch, col=East.cols[3], cex=cex)
legend("left", pch=pch, pt.cex=cex, col=East.cols[c(4:6,3)], legend=c("TAN0802","TAN1802","TAN1901","NBP1502"))
text.plot(levels(dat.list$TAN0802$transectID),dat.start.list$TAN0802,polar.dat.start.list$TAN0802, col=East.cols[4])
text.plot(levels(dat.list$TAN1802$transectID),dat.start.list$TAN1802,polar.dat.start.list$TAN1802,col=East.cols[5], adj=-0.7)
text.plot(levels(dat.list$TAN1901$transectID),dat.start.list$TAN1901,polar.dat.start.list$TAN1901,col=East.cols[6])
text(polar.dat.NBP1502,labels=dat.list$NBP1502$transectID, adj=1.5, cex=0.7,col=East.cols[3])
plot(r2,legend.only=TRUE,col=cols, breaks=breaks, axis.args=list(at=leg.args,labels=leg.args))
```


```{r, warning=FALSE, echo=FALSE, fig.width=6.8, fig.height=4.7}
## seamounts only
plot(r2, xlim=c(-250000,430000), ylim=c(-2620000,-2200000),col=cols, main="Ross Sea - seamounts", axes=FALSE, breaks=breaks, legend=FALSE)
plot(coast.proj, add=TRUE)
scalebar(100000, xy=c(0,-2590000), type="bar", label=c("0","50","100"), below="km")
points(polar.dat.start.list$TAN0802, pch=pch, col=East.cols[4], cex=cex)
points(polar.dat.start.list$TAN1802, pch=pch, col=East.cols[5], cex=cex)
points(polar.dat.start.list$TAN1901, pch=pch, col=East.cols[6], cex=cex)
legend("topright",pch=pch, pt.cex=cex, col=East.cols[c(4:6,3)], legend=c("TAN0802","TAN1802","TAN1901"))
text.plot(levels(dat.list$TAN0802$transectID),dat.start.list$TAN0802,polar.dat.start.list$TAN0802, col=East.cols[4])
text.plot(levels(dat.list$TAN1802$transectID),dat.start.list$TAN1802,polar.dat.start.list$TAN1802,col=East.cols[5], adj=-0.7)
text.plot(levels(dat.list$TAN1901$transectID),dat.start.list$TAN1901,polar.dat.start.list$TAN1901,col=East.cols[6])
plot(r2,legend.only=TRUE,col=cols, breaks=breaks, axis.args=list(at=leg.args,labels=leg.args))
```

```{r,echo=FALSE, eval=FALSE}
## these transect need to be analysed
idx.0802 <- which(polar.dat.0802@coords[,2]>-2200000)
idx.1802 <- which(polar.dat.1802@coords[,2]>-2200000)
idx.1901 <- which(polar.dat.1901@coords[,2]>-2200000)
analyse.in.0802 <- as.character(unique(dat.0802$transectID[idx.0802]))
analyse.in.1802 <- as.character(unique(dat.1802$transectID[idx.1802]))
analyse.in.1901 <- as.character(unique(dat.1901$transectID[idx.1901]))
length(c(analyse.in.0802,analyse.in.1802,analyse.in.1901))
## disregard these transect for now:
not.analyse.in.0802 <- as.character(unique(dat.0802$transectID[which(polar.dat.0802@coords[,2]<=-2200000)]))
not.analyse.in.1802 <- as.character(unique(dat.1802$transectID[which(polar.dat.1802@coords[,2]<=-2200000)]))
not.analyse.in.1901 <- as.character(unique(dat.1901$transectID[which(polar.dat.1901@coords[,2]<=-2200000)]))
length(c(not.analyse.in.0802,not.analyse.in.1802,not.analyse.in.1901))
## depth distribution of relevant transects
depth.0802 <- extract(r2,polar.dat.0802[idx.0802,])
depth.1802 <- extract(r2,polar.dat.1802[idx.1802,])
depth.1901 <- extract(r2,polar.dat.1901[idx.1901,])
plot(c(depth.0802,depth.1802,depth.1901))
```


#### Mertz

```{r, warning=FALSE, echo=FALSE, fig.width=10, fig.height=4.5}
plot(r2, xlim=c(700000,1700000), ylim=c(-2350000,-2000000),col=cols, main="Mertz", axes=FALSE, breaks=breaks, legend=FALSE)
plot(coast.proj, add=TRUE)
scalebar(20000, type="bar", label=c("0","10","20"), below="km")
points(polar.dat.start.list$AA2011 , pch=pch, col=East.cols[1], cex=cex)
text.plot(levels(dat.list$NBP1402$transectID),dat.start.list$NBP1402,polar.dat.start.list$NBP1402, col=East.cols[1])
plot(r2,legend.only=TRUE,col=cols, breaks=breaks, axis.args=list(at=leg.args,labels=leg.args))
```

#### Sabrina Coast

```{r, warning=FALSE, echo=FALSE, fig.width=6, fig.height=4.5}
plot(r2, xlim=c(2150000,2300000), ylim=c(-1350000,-1250000),col=cols, main="Sabrina Coast", axes=FALSE, breaks=breaks, legend=FALSE)
plot(coast.proj, add=TRUE)
scalebar(20000, type="bar", label=c("0","10","20"), below="km")
points(polar.dat.start.list$NBP1402 , pch=pch, col=East.cols[1], cex=cex)
text.plot(levels(dat.list$NBP1402$transectID),dat.start.list$NBP1402,polar.dat.start.list$NBP1402, col=East.cols[1])
plot(r2,legend.only=TRUE,col=cols, breaks=breaks, axis.args=list(at=leg.args,labels=leg.args))
```

#### AP - overview
```{r, warning=FALSE, echo=FALSE, fig.width=10, fig.height=20}
# library(readxl)
# ANA08D.raw <- read.csv(paste0(bio.dir,"Stills/ANA08D/ANA08D_yoyo_elog.csv"),header=TRUE)
# ANA08D <- ANA08D.raw[-c(1:3),c(1:3,7:9)]
# names(ANA08D)[3:5] <- c("transectID","lat","lon")
# ANA08D$transectID <- as.factor(ANA08D$transectID)
# 
# spatial.dat <- data.frame(cbind(ANA08D$lon, ANA08D$lat))
# names(spatial.dat) <- c("Longitude","Latitude")
# coordinates(spatial.dat) <- c("Longitude","Latitude")
# proj4string(spatial.dat) <- CRS("+proj=longlat +datum=WGS84")
# polar.dat.ANA08D <- spTransform(spatial.dat, CRS(stereo))

plot(r2, xlim=c(-2750000,-2100000), ylim=c(750000,2410000),col=cols, main="Antarctic Peninsula", axes=FALSE)
plot(coast.proj, add=TRUE)
scalebar(200000, type="bar", label=c("0","100","200"), below="km")
points(polar.dat.start.list$WAP, pch=pch, cex=cex, col=wap_cols[1])
points(polar.dat.start.list$PS61, pch=pch, cex=cex, col=wap_cols[2])
points(polar.dat.start.list$PS81, pch=pch, cex=cex, col=wap_cols[3])
points(polar.dat.start.list$PS118, pch=pch, cex=cex, col=wap_cols[4])
points(polar.dat.start.list$JR262, pch=pch, cex=cex, col=wap_cols[5])
points(polar.dat.start.list$JR15005, pch=pch, cex=cex, col=wap_cols[6])
points(polar.dat.start.list$JR17001, pch=pch, cex=cex, col=wap_cols[7])
points(polar.dat.start.list$JR17003, pch=pch, cex=cex, col=wap_cols[8])
points(polar.dat.start.list$LMG1311, pch=pch, cex=cex, col=wap_cols[9])
# points(polar.dat.ANA08D,pch=pch, cex=cex,col="pink")
legend("topleft",pch=pch, pt.cex=cex, col=wap_cols, legend=c("WAP","PS61","PS81","PS118","JR262","JR15005","JR17001","JR17003","LMG1311"))
# text.plot(levels(dat.list$WAP$transectID),dat.start.list$WAP,polar.dat.start.list$WAP, col=wap_cols[1])
# text.plot(levels(dat.list$PS61$transectID),dat.start.list$PS61,polar.dat.start.list$PS61, col=wap_cols[2])
# text.plot(levels(dat.list$PS81$transectID),dat.start.list$PS81,polar.dat.start.list$PS81, col=wap_cols[3])
# text.plot(levels(dat.list$PS118$transectID),dat.start.list$PS118,polar.dat.start.list$PS118, col=wap_cols[4])
# text.plot(levels(dat.list$JR262$transectID),dat.start.list$JR262,polar.dat.start.list$JR262, col=wap_cols[5])
# text.plot(levels(dat.list$JR15005$transectID),dat.start.list$JR15005,polar.dat.start.list$JR15005, col=wap_cols[6])
# text.plot(levels(dat.list$JR17001$transectID),dat.start.list$JR17001,polar.dat.start.list$JR17001, col=wap_cols[7])
# text.plot(levels(dat.list$JR17003$transectID),dat.start.list$JR17003,polar.dat.start.list$JR17003, col=wap_cols[8])
# text.plot(levels(dat.list$LMG1311$transectID),dat.start.list$LMG1311,polar.dat.start.list$LMG1311, col=wap_cols[9])
#text(polar.dat.ANA08D,labels=ANA08D$transectID, adj=1.5, cex=0.7,col="pink")
#plot(r2,legend.only=TRUE,col=cols, breaks=breaks, axis.args=list(at=leg.args,labels=leg.args))
```

#### AP - zoom

```{r, warning=FALSE, echo=FALSE, fig.width=10, fig.height=9}
plot(r2, xlim=c(-2550000,-2100000), ylim=c(2000000,2420000), main="Antarctic Peninsula - South Orkney", axes=FALSE, col=cols)#cols, breaks=breaks, legend=FALSE
plot(coast.proj, add=TRUE)
scalebar(20000, type="bar", label=c("0","10","20"), below="km")
points(polar.dat.start.list$PS61, pch=pch, cex=cex, col=wap_cols[2])
points(polar.dat.start.list$PS118, pch=pch, cex=cex, col=wap_cols[4])
points(polar.dat.start.list$JR262, pch=pch, cex=cex, col=wap_cols[5])
points(polar.dat.start.list$JR15005, pch=pch, cex=cex, col=wap_cols[6])
legend("topleft",pch=pch, pt.cex=cex, col=wap_cols[c(2,4:6)], legend=c("PS61","PS118","JR262","JR15005"))
# text.plot(levels(dat.list$PS61$transectID),dat.start.list$PS61,polar.dat.start.list$PS61, col=wap_cols[2])
# text.plot(levels(dat.list$PS118$transectID),dat.start.list$PS118,polar.dat.start.list$PS118, col=wap_cols[4])
# text.plot(levels(dat.list$JR262$transectID),dat.start.list$JR262,polar.dat.start.list$JR262, col=wap_cols[5])
# text.plot(levels(dat.list$JR15005$transectID),dat.start.list$JR15005,polar.dat.start.list$JR15005, col=wap_cols[6])
#plot(r2,legend.only=TRUE,col=cols, breaks=breaks, axis.args=list(at=leg.args,labels=leg.args))
```

```{r, warning=FALSE, echo=FALSE, fig.width=10, fig.height=10}
plot(r2, xlim=c(-2750000,-2300000), ylim=c(1400000,1900000), main="Antarctic Peninsula - Tip", axes=FALSE, col=cols)#cols, breaks=breaks, legend=FALSE
plot(coast.proj, add=TRUE)
scalebar(20000, type="bar", label=c("0","10","20"), below="km")
points(polar.dat.start.list$WAP, pch=pch, cex=cex, col=wap_cols[1])
points(polar.dat.start.list$PS61, pch=pch, cex=cex, col=wap_cols[2])
points(polar.dat.start.list$PS81, pch=pch, cex=cex, col=wap_cols[3])
points(polar.dat.start.list$PS118, pch=pch, cex=cex, col=wap_cols[4])
points(polar.dat.start.list$JR17001, pch=pch, cex=cex, col=wap_cols[7])
points(polar.dat.start.list$JR17003, pch=pch, cex=cex, col=wap_cols[8])
legend("topleft",pch=pch, pt.cex=cex, col=wap_cols[c(1:4,7,8)], legend=c("WAP","PS61","PS81","PS118","JR17001","JR17003"))
#text.plot(levels(dat.list$WAP$transectID),dat.start.list$WAP,polar.dat.start.list$WAP, col=wap_cols[1])
# text.plot(levels(dat.list$PS61$transectID),dat.start.list$PS61,polar.dat.start.list$PS61, col=wap_cols[2])
# text.plot(levels(dat.list$PS81$transectID),dat.start.list$PS81,polar.dat.start.list$PS81, col=wap_cols[3])
# text.plot(levels(dat.list$PS118$transectID),dat.start.list$PS118,polar.dat.start.list$PS118, col=wap_cols[4])
# text.plot(levels(dat.list$JR17001$transectID),dat.start.list$JR17001,polar.dat.start.list$JR17001, col=wap_cols[7])
# text.plot(levels(dat.list$JR17003$transectID),dat.start.list$JR17003,polar.dat.start.list$JR17003, col=wap_cols[8])
#plot(r2,legend.only=TRUE,col=cols, breaks=breaks, axis.args=list(at=leg.args,labels=leg.args))
```



```{r, warning=FALSE, echo=FALSE, fig.width=5, fig.height=9.2}
plot(r2, xlim=c(-2600000,-2375000), ylim=c(800000,1400000), main="Antarctic Peninsula - lower", axes=FALSE, col=cols)#cols, breaks=breaks, legend=FALSE
plot(coast.proj, add=TRUE)
scalebar(20000, type="bar", label=c("0","10","20"), below="km")
points(polar.dat.start.list$WAP, pch=pch, cex=cex, col=wap_cols[1])
points(polar.dat.start.list$JR17001, pch=pch, cex=cex, col=wap_cols[7])
points(polar.dat.start.list$LMG1311, pch=pch, cex=cex, col=wap_cols[9])
legend("topright",pch=pch, pt.cex=cex, col=wap_cols[c(1,7,9)], legend=c("WAP","JR17001","LMG1311"))
# text.plot(levels(dat.list$WAP$transectID),dat.start.list$WAP,polar.dat.start.list$WAP, col=wap_cols[1])
# text.plot(levels(dat.list$JR17001$transectID),dat.start.list$JR17001,polar.dat.start.list$JR17001, col=wap_cols[7])
# text.plot(levels(dat.list$LMG1311$transectID),dat.start.list$LMG1311,polar.dat.start.list$LMG1311, col=wap_cols[9])
#plot(r2,legend.only=TRUE,col=cols, breaks=breaks, axis.args=list(at=leg.args,labels=leg.args))
```

#### Weddell Sea
```{r echo=FALSE, fig.height=7.8, fig.width=10, warning=FALSE}
plot(r2, xlim=c(-1400000,600000), ylim=c(800000,2300000),col=cols, main="Weddell Sea", axes=FALSE)#, breaks=breaks, legend=FALSE)
plot(coast.proj, add=TRUE)
scalebar(200000, xy=c(0,900000), type="bar", label=c("0","100","200"), below="km")
points(polar.dat.PS06, pch=pch, cex=cex, col=wap_cols[1])
points(polar.dat.PS14, pch=pch, cex=cex, col=wap_cols[2])
points(polar.dat.PS18, pch=pch, cex=cex, col=wap_cols[3])
points(polar.dat.PS96, pch=pch, cex=cex, col=wap_cols[4])
legend("bottomright",pch=pch, pt.cex=cex, col=wap_cols, legend=c("PS06","PS14","PS18","PS96"))
text.plot(levels(dat.list$PS06$transectID),dat.list$PS06,polar.dat.PS06, col=wap_cols[1])
text.plot(levels(dat.list$PS14$transectID),dat.list$PS14,polar.dat.PS14,col=wap_cols[2], adj=-0.7)
text.plot(levels(dat.list$PS18$transectID),dat.list$PS18,polar.dat.PS18,col=wap_cols[3])
text.plot(levels(dat.list$PS96$transectID),dat.list$PS96,polar.dat.PS96,col=wap_cols[4])
#plot(r2,legend.only=TRUE,col=cols, breaks=breaks, axis.args=list(at=leg.args,labels=leg.args))
```

## Diatom Data - Sampling locations

```{r, message=FALSE, warning=FALSE}
#### Diatom survey sites on bathymetry (excluding CEAMARC sites, and splitting AP from the rest for plotting)
ceamarc <- which(diatom.dat$Cruise=="CEAMARC")
AP_cruises <- which(diatom.dat$Longitude<0&diatom.dat$Longitude>-90)
diatom.ceamarc <- diatom.dat[ceamarc,]
diatom.AP <- diatom.dat[AP_cruises,]
diatom.rest <- diatom.dat[-c(AP_cruises,ceamarc),]

diatom.AP$Cruise <- factor(diatom.AP$Cruise)
diatom.rest$Cruise <- factor(diatom.rest$Cruise)

AP.cols <- c(brewer.pal(12,"Paired"),"black")[c(3:13,1,2)]
rest.cols <- brewer.pal(8,"Set1")
rest.cols.lab <- c(rep("white",4),brewer.pal(8,"Set1")[c(3,5,8,6,1,7,2,4)])
t.labs.rest <- c("","","","",levels(diatom.rest$Cruise)[c(3,5,8,6,1,7,2,4)])
```

#### Overview

```{r, fig.width=10, fig.height=10, warning=FALSE}
#pdf(file="Circumpolar_overview_diatoms.pdf", width=10, height=10)
par(oma=c(0,0,0,0))
## main map
JJ_SOmap(ramp.col = better.ramp2, border=FALSE, graticules = TRUE, label = "depth\n(m)", leg.steps=leg.steps)
## legend top left with surveys from that area
JJ_SOleg(col=AP.cols, ticks = length(AP.cols), border_width = 0, label="Survey IDs\nAntarctic Peninsula", position = "topleft",tlabs=levels(diatom.AP$Cruise), leg.steps=leg.steps)
## legend bottom rigth with surveys from that area
JJ_SOleg(col=rest.cols.lab, ticks = length(rest.cols), border_width = 0, label="Survey IDs\nother regions", position = "bottomright",tlabs=t.labs.rest, leg.steps=leg.steps)
## add sampling points
for(i in 1:length(levels(diatom.AP$Cruise))){
  plot(polar.dat.diatom[diatom.dat$Cruise==levels(diatom.AP$Cruise)[i]], add=TRUE, pch=pch, cex=cex, col=AP.cols[i])
}
for(i in 1:length(levels(diatom.rest$Cruise))){
  plot(polar.dat.diatom[diatom.dat$Cruise==levels(diatom.rest$Cruise)[i]], add=TRUE, pch=pch, cex=cex, col=rest.cols[i])
}
## and a scalebar to get a sense of distance
scalebar(1000000, type="bar", label=c("0","500","1000"), below="km")
#dev.off()
```



### Region specific overviews {.tabset}

#### Ross Sea
```{r, warning=FALSE, echo=FALSE, fig.width=10, fig.height=7.1}
plot(r2, xlim=c(-800000,480000), ylim=c(-2000000,-1200000),col=cols, main="Ross Sea", axes=FALSE, breaks=breaks, legend=FALSE)
plot(coast.proj, add=TRUE)
scalebar(200000, type="bar", label=c("0","100","200"), below="km", col="white")
for(i in 1:length(levels(diatom.rest$Cruise))){
  plot(polar.dat.diatom[diatom.dat$Cruise==levels(diatom.rest$Cruise)[i]], add=TRUE, pch=pch, cex=cex, col=rest.cols[i])
}
legend("left", pch=pch, pt.cex=cex, col=rest.cols[-c(3,5)], legend=levels(diatom.rest$Cruise)[-c(3,5)])

plot(r2,legend.only=TRUE,col=cols, breaks=breaks, axis.args=list(at=leg.args,labels=leg.args))
```


#### Amundsen Sea
```{r, warning=FALSE, echo=FALSE, fig.width=10, fig.height=10}
plot(r2, xlim=c(-2100000,-1100000), ylim=c(-1350000,-300000),col=cols, main="Amundsen Sea", axes=FALSE, breaks=breaks, legend=FALSE)
plot(coast.proj, add=TRUE)
scalebar(200000, type="bar", label=c("0","100","200"), below="km", col="white")
for(i in 1:length(levels(diatom.rest$Cruise))){
  plot(polar.dat.diatom[diatom.dat$Cruise==levels(diatom.rest$Cruise)[i]], add=TRUE, pch=pch, cex=cex, col=rest.cols[i])
}
legend("topright", pch=pch, pt.cex=cex, col=rest.cols[4], legend=levels(diatom.rest$Cruise)[4])
plot(r2,legend.only=TRUE,col=cols, breaks=breaks, axis.args=list(at=leg.args,labels=leg.args))
```


#### East Antarctica 1
```{r, warning=FALSE, echo=FALSE, fig.width=7, fig.height=5.5}
plot(r2, xlim=c(1350000,1750000), ylim=c(-2200000,-1900000),col=cols, main="East Antarctica - CEAMARC", axes=FALSE, breaks=breaks, legend=FALSE)
plot(coast.proj, add=TRUE)
scalebar(100000, xy=c(1600000,-2150000), type="bar", label=c("0","50","100"), below="km", col="white")
for(i in 1:length(levels(diatom.rest$Cruise))){
  plot(polar.dat.diatom[diatom.dat$Cruise==levels(diatom.rest$Cruise)[i]], add=TRUE, pch=pch, cex=cex, col=rest.cols[i])
}
legend("topleft", pch=pch, pt.cex=cex, col=rest.cols[3], legend=levels(diatom.rest$Cruise)[3])

plot(r2,legend.only=TRUE,col=cols, breaks=breaks, axis.args=list(at=leg.args,labels=leg.args))
```


#### East Antarctica 2
```{r, warning=FALSE, echo=FALSE, fig.width=7, fig.height=5}
plot(r2, xlim=c(2000000,2400000), ylim=c(-1400000,-1100000),col=cols, main="East Antarctica - Totten area", axes=FALSE, breaks=breaks, legend=FALSE)
plot(coast.proj, add=TRUE)
scalebar(100000, type="bar", label=c("0","50","100"), below="km")
for(i in 1:length(levels(diatom.rest$Cruise))){
  plot(polar.dat.diatom[diatom.dat$Cruise==levels(diatom.rest$Cruise)[i]], add=TRUE, pch=pch, cex=cex, col=rest.cols[i])
}
legend("topleft", pch=pch, pt.cex=cex, col=rest.cols[5], legend=levels(diatom.rest$Cruise)[5])

plot(r2,legend.only=TRUE,col=cols, breaks=breaks, axis.args=list(at=leg.args,labels=leg.args))
```


#### East Antarctica 3
```{r, warning=FALSE, echo=FALSE, fig.width=5.8, fig.height=10}
plot(r2, xlim=c(2000000,2800000), ylim=c(-300000,1500000),col=cols, main="East Antarctica - Amery area", axes=FALSE, breaks=breaks, legend=FALSE)
plot(coast.proj, add=TRUE)
scalebar(200000, type="bar", label=c("0","100","200"), below="km")
for(i in 1:length(levels(diatom.rest$Cruise))){
  plot(polar.dat.diatom[diatom.dat$Cruise==levels(diatom.rest$Cruise)[i]], add=TRUE, pch=pch, cex=cex, col=rest.cols[i])
}
legend("topright", pch=pch, pt.cex=cex, col=rest.cols[3], legend=levels(diatom.rest$Cruise)[3])

plot(r2,legend.only=TRUE,col=cols, breaks=breaks, axis.args=list(at=leg.args,labels=leg.args))
```

#### AP
```{r, warning=FALSE, echo=FALSE, fig.width=10, fig.height=13.5}
plot(r2, xlim=c(-2750000,-2000000), ylim=c(700000,1850000),col=cols, main="Antarctic Peninsula", axes=FALSE, breaks=breaks, legend=FALSE)
plot(coast.proj, add=TRUE)
scalebar(200000, type="bar", label=c("0","100","200"), below="km")

for(i in 1:length(levels(diatom.AP$Cruise))){
  plot(polar.dat.diatom[diatom.dat$Cruise==levels(diatom.AP$Cruise)[i]], add=TRUE, pch=pch, cex=cex, col=AP.cols[i])
}

legend("topright",pch=pch, pt.cex=cex, col=AP.cols, legend=levels(diatom.AP$Cruise))

# text.plot(levels(dat.CRS$transectID),dat.start.list$WAP,polar.dat.start.list$WAP, col=wap_cols[1])
# text.plot(levels(dat.PS61$TransectID),dat.start.list$PS61,polar.dat.start.list$PS61, col=wap_cols[2])
# text.plot(levels(dat.PS81$transectID),dat.start.list$PS81,polar.dat.start.list$PS81, col=wap_cols[3])
# text.plot(levels(dat.PS118$transectID),dat.start.list$PS118,polar.dat.start.list$PS118, col=wap_cols[4])
plot(r2,legend.only=TRUE,col=cols, breaks=breaks, axis.args=list(at=leg.args,labels=leg.args))
```

## Diatom Data - Abundances

```{r}
## specify colour gradient
g.cols <- rev(heat.colors(10))
## split range of diatom values into 10
val.10 <- cut(log(diatom.dat$Absolute_diatom_abundance_in_million_valves_per_gram_dry_sediment),breaks = 10)
val.range <- range(log(diatom.dat[,7]))
val.seq <- seq(val.range[1],val.range[2],length.out=length(g.cols))
val.labels <- round(exp(val.seq),3)

## colour by abundance of diatoms
diatom.dat$Col <- g.cols[as.numeric(val.10)]
```

#### Overview

```{r, fig.width=10, fig.height=10, warning=FALSE}
#pdf(file="Circumpolar_overview_diatoms_abundances.pdf", width=10, height=10)
par(oma=c(0,0,0,0))
## main map
JJ_SOmap(ramp.col=better.ramp2, border=FALSE, graticules=TRUE, label="depth\n(m)", leg.steps=leg.steps)
## legend top right with surveys from that area
JJ_SOleg(col=g.cols, ticks=length(val.seq), border_width=0, label="Diatoms\n(millions/gram)", position="bottomright", tlabs=val.labels, leg.steps=leg.steps, type="continuous")
## add sampling points
points(polar.dat.diatom, pch=20, col=diatom.dat$Col)
## and a scalebar to get a sense of distance
scalebar(1000000, type="bar", label=c("0","500","1000"), below="km")
#dev.off()
```

### Region specific overviews {.tabset}

```{r}
## setup legend
leg.vals <- round(exp(seq(min(log(diatom.dat[,7])),max(log(diatom.dat[,7])),length.out=10)),2)
leg.title <- "counts in million"
pt.cex <- 2
pt.pch <- 20
main <- ""

## specify plotting variables:
lons <- seq(-70, -50, by = 10)
lats <- seq(-75, -50, by = 2)
xl <-  range(lons) + c(-0.4, 0.4)
yl <- range(lats) + c(-0.4, 0.4)
grat <- graticule(lons, lats, proj = stereo, ylim = yl, xlim = xl)
labs <- graticule_labels(lons, lats, xline = min(xl), yline = max(yl), proj = stereo) #
```

#### Antarctic Peninsula
```{r, fig.height=10, fig.width=8, echo=FALSE}
plot(r2, xlim=c(-2800000,-1800000), ylim=c(600000,2000000),col=cols, main=main)
plot(coast.proj, add=TRUE)
scalebar(200000, type="bar", label=c("0","100","200"), below="km")
points(polar.dat.diatom, col=diatom.dat$Col, cex=pt.cex, pch=pt.pch)
legend("topright",title=leg.title, legend=leg.vals, col=g.cols, pt.cex=pt.cex, pch=pt.pch)
```

```{r, fig.height=10, fig.width=8, echo=FALSE}
plot(r2, xlim=c(-2800000,-1800000), ylim=c(600000,2000000),col=cols, main=main, axes=FALSE)
plot(coast.proj, add=TRUE)
scalebar(200000, type="bar", label=c("0","100","200"), below="km")
points(polar.dat.diatom, col=diatom.dat$Col, cex=pt.cex, pch=pt.pch)
## the lines are a SpatialLinesDataFrame
plot(grat, add = TRUE, lty = 5, col = rgb(0, 0, 0, 0.8))
## the labels are a SpatialPointsDataFrame, and islon tells us which kind
text(subset(labs, labs$islon), lab = parse(text = labs$lab[labs$islon]), pos = 3)
text(subset(labs, !labs$islon), lab = parse(text = labs$lab[!labs$islon]), pos = 2)

legend("topright",title=leg.title, legend=leg.vals, col=g.cols, pt.cex=pt.cex, pch=pt.pch)
```


##### Tip only
```{r, fig.height=7.5, fig.width=8, echo=FALSE}
plot(r2, xlim=c(-2800000,-2200000), ylim=c(1100000,1700000),col=cols, main=main)
plot(coast.proj, add=TRUE)
scalebar(200000, type="bar", label=c("0","100","200"), below="km")
points(polar.dat.diatom, col=diatom.dat$Col, cex=pt.cex, pch=pt.pch)
legend("topright",title=leg.title, legend=leg.vals, col=g.cols, pt.cex=pt.cex, pch=pt.pch)
```

#### Amundsen
```{r, fig.height=7, fig.width=8, echo=FALSE}
plot(r2, xlim=c(-2100000,-1000000), ylim=c(-1300000,-300000),col=cols, main=main)
plot(coast.proj, add=TRUE)
scalebar(200000, type="bar", label=c("0","100","200"), below="km")
points(polar.dat.diatom, col=diatom.dat$Col, cex=pt.cex, pch=pt.pch)
legend("topright",title=leg.title, legend=leg.vals, col=g.cols, pt.cex=pt.cex, pch=pt.pch)
```

#### Ross Sea
```{r, fig.height=7, fig.width=10, echo=FALSE}
plot(r2, xlim=c(-800000,500000), ylim=c(-2000000,-1200000),col=cols, main=main)
plot(coast.proj, add=TRUE)
scalebar(200000, type="bar", label=c("0","100","200"), below="km")
points(polar.dat.diatom, col=diatom.dat$Col, cex=pt.cex, pch=pt.pch)
legend("left",title=leg.title, legend=leg.vals, col=g.cols, pt.cex=pt.cex, pch=pt.pch)
```

##### Ross Sea - including sample IDs
```{r, fig.height=7, fig.width=10, echo=FALSE}
plot(r2, xlim=c(-800000,500000), ylim=c(-2000000,-1200000),col=cols, main=main)
plot(coast.proj, add=TRUE)
scalebar(200000, type="bar", label=c("0","100","200"), below="km")
points(polar.dat.diatom, col=diatom.dat$Col, cex=pt.cex, pch=pt.pch)
text(polar.dat.diatom, labels=diatom.dat$SampleID, adj=c(0,-1))
legend("left",title=leg.title, legend=leg.vals, col=g.cols, pt.cex=pt.cex, pch=pt.pch)
```

#### Sabrina Coast & CEAMARC
```{r, fig.height=7, fig.width=7, echo=FALSE}
plot(r2, xlim=c(1300000,2500000), ylim=c(-2300000,-1100000),col=cols, main=main)
plot(coast.proj, add=TRUE)
scalebar(200000, type="bar", label=c("0","100","200"), below="km",xy=c(2300000,-2200000))
points(polar.dat.diatom, col=diatom.dat$Col, cex=pt.cex, pch=pt.pch)
legend("topleft",title=leg.title, legend=leg.vals, col=g.cols, pt.cex=pt.cex, pch=pt.pch)
```

```{r, fig.height=6, fig.width=10, echo=FALSE}
plot(r2, xlim=c(1370000,1800000), ylim=c(-2200000,-1950000),col=cols, main=main)
plot(coast.proj, add=TRUE)
scalebar(200000, type="bar", label=c("0","100","200"), below="km",xy=c(2300000,-2200000))
points(polar.dat.diatom, col=diatom.dat$Col, cex=pt.cex, pch=pt.pch)
text(polar.dat.diatom, labels=diatom.dat$SampleID, adj=c(0,-1),cex=0.5)
legend("topleft",title=leg.title, legend=leg.vals, col=g.cols, pt.cex=pt.cex, pch=pt.pch)
```

#### East-Antarctic
```{r, fig.height=10, fig.width=6, echo=FALSE}
plot(r2, xlim=c(2000000,3000000), ylim=c(-400000,1500000),col=cols, main=main)
plot(coast.proj, add=TRUE)
scalebar(200000, type="bar", label=c("0","100","200"), below="km")
points(polar.dat.diatom, col=diatom.dat$Col, cex=pt.cex, pch=pt.pch)
legend("topright",title=leg.title, legend=leg.vals, col=g.cols, pt.cex=pt.cex, pch=pt.pch)
```















```{r, warning=FALSE, echo=FALSE, eval=FALSE}
## for Nicole / Weddell Sea MPA
dat <- polar.dat.subset.list$PS18[which(dat.subset.list$PS18$lon>=0),]

library(graticule)
lons <- seq(0, 30, length = 7)
lats <- seq(-71, -69, length = 2)
prj <- crs(r2)
## optionally, specify the extents of the meridians and parallels
## here we push them out a little on each side
xl <-  range(lons) + c(-0.4, 0.4)
yl <- range(lats) + c(-0.4, 0.4)
## build the lines with our precise locations and ranges
grat <- graticule(lons, lats, proj = prj, xlim = xl, ylim = yl)
labs <- graticule_labels(lons, lats, xline = min(xl), yline = max(yl), proj = prj)
```

```{r echo=FALSE, eval=FALSE, fig.height=7, fig.width=10, warning=FALSE}
lg.col <- "red"
plot(r2, xlim=c(-50000,500000), ylim=c(2000000,2400000),col=cols, main="", axes=FALSE)#, breaks=breaks, legend=FALSE)
plot(coast.proj, add=TRUE)
scalebar(200000, xy=c(0,900000), type="bar", label=c("0","100","200"), below="km")
points(polar.dat.PS18, pch=pch, cex=cex, col=lg.col)
legend("topleft",pch=pch, pt.cex=cex, col=lg.col, legend="PS18")
text.plot(levels(dat.list$PS18$transectID),dat.list$PS18,polar.dat.PS18)#,col=lg.col)
plot(grat, add=TRUE, lty=2)
text(subset(labs, labs$islon), lab = parse(text = labs$lab[labs$islon]), pos = 3)
text(subset(labs, !labs$islon), lab = parse(text = labs$lab[!labs$islon]), pos = 2)
```

```{r echo=FALSE, eval=FALSE, fig.height=10, fig.width=10, warning=FALSE}
lg.col <- "red"
lons <- seq(0, 30, length = 4)
lats <- c(seq(-75, -65, length = 6),-60)
prj <- crs(r2)
## optionally, specify the extents of the meridians and parallels
## here we push them out a little on each side
xl <-  range(lons) + c(-0.4, 0.4)
yl <- range(lats) + c(-0.4, 0.4)
## build the lines with our precise locations and ranges
grat <- graticule(lons, lats, proj = prj, xlim = xl, ylim = yl)
labs <- graticule_labels(lons, lats, xline = min(xl), yline = max(yl), proj = prj)

#png("WeddellSeaMPA_Phase2_BenthicImagery.png", width=700, height=700)
pdf("WeddellSeaMPA_Phase2_BenthicImagery.pdf", width=9, height=9)
plot(r2, xlim=c(-200000,1800000), ylim=c(1500000,3900000),col=cols, main="", axes=FALSE)#, breaks=breaks, legend=FALSE)
plot(coast.proj, add=TRUE)
scalebar(200000, xy=c(0,900000), type="bar", label=c("0","100","200"), below="km")
points(polar.dat.PS18, pch=pch, cex=1.5, col=lg.col)
legend("bottomright",pch=pch, pt.cex=1.5, col=lg.col, legend="PS18")
#text.plot(levels(dat.list$PS18$transectID),dat.list$PS18,polar.dat.PS18)#,col=lg.col)
plot(grat, add=TRUE, lty=2)
text(subset(labs, labs$islon), lab = parse(text = labs$lab[labs$islon]), pos = 3)
text(subset(labs, !labs$islon), lab = parse(text = labs$lab[!labs$islon]), pos = 2)
dev.off()
```


```{r, warning=FALSE, echo=FALSE, fig.height=10, fig.width=10, eval=FALSE}
## for Fanny:

F.cruise <- c("PS81","PS96","TAN0802","TAN1802","TAN1901")
F.cols <- brewer.pal(6,"Set1")[-2]
F.cols.pts <- c(brewer.pal(6,"Set1")[1],brewer.pal(6,"Set1")[-2])
F.list <- polar.dat.list.clean[c(6:8,10:12)]

leg.steps <- c(1,2,2.75,3)
#### Image survey sites on bathymetry
#pdf(file="Circumpolar_overview_images_forFanny.pdf", width=10, height=10)
par(oma=c(0,0,0,0))
JJ_SOmap(ramp.col = better.ramp2, border=FALSE, graticules = TRUE, label = "depth\n(m)", leg.steps=leg.steps)
JJ_SOleg(col=F.cols, ticks=length(F.cols), border_width=0, label="Survey IDs", position="topright", tlabs=F.cruise, leg.steps=leg.steps)
for(i in 1:length(F.list)){
  plot(F.list[[i]], add=TRUE, pch=pch, cex=cex, col=F.cols.pts[i])
}
scalebar(1000000, type="bar", label=c("0","500","1000"), below="km")
#dev.off()
```


