---
title: "display biodiversity maps"
output: html_document
---

```{r}
# Load packages
library(terra)
library(viridis)
# Convenience operator
'%!in%' <- function(x, y) !('%in%'(x, y))
# Load shared directory definitions

usr <- "VM"
source(file.path("..", "0_SourceFile.R"))

# Resolution label used in environmental and predictive-map filenames
res <- "2km"
# Select one model to display in this report
model_id <- "fam_cafe"

# Environmental raster directory
env.derived <- file.path(usr.dropbox.dir, "data_environmental/derived")
# Optional tools directory (only needed later for SOmap plotting)
tools.dir <- file.path(usr.dropbox.dir, "SouthernOceanBiodiversityMapping/Useful_Functions_Tools")
# Predictive-map root directory
# Using the local copied version for faster raster IO
pred_base_dir <- file.path(usr.main.dir, "4_model_prediction/copy_of_predictive_maps/circum_antarctic")
# Directory containing predictive maps for the selected model
model_dir <- file.path(pred_base_dir, model_id)
# Load environmental raster stack for plotting context / extent / CRS
r.stack <- rast(file.path(env.derived, paste0("Circumpolar_EnvData_", res, "_shelf_mask_unscaled_variables.tif")))
r2 <- r.stack$depth
```


```{r}
# MCMC settings retained for file naming context only
thin <- 10
samples <- 800
nChains <- 4

# Selected species index for species-level map panels
# Change this number to inspect a different morphospecies
species_index <- 19
```


```{r}
# List all species presence/absence raster files for the selected model
pa_files <- list.files(
  model_dir,
  pattern = "^PA_.*\\.tif$",
  full.names = TRUE
)

# Load only the layers needed for this report:
# layer 1 = mean
# layer 2 = median
# layer 3 = standard error
pa_mean   <- rast(pa_files, lyrs = 1)
pa_median <- rast(pa_files, lyrs = 2)
pa_se     <- rast(pa_files, lyrs = 3)

# Calculate richness as the sum of species occurrence probabilities
# Here the median predictions are used for richness display
hmsc.maps.richness <- sum(pa_median, na.rm = TRUE)

# Load total abundance raster and extract useful layers
abundance_file <- file.path(model_dir, "total_abundance.tif")
abundance_maps <- rast(abundance_file)

abundance_mean   <- abundance_maps[[1]]
abundance_median <- abundance_maps[[2]]
abundance_se     <- abundance_maps[[3]]
```

## p/a maps {.tabset}

### p/a - overview

```{r, fig.width=10, fig.height=10}
# Plot predicted richness (sum of species median occurrence probabilities)
plot(hmsc.maps.richness, col = viridis(99))
```

### p/a - individual species

```{r, fig.width=13, fig.height=7}
# Display one species at a time:
# left panel = median probability of presence
# right panel = standard error of the prediction
par(mfrow = c(1, 2), oma = c(2, 0, 2, 3))
mar <- c(1, 1, 1, 0)

for (i in seq_len(nlyr(pa_median))) {
  plot(pa_median[[i]],
       main = names(pa_median)[i],
       mar = mar,
       legend = FALSE,
       col = viridis(99))
  
  plot(pa_se[[i]],
       main = paste0(names(pa_median)[i], "_se"),
       mar = mar,
       col = viridis(99))
}
```


### p/a uncertainty hotspots

requires more thinking about how to detect the hotspots  
essentially need to look at where a standardised prediction error is the smallest, because larger predictions will have larger errors... And for p/a data small and large probabilities will have small errors, so can't simply divide occurrence by the error...  

```{r, eval=FALSE, echo=FALSE, fig.width=13, fig.height=7}
# Example uncertainty exploration for one selected species
par(mfrow = c(1, 2), oma = c(2, 0, 2, 3))
mar <- c(1, 1, 1, 0)

m <- pa_median[[species_index]]
s <- pa_se[[species_index]]

# A simple proportional uncertainty metric
uncertainty_map <- m / s
plot(m, main = names(pa_median)[species_index], mar = mar, legend = FALSE, col = viridis(99))
plot(uncertainty_map, main = "proportional uncertainty", mar = mar, col = viridis(99))

# Alternative scaling of uncertainty
par(mfrow = c(1, 1), oma = c(2, 0, 2, 3))
uncertainty_map2 <- m / (s^2)

plot(uncertainty_map2,
     main = "proportional uncertainty",
     mar = mar,
     col = viridis(99))

# Example hotspot-style thresholds for probability and uncertainty
# These remain exploratory and are not part of the main pipeline
unc.map2 <- uncertainty_map
unc.map2[!is.na(uncertainty_map[])] <- 0
sel.uncert.top1 <- selectHighest(uncertainty_map, 50000)
unc.map2[sel.uncert.top1] <- 1

prob.map2 <- m
prob.map2[!is.na(m[])] <- 0
sel.prob.top1 <- selectHighest(m, 50000)
prob.map2[sel.prob.top1] <- 1

par(mfrow = c(1, 1))
plot(unc.map2, main = "top certainty cells")
plot(prob.map2, main = "top probability-of-presence cells")
```


```{r, eval=FALSE, echo=FALSE, fig.width=13, fig.height=7}
# Repeat the exploratory uncertainty analysis for a different species
# Change species_index here if a different morphospecies is of interest
m <- pa_median[[species_index]]
s <- pa_se[[species_index]]

par(mfrow = c(1, 2), oma = c(2, 0, 2, 3))
mar <- c(1, 1, 1, 0)

uncertainty_map <- m / s
plot(m, main = names(pa_median)[species_index], mar = mar, legend = FALSE, col = viridis(99))

plot(uncertainty_map, main = "proportional uncertainty", mar = mar, col = viridis(99))

par(mfrow = c(1, 1), oma = c(2, 0, 2, 3))
uncertainty_map <- m / (s^2)
plot(uncertainty_map, main = "proportional uncertainty", mar = mar, col = viridis(99))

# Identify cells with high occurrence probability and low uncertainty
unc.map2 <- s
unc.map2[!is.na(s[])] <- 0
sel.uncert.top1 <- selectHighest(s, 200000, low = TRUE)
unc.map2[sel.uncert.top1] <- 1

prob.map2 <- m
prob.map2[!is.na(m[])] <- 0
sel.prob.top1 <- selectHighest(m, 200000)
prob.map2[sel.prob.top1] <- 1

par(mfrow = c(1, 1))
plot(unc.map2, main = "highest-certainty cells")
plot(prob.map2, main = "highest-probability cells")
plot(unc.map2 + prob.map2, main = "overlap")
```

### p/a - regional snapshots

#### Antarctic Peninsula

```{r, fig.height=8, fig.width=10}
maxcell=10000000
xlim=c(-2800000,-1800000)
ylim=c(1350000,2350000)
plot(hmsc.maps.richness, xlim=xlim, ylim=ylim, maxcell=maxcell, col=viridis(99))
```

#### Ross Sea

```{r, fig.height=8, fig.width=10}
maxcell=10000000
xlim=c(-800000,500000)
ylim=c(-2350000,-1250000)
plot(hmsc.maps.richness, xlim=xlim, ylim=ylim, maxcell=maxcell, col=viridis(99))
```

### p/a - East Antarctica - individual species

```{r, eval=FALSE}
# Load custom Southern Ocean mapping utilities if needed
source(file.path(tools.dir, "SOmap_functions_JJ.R"))

# JJ_SOmap expects a RasterLayer object, so convert from terra if needed
library(raster)
JJ_SOmap(env.var.raster = raster(pa_median[[species_index]]),
  ramp.col = viridis(99), border = FALSE, graticules = TRUE, label = "probability of presence"
)
```


```{r, fig.width=7, fig.height=10}
# East Antarctica zoom for one selected species:
# top panel = median probability of presence
# bottom panel = standard error
maxcell <- 10000000
xlim <- c(0, 2500000)
ylim <- c(-2400000, -1000000)

par(mfrow = c(2, 1), bg = NA)

for (i in species_index) {
  plot(pa_median[[i]], xlim = xlim, ylim = ylim, maxcell = maxcell, main = names(pa_median)[i], mar = mar,
       legend = FALSE, col = viridis(99))
  
  plot(pa_se[[i]], xlim = xlim, ylim = ylim, maxcell = maxcell, main = paste0(names(pa_median)[i], "_se"),
       mar = mar, col = viridis(99))
}
```

```{r, fig.width=7, fig.height=10}
# Export a richness zoom for East Antarctica
maxcell <- 10000000
xlim <- c(1000000, 3000000)
ylim <- c(-2400000, 1600000)

png(file.path(getwd(), "Circumpolar_richness_zoom.png"),
    width = 2400, height = 2400, res = 300)

par(mfrow = c(1, 1), bg = NA)

plot(hmsc.maps.richness, xlim = xlim, ylim = ylim, maxcell = maxcell, mar = mar,
     col = viridis(99), axes = FALSE)

dev.off()
```
