---
title: "display circumpolar biodiversity maps"
output:
  html_document:
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = FALSE, dev = "png",
  fig.height = 10, fig.width = 10, dpi = 300, fig.path = "display_predictive_maps_figures/"
)

############################################################
# USER OPTIONS
############################################################
# Toggle writing figures to file
pngs <- FALSE
pdfs <- FALSE

# Spatial resolution
res <- "2km"

# Select the model folder to display in this report
# Examples:
#   "fam_cafe", "fam_cbpm", "fam_eppl", "fam_vpmg",
#   "npp_cafe", "npp_cbpm", "npp_eppl", "npp_vpmg"
model_id <- "fam_cafe"

# Plot export settings
png.height <- 2400
png.width  <- 2400
png.res    <- 300
pdf.height <- 8
pdf.width  <- 8


############################################################
# LOAD SHARED PROJECT SETUP FILES
############################################################
# Look for 0_SourceFile.R and 0_Functions.R either in the current folder or one level above. 
source_file <- if (file.exists("0_SourceFile.R")) {"0_SourceFile.R"
} else if (file.exists(file.path("..", "0_SourceFile.R"))) {file.path("..", "0_SourceFile.R")
} else {stop("Could not find 0_SourceFile.R in the current or parent directory.")
}
functions_file <- if (file.exists("0_Functions.R")) {"0_Functions.R"
} else if (file.exists(file.path("..", "0_Functions.R"))) {file.path("..", "0_Functions.R")
} else {stop("Could not find 0_Functions.R in the current or parent directory.")
}
source(source_file)
source(functions_file)

############################################################
# PACKAGES
############################################################
library(Hmsc)
library(terra)
library(SOmap)
library(viridis)
library(knitr)
library(RColorBrewer)
library(vioplot)
'%!in%' <- function(x, y) !('%in%'(x, y))

############################################################
# DIRECTORY DEFINITIONS
############################################################
# Environmental rasters
env.derived <- file.path(usr.dropbox.dir, "data_environmental/derived")

# Biological data
bio.dir <- file.path(usr.main.dir, "data_biological")

# Existing project folders used elsewhere in this report
tools.dir    <- file.path(usr.dropbox.dir, "SouthernOceanBiodiversityMapping/Useful_Functions_Tools")
DP.dir       <- file.path(usr.dropbox.dir, "DP190101858_MappingAntarcticSeafloorBiodiversity/")
ARC_Data.dir <- file.path(usr.dropbox.dir, "SouthernOceanBiodiversityMapping/ARC_Data/")

# Local copy of predictive maps for fast raster IO
pred.copy.base <- file.path(usr.main.dir,"4_model_prediction/copy_of_predictive_maps/circum_antarctic")

# Model-specific directory containing species maps and total abundance
pred.model.dir <- file.path(pred.copy.base, model_id)

# Directory for derived hotspot / percentile outputs
# change this line if those files are stored in a subfolder (e.g. "hotspots")
derived.dir <- pred.model.dir

# Directory for any figure exports created by this Rmd
pred.dir <- file.path(getwd(), "display_predictive_maps_outputs", model_id, "")
dir.create(pred.dir, recursive = TRUE, showWarnings = FALSE)


############################################################
# BASIC CHECKS
############################################################
if (!dir.exists(pred.model.dir)) {stop("Predictive map directory not found: ", pred.model.dir)}
if (!dir.exists(derived.dir)) {stop("Derived output directory not found: ", derived.dir)}

############################################################
# LOAD ENVIRONMENTAL CONTEXT RASTERS
############################################################
r.stack <- rast(file.path(env.derived, paste0("Circumpolar_EnvData_", res, "_shelf_mask_unscaled_variables.tif")))
r2 <- r.stack$depth
# r2[1:350, ] <- NA
# r.stack.subset <- r.stack[[c(1:3, 14, 26, 42:45, 51, 56)]]
# rm(r.stack)

############################################################
# LOAD SPECIES-LEVEL PREDICTIVE MAPS
############################################################

# Load all PA raster files for the selected model
pa.files <- list.files(
  pred.model.dir,
  pattern = "^PA_.*\\.tif$",
  full.names = TRUE
)

if (length(pa.files) == 0) {
  stop("No PA_*.tif files found in: ", pred.model.dir)
}

# Load only the layers needed in this document:
# 1 = mean, 2 = median, 3 = standard error
hmsc.maps.pa.env.mean   <- rast(pa.files, lyrs = 1)
hmsc.maps.pa.env.median <- rast(pa.files, lyrs = 2)
hmsc.maps.pa.env.se     <- rast(pa.files, lyrs = 3)

# Richness reconstructed directly from species median predictions
# This keeps the same object name used later in the report.
hmsc.maps.richness.env <- rast(sum(hmsc.maps.pa.env.median, na.rm = TRUE))
names(hmsc.maps.richness.env) <- "median"


############################################################
# LOAD TOTAL ABUNDANCE
############################################################

abundance.file <- file.path(pred.model.dir, "total_abundance.tif")

if (!file.exists(abundance.file)) {
  stop("Could not find total_abundance.tif in: ", pred.model.dir)
}

hmsc.maps.totalabundance.env <- rast(abundance.file)

# Keep the same convenience layer name used later in the report
# by capping median abundance at 100
hmsc.maps.totalabundance.env$median.100 <- hmsc.maps.totalabundance.env[[2]]
values(hmsc.maps.totalabundance.env$median.100)[
  values(hmsc.maps.totalabundance.env[[2]]) > 100
] <- 100


############################################################
# LOAD ADDITIONAL DERIVED PRODUCTS
############################################################

# The following files were used in the original workflow and are kept
# with the same filenames. Only the root directory has been updated.
# If these files live in a different subfolder, update 'derived.dir' above.

# Richness / abundance percentiles
perc.richness.median  <- rast(file.path(derived.dir, "percentiles_richness_median.tif"))
perc.abundance.median <- rast(file.path(derived.dir, "percentiles_abundance_median.tif"))
perc.rich_per_abund   <- rast(file.path(derived.dir, "percentiles_richness_per_abundance.tif"))

# Domain-level percentiles
perc.richness.domains  <- rast(file.path(derived.dir, "percentiles_richness_domains.tif"))
perc.abundance.domains <- rast(file.path(derived.dir, "percentiles_abundance_domains.tif"))

# Combined hotspot rasters
top.percentiles         <- rast(file.path(derived.dir, "top_percentiles.tif"))
top.percentiles.domains <- rast(file.path(derived.dir, "top_percentiles_per_domains.tif"))

# Beta-diversity and bootstrapped richness products
median.betadiv    <- rast(file.path(derived.dir, "median_BetaDiversity_LCBD.tif"))
boots.rich.median <- rast(file.path(derived.dir, "bootstrapped_median.tif"))
boots.rich.sd     <- rast(file.path(derived.dir, "bootstrapped_sd.tif"))

# Beta-diversity percentiles
perc.betadiv.median <- rast(file.path(derived.dir, "median_BetaDiversity_LCBD_percentiles.tif"))


############################################################
# MATCH NA STRUCTURE ACROSS DERIVED RASTERS
############################################################

ra.na.sel <- which(is.na(hmsc.maps.richness.env$median[]))

median.betadiv[ra.na.sel]    <- NA
boots.rich.median[ra.na.sel] <- NA
boots.rich.sd[ra.na.sel]     <- NA


############################################################
# LOAD SAMPLE LOCATIONS
############################################################

load(file = paste0(ARC_Data.dir, "Image_level_bio.Rdata"))


############################################################
# LOAD COASTLINE
############################################################

coast.unprojected <- vect(
  paste0(
    usr.dropbox.dir,
    "data_environmental/antarctic_coastline_2023/add_coastline_high_res_polygon_v7_8.shp"
  )
)

coast.proj <- project(coast.unprojected, r2)


############################################################
# SETUP COLOURS
############################################################

better.ramp <- grDevices::colorRampPalette(rev(blues9))
better.ramp <- viridis
lcex <- 1

# Colours for discrete hotspot categories
perc.ramp <- grDevices::colorRampPalette(c("grey90", "grey65", rev(viridis(3))))
cols <- perc.ramp(5)
leg.steps <- c(1, 2, 2.75, 3)

# Colours for overlap categories
cols.discr <- c("#0072B2", "#009E73", "#E69F00")

# Continuous colour scales
cols.c <- viridis(99)
cols.grey <- grey.colors(99, end = 0.99)

# Continuous - logged
log_data <- log10(seq(0.01, 1, length.out = 100))
normalized_log_data <- (log_data - min(log_data)) / (max(log_data) - min(log_data))
cols.c.log <- viridis(99)[as.numeric(cut(normalized_log_data, breaks = 99))]
cols.grey.rev.log <- rev(grey.colors(99, end = 0.99))[as.numeric(cut(normalized_log_data, breaks = 99))]

# Continuous - power transformed
power_data <- seq(0, 1, length.out = 100)^0.5
normalized_power_data <- (power_data - min(power_data)) / (max(power_data) - min(power_data))
cols.c.power <- viridis(99)[as.numeric(cut(normalized_power_data, breaks = 99))]
cols.grey.rev.power <- rev(grey.colors(99, end = 0.99))[as.numeric(cut(normalized_power_data, breaks = 99))]

# Coast/land colours
coast.col <- "grey40"
land.col  <- "grey60"

# Violin plot colours
col.in  <- cols.discr[1]
col.out <- cols[1]

# Plotting ranges
range.pa <- c(0, 1)
range.ab <- c(0, 100)

# Common plotting parameters
plot.par <- par(mfrow = c(1, 1), mar = c(0, 0, 0, 0), oma = c(0, 0, 0, 0))


############################################################
# CREATE HOTSPOT POLYGONS
############################################################

reclass_matrix <- matrix(c(0, NA, 1, NA, 2, 1, 3, 1, 4, 1), ncol = 2, byrow = TRUE)

top.biodiv.reclass    <- classify(top.percentiles$biodiversity, reclass_matrix)
top.biodiv.polygons   <- as.polygons(top.biodiv.reclass, dissolve = TRUE)
top.richness.polygons <- as.polygons(classify(top.percentiles$richness, reclass_matrix), dissolve = TRUE)
top.abundance.polygons <- as.polygons(classify(top.percentiles$abundance, reclass_matrix), dissolve = TRUE)

top.dom.rich.polygons <- as.polygons(classify(top.percentiles.domains[[1]], reclass_matrix), dissolve = TRUE)
for (i in 2:6) {
  top.dom.rich.polygons <- c(
    top.dom.rich.polygons,
    as.polygons(classify(top.percentiles.domains[[i]], reclass_matrix), dissolve = TRUE)
  )
}

top.dom.abund.polygons <- as.polygons(classify(top.percentiles.domains[[7]], reclass_matrix), dissolve = TRUE)
for (i in 8:12) {
  top.dom.abund.polygons <- c(
    top.dom.abund.polygons,
    as.polygons(classify(top.percentiles.domains[[i]], reclass_matrix), dissolve = TRUE)
  )
}

top.dom.biodiv.polygons <- as.polygons(classify(top.percentiles.domains[[13]], reclass_matrix), dissolve = TRUE)
for (i in 14:18) {
  top.dom.biodiv.polygons <- c(
    top.dom.biodiv.polygons,
    as.polygons(classify(top.percentiles.domains[[i]], reclass_matrix), dissolve = TRUE)
  )
}


############################################################
# INSET MAP SETUP
############################################################

# Zoom extents
AP.ext <- ext(-2750000, -2250000, 1050000, 1900000)
PB.ext <- ext(2100000, 2600000, 300000, 1500000)
RS.ext <- ext(-450000, 500000, -1900000, -1250000)
MG.ext <- ext(1300000, 1800000, -2300000, -1900000)

# Shift positions for inset plotting
AP.dx <-  2700000
PB.dx <- -1100000
RS.dx <-  -275000
MG.dx <-  -300000

AP.dy <- -400000
PB.dy <- -500000
RS.dy <- 1100000
MG.dy <-  850000

# Legend positions
l.x  <- -2800000
l.y  <- -1750000
l.xt <- l.x + 480000
l.yt <- l.y + 50000

l.x2 <- -1600000
l.y2 <- -1950000
plg.c <- list(size = c(0.3, 1), ext = c(-5000000, 1800000, -2200000, -2100000), loc = "bottom")

# Title position
loc.main <- c(0, 2800000)

# Coastline insets
AP.coast <- rescale(crop(coast.proj, AP.ext), fx = 2, x0 = mean(AP.ext[1:2]), y0 = mean(AP.ext[3:4]))
PB.coast <- rescale(crop(coast.proj, PB.ext), fx = 2, x0 = mean(PB.ext[1:2]), y0 = mean(PB.ext[3:4]))
RS.coast <- rescale(crop(coast.proj, RS.ext), fx = 2, x0 = mean(RS.ext[1:2]), y0 = mean(RS.ext[3:4]))
MG.coast <- rescale(crop(coast.proj, MG.ext), fx = 2, x0 = mean(MG.ext[1:2]), y0 = mean(MG.ext[3:4]))

AP.coast.b <- shift(AP.coast, AP.dx, AP.dy)
PB.coast.b <- shift(PB.coast, PB.dx, PB.dy)
RS.coast.b <- shift(RS.coast, RS.dx, RS.dy)
MG.coast.b <- shift(MG.coast, MG.dx, MG.dy)

# Sample points for inset maps
sp.samples <- vect(img.metadata[, 5:6], geom = c("proj_coord_x", "proj_coord_y"), crs = crs(hmsc.maps.richness.env$median))

AP.samples <- rescale(crop(sp.samples, AP.ext), fx = 2)
PB.samples <- rescale(crop(sp.samples, PB.ext), fx = 2)
RS.samples <- rescale(crop(sp.samples, RS.ext), fx = 2)
MG.samples <- rescale(crop(sp.samples, MG.ext), fx = 2)

AP.b.samples <- shift(AP.samples, AP.dx, AP.dy)
PB.b.samples <- shift(PB.samples, PB.dx, PB.dy)
RS.b.samples <- shift(RS.samples, RS.dx, RS.dy)
MG.b.samples <- shift(MG.samples, MG.dx, MG.dy)
``
```

















































```{r ReadAndPrepData_EnvGaps}
##########################
#### Environmental gaps
## load environmental gaps
gaps.hv.a <- rast(paste0(DP.dir,"Circumpolar_Analysis_GapHypervolume_2km_All21SurveysCombined_AllVariablesExceptDepth2Canyons_Quantile095.tif"))
levels(gaps.hv.a) <- data.frame(id=0:1, cover=c("outside of environmental hypervolume","inside of environmental hypervolume"))
```

```{r}
##########################
#### SO-Predators
## load SO-predator data
predator.raw <- rast("C:/Users/jjansen/OneDrive - University of Tasmania/science/data_biological/Hi2023-aes_colony_weighted_cog.tif")

#### upper decile for all of the Southern Ocean
predator.top10 <- topX.calc(predator.raw, x=0.1)
## project upper decile to polar stereographic
predator.top10.proj <- project(predator.top10, top.percentiles$biodiversity, method="near")
## calculate polygons doesn't work, maybe because we project after the calculations
#predator.top10.polygons <- as.polygons(predator.top10.proj, dissolve=TRUE)

#### upper decile for shelf only
## project raw data to polar stereographic
predator.proj <- project(predator.raw, top.percentiles$biodiversity)
## reduce to shelf only
predator.shelf <- predator.proj
predator.shelf[is.na(top.percentiles$biodiversity[])] <- NA
## calculate upper decile like in the Hindell paper
predator.shelf.top25 <- topX.calc(predator.shelf, x=0.25)
predator.shelf.top10 <- topX.calc(predator.shelf, x=0.1)
predator.shelf.top5  <- topX.calc(predator.shelf, x=0.05)
predator.shelf.top1  <- topX.calc(predator.shelf, x=0.01)
predator.shelf.top   <- sum(predator.shelf.top1, predator.shelf.top5, predator.shelf.top10, predator.shelf.top25)
## calculate polygons
reclass_matrix <- matrix(c(0, NA, 1, NA, 2, 1, 3, 1, 4, 1), ncol=2, byrow=TRUE)
predator.shelf.top.reclass <- classify(predator.shelf.top, reclass_matrix)
predator.shelf.top.polygons <- as.polygons(predator.shelf.top.reclass, dissolve=TRUE)
## overlapping areas
top.combined.only <- app(c(top.biodiv.reclass, predator.shelf.top.reclass), fun=sum, na.rm=FALSE)
top.combined.only.polygons <- as.polygons(top.combined.only, dissolve=TRUE)
## combine to a single raster with 3 levels
top.combined <- app(c(top.biodiv.reclass, predator.shelf.top.reclass, predator.shelf.top.reclass), fun=sum, na.rm=TRUE)
```

# Displaying predictive maps {.tabset}

## Paper figures {.tabset}

### Fig 1

#### Biodiversity Hotspots

```{r Fig1}
output.basename <- "Circumpolar_biodiversity_percentiles"

ra <- top.percentiles$biodiversity
levels(ra)[[1]]$biodiversity[2:5] <- paste0("   ",levels(ra)[[1]]$biodiversity[2:5])

## extract, zoom and shift inset maps
crop_and_shift(ra)
#### plot everything
if(pngs){png(paste0(output.basename, ".png"), width=png.width, height=png.height, res=png.res)
  }else if(pdfs){pdf(paste0(output.basename, ".pdf"), width=pdf.width, height=pdf.height)}
par(plot.par)
plot(ra, main="Biodiversity Hotspots", col=cols, axes=FALSE, plg=list(x=l.x, y=l.y-80000), loc.main=loc.main)
text(x=l.xt, y=l.yt-80000, "percentiles")
add.plotsandlines()
if(pngs|pdfs){dev.off()}
```

#### Alternatives

```{r}
#### hotspots on bathymetry
ra <- r2
range=range(values(ra), na.rm=TRUE)

## extract, zoom and shift inset maps
crop_and_shift(ra, top.biodiv.polygons)
```

```{r}
if(pngs){png(paste0(pred.dir,"Circumpolar_biodiversity_percentiles_AES.png"), width=png.width, height=png.height, res=png.res)}
par(plot.par)
plot(ra, main="Biodiversity Hotspots", col=cols.grey, axes=FALSE, plg=plg.c, loc.main=loc.main)
plot(top.biodiv.polygons, add=TRUE, col=cols.discr[1], border=NA)#, alpha=0.5)
text(x=l.x2, y=l.y2, "Depth (m)")
add.plotsandlines(inset.cols=cols.grey, plot.points=TRUE, range=range, polygons=TRUE, fill.col=cols.discr[1], border.col=NA)#, alpha=0.5)
if(pngs){dev.off()}
```

```{r}
#if(pngs){png(paste0(pred.dir,"Circumpolar_biodiversity_percentiles_AES.png"), width=png.width, height=png.height, res=png.res)}
par(plot.par)
plot(ra, main="Biodiversity Hotspots", col=cols.grey, axes=FALSE, plg=plg.c, loc.main=loc.main)
plot(top.biodiv.polygons, add=TRUE, col=cols.discr[1], border=NA)#, alpha=0.5)
text(x=l.x2, y=l.y2, "Depth (m)")
# add.plotsandlines(inset.cols=cols.grey, plot.points=FALSE, range=range, polygons=TRUE, fill.col="steelblue1", border.col=NA)#, alpha=0.5)
plot(coast.proj[values(coast.proj)=="land"], col=land.col, add=TRUE)
lines(coast.proj, col=coast.col)
#if(pngs){dev.off()}
```

### Fig 2

#### Richness

```{r}
ra <- hmsc.maps.richness.env$median
range=c(0,max(values(ra), na.rm=TRUE))

## extract, zoom and shift inset maps
crop_and_shift(ra)

#### plot everything
if(pngs){png(paste0(pred.dir,"Circumpolar_Richness_median.png"), width=png.width, height=png.height, res=png.res)}
par(plot.par)
plot(ra, main="Richness - Median", col=cols.c.power, axes=FALSE, plg=plg.c, loc.main=loc.main)
text(x=l.x2, y=l.y2, "number of morphospecies")
# points(sp.samples, col="black", pch=19, cex=0.8)
add.plotsandlines(inset.cols=cols.c.power, plot.points=FALSE, range=range)
if(pngs){dev.off()}
```

```{r}
ra <- hmsc.maps.richness.env$se
range=c(0,max(values(ra), na.rm=TRUE))

## extract, zoom and shift inset maps
crop_and_shift(ra)

#### plot everything
if(pngs){png(paste0(pred.dir,"Circumpolar_Richness_se.png"), width=png.width, height=png.height, res=png.res)}
par(plot.par)
plot(ra, main="Richness - Standard Error", col=rev(cols.grey), axes=FALSE, plg=plg.c, loc.main=loc.main)
text(x=l.x2, y=l.y2, "number of morphospecies")
# points(sp.samples, col="black", pch=19, cex=0.8)
add.plotsandlines(inset.cols=rev(cols.grey), plot.points=FALSE, range=range)
if(pngs){dev.off()}
```

```{r}
ra <- boots.rich.median
range=c(0,max(values(ra), na.rm=TRUE))

## extract, zoom and shift inset maps
crop_and_shift(ra)

#### plot everything
if(pngs){png(paste0(pred.dir,"Circumpolar_Richness_bootstrapped_median.png"), width=png.width, height=png.height, res=png.res)}
par(plot.par)
plot(ra, main="Richness - Bootstrapped Median", col=cols.c.power, axes=FALSE, plg=plg.c, loc.main=loc.main)
text(x=l.x2, y=l.y2, "number of morphospecies")
# points(sp.samples, col="black", pch=19, cex=0.8)
add.plotsandlines(inset.cols=cols.c.power, plot.points=FALSE, range=range)
if(pngs){dev.off()}
```

```{r}
ra <- boots.rich.sd
range=c(0,max(values(ra), na.rm=TRUE))

## extract, zoom and shift inset maps
crop_and_shift(ra)

#### plot everything
if(pngs){png(paste0(pred.dir,"Circumpolar_Richness_Bootstrapped_sd.png"), width=png.width, height=png.height, res=png.res)}
par(plot.par)
plot(ra, main="Richness - Bootstrapped Standard Error", col=rev(cols.grey), axes=FALSE, plg=plg.c, loc.main=loc.main)
text(x=l.x2, y=l.y2, "number of morphospecies")
# points(sp.samples, col="black", pch=19, cex=0.8)
add.plotsandlines(inset.cols=rev(cols.grey), plot.points=FALSE, range=range)
if(pngs){dev.off()}
```

### Fig 3

#### Total Abundance

```{r}
ra <- hmsc.maps.totalabundance.env$median.100
## extract, zoom and shift inset maps
crop_and_shift(ra)

#### plot everything
if(pngs){png(paste0(pred.dir,"Circumpolar_Totalabundance_median.png"), width=png.width, height=png.height, res=png.res)}
par(plot.par)
plot(ra, main="Abundance - Median", col=cols.c.power, axes=FALSE, plg=plg.c, range=range.ab, loc.main=loc.main)
text(x=l.x2, y=l.y2, "%-cover")
# points(sp.samples, col="black", pch=19, cex=0.8)
add.plotsandlines(inset.cols=cols.c.power, plot.points=FALSE, range=range.ab)
if(pngs){dev.off()}
```

```{r}
ra <- hmsc.maps.totalabundance.env$se
ra[which(ra[]>100)] <- NA

## extract, zoom and shift inset maps
crop_and_shift(ra)

#### plot everything
if(pngs){png(paste0(pred.dir,"Circumpolar_Totalabundance_se.png"), width=png.width, height=png.height, res=png.res)}
par(plot.par)
plot(ra, main="Abundance - Standard Error", col=cols.grey.rev.power, axes=FALSE, plg=plg.c, range=range.ab, loc.main=loc.main)
text(x=l.x2, y=l.y2, "%-cover")
# points(sp.samples, col="black", pch=19, cex=0.8)
add.plotsandlines(inset.cols=cols.grey.rev.power, plot.points=FALSE, range=range.ab)
if(pngs){dev.off()}
```

### Fig 4 {.tabset}

environmental conditions inside upper decile

```{r}
viol.names <- "    Hotspots   Other areas"
# viol.names <- " Biodiv.-AES   Other areas"

## load environmental values for biodiversity hotspots vs all other areas
load("C:/Users/jjansen/OneDrive - University of Tasmania/science/data_biological/BiodiversityHotspots_EnvironmentalSetting.Rdata")
```

#### main differences: top 10% biodiversity hotspots vs all other areas

```{r, fig.height=4, fig.width=8}
sel <- c(1, 6, 9, 10)
y_labels <- c(
 "Depth (m)",
 "Mean seafloor current speed (m/s)",
 "Seafloor temperature (\u00B0C)",
 "log(seafloor particle flux)"
)

par(mfrow=c(1,4), mgp=c(1, 1, 0))
for(i in 1:4){
  ## Plot violin plots
  vioplot(vals.in[[sel[i]]],  col=col.in, side="left", names=viol.names)
  vioplot(vals.out[[sel[i]]], col=col.out, side="right", add=TRUE)#, h=bw
  title(ylab=y_labels[i], line=3)
}

#### adjustments
## smoothness can be adjusted with e.g. "h=20"
# vioplot(vals.in.depth,  col=col.in, side="left", h=20)
# vioplot(vals.out.depth, col=col.out, side="right", add=TRUE, h=20)
## width can be made proportional with "wex"
# count_ratio <- length(vals.in.depth) / length(vals.out.depth)
# vioplot(vals.in.depth,  col=col.in, side="left", wex=count_ratio)
# vioplot(vals.out.depth, col=col.out, side="right", add=TRUE, wex=1)
```

```{r, fig.height=4, fig.width=8}
count_ratio <- length(vals.in[[1]]) / length(vals.out[[1]])

par(mfrow=c(1,4), mgp=c(1, 1, 0))
for(i in 1:4){
  ## Plot violin plots
  vioplot(vals.in[[sel[i]]],  col=col.in, side="left", names=viol.names, wex=count_ratio)
  vioplot(vals.out[[sel[i]]], col=col.out, side="right", add=TRUE, wex=1)#, h=bw
  title(ylab=y_labels[i], line=3)
}
```

Histograms work for some but not easy to interpret for others

```{r, fig.height=8, fig.width=8}
prep.hist <- function(x.in, x.out,breaks="Sturges"){
  hist_in  <- hist(x.in, plot = FALSE, breaks=breaks)
  hist_out <- hist(x.out, plot = FALSE, breaks=breaks)
  ## Extract midpoints, assuming both histograms have the same breaks
  midpoints  <<- hist_in$mids
  ## Combine the counts into a matrix for side-by-side plotting
  counts_matrix <<- rbind(hist_in$counts, hist_out$counts)
}
par(mfrow=c(2,2))
prep.hist(vals.in[[sel[1]]], vals.out[[sel[1]]])
barplot(counts_matrix, beside = TRUE, horiz = TRUE, col = c(col.in, col.out), 
        names.arg = midpoints, xlab = "Frequency", ylab = "Depth", 
        legend.text = c("In", "Out"))

prep.hist(vals.in[[sel[3]]], vals.out[[sel[3]]])
barplot(counts_matrix, beside = TRUE, horiz = TRUE, col = c(col.in, col.out), 
        names.arg = midpoints, xlab = "Frequency", ylab = "Temperature", 
        legend.text = c("In", "Out"))

prep.hist(vals.in[[sel[2]]], vals.out[[sel[2]]], breaks <- c(seq(0, 0.2, by= 0.02), 0.65))#seq(0.25, 0.65, by= 0.05)))
# Create a horizontal barplot
barplot(counts_matrix, beside = TRUE, horiz = TRUE, col = c(col.in, col.out), 
        names.arg = midpoints, xlab = "Frequency", ylab = "Current speed")

prep.hist(vals.in[[sel[4]]], vals.out[[sel[4]]], breaks <- seq(0, 15, by= 1))
# Create a horizontal barplot
barplot(counts_matrix, beside = TRUE, horiz = TRUE, col = c(col.in, col.out), 
        names.arg = midpoints, xlab = "Frequency", ylab = "log(flux)")
```

What proportion of the shelf falls into the bottom half temperature?

```{r}
## extract values
vals_everywhere <- c(vals.in$seafloortemperature, vals.out$seafloortemperature)
vals_inside     <- vals.in$seafloortemperature
vals_outside    <- vals.out$seafloortemperature
```

```{r}
##
paste0("median temperature: ", round(median(vals_everywhere), 3))
##
paste0("Upper decile hotspots in colder half: ", round(length(which(vals_inside<median(vals_everywhere)))/length(vals_inside),3))
paste0("Upper decile hotspots in warmer half: ", round(length(which(vals_inside>median(vals_everywhere)))/length(vals_inside),3))
```

```{r}
paste0("prop. of shelf below 0 degrees: ", round(length(which(vals_everywhere<(0)))/length(vals_everywhere), 3))
##
paste0("Upper decile hotspots in colder part: ", round(length(which(vals_inside<(0)))/length(vals_inside),3))
paste0("Upper decile hotspots in warmer part: ", round(length(which(vals_inside>(0)))/length(vals_inside),3))
```

```{r}
paste0("prop. of shelf below -1.55 degrees: ", round(length(which(vals_everywhere<(-1.55)))/length(vals_everywhere), 3))
##
paste0("Upper decile hotspots in colder part: ", round(length(which(vals_inside<(-1.55)))/length(vals_inside),3))
paste0("Upper decile hotspots in warmer part: ", round(length(which(vals_inside>(-1.55)))/length(vals_inside),3))
```

#### all other comparisons
```{r, fig.height=4, fig.width=8}
sel <- c(2:5,7:8,11)
y_labels <- c(
  "Slope (\u00B0)",
  "Topographic position index",
  "Distance to canyons (m)",
  "Net primary productivity (mg C/m\u2/day)",
  "Residual seafloor current speed (m/s)",
  "Seafloor salinity",
  "Seafloor particle sedimentation"
)

par(mfrow=c(1,4), mgp=c(1, 1, 0))
for(i in 1:7){
  ## Plot violin plots
  vioplot(vals.in[[sel[i]]],  col=col.in, side="left", names=viol.names)
  vioplot(vals.out[[sel[i]]], col=col.out, side="right", add=TRUE)#, h=bw
  title(ylab=y_labels[i], line=3)
}
```


#### bathymetry

```{r, fig.height=4, fig.width=8}
par(mfrow=c(1,4))
for(i in 1:4){
  vioplot(vals.in[[i]], col=col.in, side="left")#, names="", ylab="depth")
  vioplot(vals.out[[i]], col=col.out, side="right", add=TRUE)
}
```

#### water

```{r, fig.height=4, fig.width=8}
par(mfrow=c(1,4))
for(i in 6:9){
  vioplot(vals.in[[i]], col=col.in, side="left")
  vioplot(vals.out[[i]], col=col.out, side="right", add=TRUE)
}
```

```{r}
plot(density(vals.out[[6]]), col=col.out)
lines(density(vals.in[[6]]), col=col.in)

plot(density(vals.in[[8]]), col=col.in, xlim=c(33.9,35.1))
lines(density(vals.out[[8]]), col=col.out)
```

#### food

```{r, fig.height=4, fig.width=8}
par(mfrow=c(1,4))
for(i in c(5,10,11)){
  vioplot(vals.in[[i]], col=col.in, side="left")
  vioplot(vals.out[[i]], col=col.out, side="right", add=TRUE)
}
```


#### hotspot stats

```{r}
rich.vals.in  <- mask(hmsc.maps.richness.env$median, top.biodiv.polygons)[]
rich.vals.out <- mask(hmsc.maps.richness.env$median, top.biodiv.polygons, inverse=TRUE)[]
abund.vals.in  <- mask(hmsc.maps.totalabundance.env$median.100, top.biodiv.polygons)[]
abund.vals.out <- mask(hmsc.maps.totalabundance.env$median.100, top.biodiv.polygons, inverse=TRUE)[]
message("hotspot richness mean & sd =")
mean(rich.vals.in, na.rm=TRUE)
sd(rich.vals.in, na.rm=TRUE)
message("non-hotspot richness mean & sd =")
mean(rich.vals.out, na.rm=TRUE)
sd(rich.vals.out, na.rm=TRUE)
message("hotspot abundance mean & sd =")
mean(abund.vals.in, na.rm=TRUE)
sd(abund.vals.in, na.rm=TRUE)
message("non-hotspot abundance mean & sd =")
mean(abund.vals.out, na.rm=TRUE)
sd(abund.vals.out, na.rm=TRUE)
```

```{r}
#### 98321 cells at 4km^2 -> 393284km^2 covered at ~26% -> 102000km^2 at 100%
sel <- which(!is.na(abund.vals.in))
cell_area <- 4  # km²
n.cells <- length(sel)
hotspot.area <-  n.cells * cell_area
div.val <- mean(abund.vals.in[sel])
hotspot.area/(100/div.val)

#### uncertainty
set.seed(42)
## Parameters
n.iterations <- 1000
## Bootstrap loop
bootstrap.estimates <- numeric(n.iterations)
for (i in 1:n.iterations) {
  sample.vals <- sample(abund.vals.in[sel], size = n.cells, replace = TRUE)
  mean.coverage <- mean(sample.vals, na.rm = TRUE)
  bootstrap.estimates[i] <- hotspot.area / (100 / mean.coverage)
}

# Results
mean.estimate <- mean(bootstrap.estimates)
ci <- quantile(bootstrap.estimates, probs = c(0.025, 0.975))

cat("Estimated fully covered area:", round(mean.estimate, 2), "km²\n")
cat("95% Confidence Interval:", round(ci[1], 2), "-", round(ci[2], 2), "km²\n")
```

```{r}
#### 98321 cells at 4km^2 -> 393284km^2 covered at ~26% -> 102000km^2 at 100%
sel <- which(!is.na(abund.vals.out))
cell_area <- 4  # km²
n.cells <- length(sel)
hotspot.area <-  n.cells * cell_area
div.val <- mean(abund.vals.out, na.rm=TRUE)
hotspot.area/(100/div.val)

#### uncertainty
set.seed(42)
## Parameters
n.iterations <- 1000
## Bootstrap loop
bootstrap.estimates <- numeric(n.iterations)
for (i in 1:n.iterations) {
  sample.vals <- sample(abund.vals.out[sel], size = n.cells, replace = TRUE)
  mean.coverage <- mean(sample.vals, na.rm = TRUE)
  bootstrap.estimates[i] <- hotspot.area / (100 / mean.coverage)
}

# Results
mean.estimate <- mean(bootstrap.estimates)
ci <- quantile(bootstrap.estimates, probs = c(0.025, 0.975))

cat("Estimated fully covered area:", round(mean.estimate, 2), "km²\n")
cat("95% Confidence Interval:", round(ci[1], 2), "-", round(ci[2], 2), "km²\n")
```

### Fig 6

#### Overlap with Predator Hotspots

```{r}
if(pngs){png(paste0(pred.dir,"Circumpolar_biodiversity_predator_AES.png"), width=png.width, height=png.height, res=png.res)}
plot(r2, main="Overlap in key habitats", col=cols.grey, axes=FALSE, plg=plg.c, loc.main=loc.main)
plot(top.combined, add=TRUE, col=cols.discr, legend=FALSE)
legend("topright", legend=c("Seafloor", "Predators", "Both"), fill=cols.discr, bty="n")
crop_and_shift(r2, top.biodiv.polygons)
add.plotsandlines(inset.cols=cols.grey, plot.points=FALSE, polygons=TRUE, fill.col=cols.discr[1], border.col=NA)
crop_and_shift(r2, predator.shelf.top.polygons)
add.polygons(fill.col=cols.discr[2], border.col=NA)
crop_and_shift(r2, top.combined.only.polygons)
add.polygons(fill.col=cols.discr[3], border.col=NA)
text(x=l.x2, y=l.y2, "Depth (m)")
if(pngs){dev.off()}
```

#### Quantifying the overlap

```{r}
ncell.biodiv  <- length(which(top.combined[]==1))
ncell.preds   <- length(which(top.combined[]==2))
ncell.overlap <- length(which(top.combined[]==3))
ncell.total   <- length(which(top.combined[]>=1))

paste0(round(ncell.overlap/(ncell.biodiv+ncell.overlap),3)," = prop. of biodiversity hotspots that are also predator hotspots")
paste0(round(ncell.overlap/(ncell.preds+ncell.overlap),3)," = prop. of predator hotspots that are also biodiversity hotspots")
paste0(round(ncell.overlap/ncell.total,3)," = prop. of overlap across all hotspots areas")
```

##### Comparing differences: overlapping biodiversity-predator-hotspots vs all biodiversity hotspots

```{r}
## load environmental values for biodiversity hotspots vs all other areas
load("C:/Users/jjansen/OneDrive - University of Tasmania/science/data_biological/BiodiversityHotspots_EnvironmentalSetting.Rdata")

## load environmental values for predator hotspots and overlapping areas with benthic hotspots
load("C:/Users/jjansen/OneDrive - University of Tasmania/science/data_biological/BiodiversityHotspots_EnvironmentalSetting_PredatorsAndOverlap.Rdata")

viol.names2 <- " Biodiv.-AES   Pred-Overlap"
```

```{r, fig.height=4, fig.width=8}
par(mfrow=c(1,4), mgp=c(1, 1, 0))
for(i in 1:nlyr(r.stack.subset)){
  ## Plot violin plots
vioplot(vals.in[[i]], col=col.in, side="left", names=viol.names2)
vioplot(vals.o.in[[i]],  col=col.out, side="right", add=TRUE)
title(ylab=names(r.stack.subset)[i], line=3)
}
```

```{r, fig.height=4, fig.width=8}
par(mfrow=c(1,4), mgp=c(1, 1, 0))
for(i in 1:4){
  ## Plot violin plots
vioplot(vals.in[[sel[i]]], col=col.in, side="left", names=viol.names2)
vioplot(vals.o.in[[sel[i]]],  col=col.out, side="right", add=TRUE)
title(ylab=y_labels[i], line=3)
}
```

##### Comparing differences: biodiversity hotspots vs predator hotspots

```{r}
viol.names3 <- "Biodiv.-AES   Pred.-AES  "
```

```{r, fig.height=4, fig.width=8}
par(mfrow=c(1,4), mgp=c(1, 1, 0))
for(i in 1:nlyr(r.stack.subset)){
  ## Plot violin plots
vioplot(vals.in[[i]], col=col.in, side="left", names=viol.names3)
vioplot(vals.p.in[[i]],  col=col.out, side="right", add=TRUE)
title(ylab=names(r.stack.subset)[i], line=3)
}
```

#### Alternatives

```{r}
if(pngs){png(paste0(pred.dir,"Circumpolar_biodiversity_predator_overlappingAES.png"), width=png.width, height=png.height, res=png.res)}
plot(r2, main="Overlap in key habitats", col=cols.grey, axes=FALSE, plg=plg.c, loc.main=loc.main)
plot(top.biodiv.polygons, add=TRUE, col=cols.discr[1], border=NA, legend=FALSE)
plot(top.combined.only, add=TRUE, col=cols.discr[3], legend=FALSE)
legend("topright", legend=c("Seafloor hotspots", "Overlap with Predator hotspots"), fill=cols.discr[c(1,3)], bty="n")
crop_and_shift(r2, top.biodiv.polygons)
add.plotsandlines(inset.cols=cols.grey, plot.points=FALSE, polygons=TRUE, fill.col=cols.discr[1], border.col=NA)
crop_and_shift(r2, top.combined.only.polygons)
add.polygons(fill.col=cols.discr[3], border.col=NA)
text(x=l.x2, y=l.y2, "Depth (m)")
if(pngs){dev.off()}
```

```{r, echo=FALSE}
## extract, zoom and shift inset maps
crop_and_shift(r2, predator.shelf.top.polygons)
range=range(values(r2), na.rm=TRUE)

#### plot everything
if(pngs){png(paste0(pred.dir,"Circumpolar_predator_AES.png"), width=png.width, height=png.height, res=png.res)}
par(plot.par)
plot(r2, main="Top 10% of Antarctic shelf habitat importance for mammal/seabird predators\nfrom Hindell et al (2020)", col=cols.grey, axes=FALSE, plg=plg.c, loc.main=loc.main)
# plot(predator.shelf.top, col=cols, axes=FALSE, plg=list(x=l.x, y=l.y), loc.main=loc.main)
plot(predator.shelf.top.polygons, add=TRUE, col=cols.discr[2], border=NA)
text(x=l.x2, y=l.y2, "Depth (m)")
add.plotsandlines(inset.cols=cols.grey, plot.points=FALSE, range=range, polygons=TRUE, fill.col=cols.discr[2], border.col=NA)#, alpha=0.5)
if(pngs){dev.off()}
```

```{r}
plot(predator.top10.proj, col=c("white",cols.discr[2]))
lines(coast.proj, add=TRUE, col="grey40")
#plot(predator.top10.polygons, col="red")
```

```{r, eval=FALSE, echo=FALSE}
ra <- predator.shelf
## extract, zoom and shift inset maps
crop_and_shift(ra, predator.shelf.top.polygons)
#### plot everything
par(plot.par)
plot(ra, main="Predator habitat importance", col=cols.c.power, axes=FALSE, plg=plg.c, loc.main=loc.main)
plot(predator.shelf.top.polygons, add=TRUE, border="white")
text(x=l.x2, y=l.y2, "habitat importance")
add.plotsandlines(inset.cols=cols.c.power, plot.points=FALSE, polygons=TRUE, fill.col=NA, border.col="white")
```

#### overlap in individual regions

```{r}
#### overlap in AP
AP.crop <- crop(top.combined, AP.ext)
# length(which(AP.crop[]==1)) #  9325
# length(which(AP.crop[]==2)) # 10709
# length(which(AP.crop[]==3)) #  5910
print("% of biodiversity hotspots (AP) that overlap with predator hotspots")
length(which(AP.crop[]==3))/length(which(AP.crop[]!=2))

#### overlap in PB
PB.crop <- crop(top.combined, PB.ext)
# length(which(PB.crop[]==1)) #  4936
# length(which(PB.crop[]==2)) # 24517
# length(which(PB.crop[]==3)) # 12068
print("% of biodiversity hotspots (PB) that overlap with predator hotspots")
length(which(PB.crop[]==3))/length(which(PB.crop[]!=2))
```

## Suppl Figures {.tabset}

### Suppl Fig - Hotspots by depth bins

```{r}
# Load libraries
library(terra)
library(dplyr)
library(ggplot2)
library(tidyr)
library(scales)

# 1. Extract values and filter NAs
depth_vals <- values(r2)
hotspot_vals <- values(top.combined)
valid_idx <- which(!is.na(depth_vals))# & !is.na(hotspot_vals))
depth_vals <- depth_vals[valid_idx]
hotspot_vals <- hotspot_vals[valid_idx]

# 2. Define depth bins
depth_bins <- seq(-2500, 0, by = 100)
depth_labels <- paste(head(depth_bins, -1), "to", tail(depth_bins, -1), "m")
depth_bin_factor <- cut(depth_vals, breaks = depth_bins, labels = depth_labels, include.lowest = TRUE)

# 3. Create data frame
df <- data.frame(
  depth = depth_vals,
  depth_bin = depth_bin_factor,
  hotspot = hotspot_vals
)

# 4. Summarize counts
summary_df <- df %>%
  group_by(depth_bin) %>%
  summarise(
    total_cells = n(),
    seafloor_cells = sum(hotspot %in% c(1, 3)),
    overlap_cells = sum(hotspot %in% 3),
    prop_seafloor = seafloor_cells / total_cells,
    prop_overlap = ifelse(seafloor_cells > 0, overlap_cells / seafloor_cells, NA)
  )

# Scale proportions to match the primary y-axis (cell counts)
max_count <- max(summary_df$total_cells, na.rm = TRUE)
summary_df <- summary_df %>%
  mutate(
    prop_seafloor_scaled = prop_seafloor * max_count,
    prop_overlap_scaled = prop_overlap * max_count
  )

plot_data <- summary_df %>%
  select(depth_bin, total_cells, seafloor_cells, overlap_cells) %>%
  pivot_longer(cols = c(total_cells, seafloor_cells, overlap_cells),
               names_to = "category", values_to = "value") %>%
  mutate(category = recode(category,
    "total_cells" = "Total number of cells",
    "seafloor_cells" = "Cells with seafloor biodiversity hotspots",
    "overlap_cells" = "Cells with both seafloor biodiversity & predator hotspots"
  )) %>%
  mutate(category = factor(category, levels = c(
    "Total number of cells",
    "Cells with seafloor biodiversity hotspots",
    "Cells with both seafloor biodiversity & predator hotspots"
  )))
```

##### seafloor hotspots vs number of cells - raw counts

```{r}
# 5. Plot: Counts of total vs seafloor hotspots
ggplot(summary_df, aes(x = depth_bin)) +
  geom_col(aes(y = total_cells), fill = "lightgray", width = 0.8) +
  geom_col(aes(y = seafloor_cells), fill = cols.discr[1], width = 0.4) +
  labs(
    x = "Depth Range (m)",
    y = "Number of Cells",
    title = "Seafloor Hotspot Cells vs Total Cells by Depth Bin"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

##### seafloor hotspots vs overlap vs number of cells - counts and proportions

```{r}
if(pngs){png(paste0(pred.dir,"Circumpolar_biodiversity_depthstats.png"), width=png.width, height=png.height, res=png.res)}
# Plot
ggplot() +
  # Grey bars (Total number of cells - widest)
  geom_col(data = filter(plot_data, category == "Total number of cells"), 
           aes(x = depth_bin, y = value, fill = category), width = 0.8) +
  # Blue bars (Seafloor biodiversity hotspots - medium)
  geom_col(data = filter(plot_data, category == "Cells with seafloor biodiversity hotspots"),
           aes(x = depth_bin, y = value, fill = category), width = 0.6) +
  # Yellow bars (Overlap with predator hotspots - narrowest)
  geom_col(data = filter(plot_data, category == "Cells with both seafloor biodiversity & predator hotspots"),
           aes(x = depth_bin, y = value, fill = category), width = 0.4) +
  # Proportion lines and points with color aesthetic mapped to labels
  geom_line(data = summary_df, aes(x = depth_bin, y = prop_seafloor_scaled, group = 1,
                                   color = "% of all cells that are seafloor biodiversity hotspots"), size = 1.2) +
  geom_point(data = summary_df, aes(x = depth_bin, y = prop_seafloor_scaled,
                                    color = "% of all cells that are seafloor biodiversity hotspots"), size = 2) +
  geom_line(data = summary_df, aes(x = depth_bin, y = prop_overlap_scaled, group = 1,
                                   color = "% of seafloor biodiversity hotspots that are predator hotspots"), size = 1.2) +
  geom_point(data = summary_df, aes(x = depth_bin, y = prop_overlap_scaled,
                                    color = "% of seafloor biodiversity hotspots that are predator hotspots"), size = 2) +
  # Y-axis scaling
  scale_y_continuous(
    name = "Number of Cells",
    breaks = seq(0, max_count, by = max_count / 10),
    labels = number_format(accuracy = 1),
    limits = c(0, max_count),
    sec.axis = sec_axis(
      ~ . / max_count,
      name = "Proportion",
      breaks = seq(0, 1, by = 0.1),
      labels = percent_format(accuracy = 1)
    )
  ) +
  # Manual fill colors for bars
  scale_fill_manual(
    values = c(
      "Total number of cells" = "lightgray",
      "Cells with seafloor biodiversity hotspots" = cols.discr[1],
      "Cells with both seafloor biodiversity & predator hotspots" = cols.discr[3]
    )
  ) +
  # Manual color scale for lines with swapped order
  scale_color_manual(
    values = c(
      "% of all cells that are seafloor biodiversity hotspots" = cols.discr[1],
      "% of seafloor biodiversity hotspots that are predator hotspots" = cols.discr[3]
    ),
    breaks = c(
      "% of all cells that are seafloor biodiversity hotspots",
      "% of seafloor biodiversity hotspots that are predator hotspots"
    )
  ) +
  # Guides to stack legends
  guides(
    fill = guide_legend(order = 1),
    color = guide_legend(order = 2)
  ) +
  # Labels and theme
  labs(
    x = "Depth Range (m)",
    title = "Seafloor Hotspot Counts and Proportions by Depth Bin"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = c(0.05, 0.95),
    legend.justification = c("left", "top"),
    legend.title = element_blank()
  )
if(pngs){dev.off()}
detach("package:scales")
```

### Suppl Fig - Richness Hotspots

```{r Fig2}
ra <- top.percentiles$richness
levels(ra)[[1]]$richness[2:5] <- paste0("   ",levels(ra)[[1]]$richness[2:5])

## extract, zoom and shift inset maps
crop_and_shift(ra)

#### plot everything
if(pngs){png(paste0(pred.dir,"Circumpolar_richness_percentiles.png"), width=png.width, height=png.height, res=png.res)}
par(plot.par)
plot(ra, main="Richness hotspots", col=cols, axes=FALSE, plg=list(x=l.x, y=l.y), loc.main=loc.main)
text(x=l.xt, y=l.yt, "percentiles")
add.plotsandlines()
if(pngs){dev.off()}
```

```{r}
ra <- hmsc.maps.richness.env$median
range=c(0,max(values(ra), na.rm=TRUE))

## extract, zoom and shift inset maps
crop_and_shift(ra,top.richness.polygons)

#### plot everything
if(pngs){png(paste0(pred.dir,"Circumpolar_richness_percentiles_AES.png"), width=png.width, height=png.height, res=png.res)}
par(plot.par)
plot(ra, main="Richness median", col=cols.c.power, axes=FALSE, plg=plg.c, loc.main=loc.main)
plot(top.richness.polygons, add=TRUE, border="white")
text(x=l.x2, y=l.y2, "number of species")
add.plotsandlines(inset.cols=cols.c.power, plot.points=FALSE, range=range, polygons=TRUE, fill.col=NA, border.col="white")
if(pngs){dev.off()}
```

```{r}
ra <- hmsc.maps.richness.env$median
range=c(0,max(values(ra), na.rm=TRUE))

## extract, zoom and shift inset maps
crop_and_shift(ra, top.richness.polygons)

#### plot everything
#if(pngs){png(paste0(pred.dir,"Circumpolar_richness_percentiles_AES.png"), width=png.width, height=png.height, res=png.res)}
par(plot.par)
plot(ra, main="Richness median", col=cols.c.power, axes=FALSE, plg=plg.c, loc.main=loc.main)
plot(top.richness.polygons, add=TRUE, border=NA, col="gold")
text(x=l.x2, y=l.y2, "number of species")
add.plotsandlines(inset.cols=cols.c.power, plot.points=FALSE, range=range, polygons=TRUE, border.col=NA, fill.col="gold")
#if(pngs){dev.off()}
```

### Suppl Fig - Abundance Hotspots

```{r Fig3}
ra <- top.percentiles$abundance
levels(ra)[[1]]$abundance[2:5] <- paste0("   ",levels(ra)[[1]]$abundance[2:5])

## extract, zoom and shift inset maps
crop_and_shift(ra)

#### plot everything
if(pngs){png(paste0(pred.dir,"Circumpolar_abundance_percentiles.png"), width=png.width, height=png.height, res=png.res)}
par(plot.par)
plot(ra, main="Abundance hotspots", col=cols, axes=FALSE, plg=list(x=l.x, y=l.y), loc.main=loc.main)
text(x=l.xt, y=l.yt, "percentiles")
add.plotsandlines()
if(pngs){dev.off()}
```

### Suppl Fig - Richness-Only Hotspots

```{r Fig4a}
ra <- top.percentiles$`richness-only`
## extract, zoom and shift inset maps
crop_and_shift(ra)

cols3 <- cols[1:4]#cols[c(1,4,5)]#

#### plot everything
if(pngs){png(paste0(pred.dir,"Circumpolar_richness_only_hotspots.png"), width=png.width, height=png.height, res=png.res)}
par(plot.par)
plot(ra, main="high-richness only hotspots", col=cols3, axes=FALSE, plg=list(x=l.x, y=l.y), loc.main=loc.main)
text(x=l.xt+290000, y=l.yt, "combined percentiles")
add.plotsandlines(inset.cols=cols3)
if(pngs){dev.off()}
```

### Suppl Fig - Abundance-Only Hotspots

```{r Fig4b}
ra <- top.percentiles$`abundance-only`
## extract, zoom and shift inset maps
crop_and_shift(ra)

cols4 <- cols[1:4]#cols[c(1,2,4:5)]#

#### plot everything
if(pngs){png(paste0(pred.dir,"Circumpolar_abundance_only_hotspots.png"), width=png.width, height=png.height, res=png.res)}
par(plot.par)
plot(ra, main="high-abundance only hotspots", col=cols4, axes=FALSE, plg=list(x=l.x, y=l.y), loc.main=loc.main)
text(x=l.xt+290000, y=l.yt, "combined percentiles")
add.plotsandlines(inset.cols=cols4)
if(pngs){dev.off()}
```

### Suppl Fig - Proportionally High Richness

```{r Suppl Fig4a}
ra <- top.percentiles$`richness-only raw`
## extract, zoom and shift inset maps
crop_and_shift(ra)

#### plot everything
if(pngs){png(paste0(pred.dir,"Circumpolar_richness_only_hotspots_percentiles.png"), width=png.width, height=png.height, res=png.res)}
par(plot.par)
plot(ra, main="proportionally high richness", col=cols, axes=FALSE, plg=list(x=l.x, y=l.y), loc.main=loc.main)
text(x=l.xt, y=l.yt, "percentiles")
add.plotsandlines(inset.cols=cols)
if(pngs){dev.off()}
```

### Suppl Fig - Proportionally High Abundance

```{r Suppl Fig4b}
ra <- top.percentiles$`abundance-only raw`
## extract, zoom and shift inset maps
crop_and_shift(ra)

#### plot everything
if(pngs){png(paste0(pred.dir,"Circumpolar_abundance_only_hotspots_percentiles.png"), width=png.width, height=png.height, res=png.res)}
par(plot.par)
plot(ra, main="proportionally high abundance", col=cols, axes=FALSE, plg=list(x=l.x, y=l.y), loc.main=loc.main)
text(x=l.xt, y=l.yt, "percentiles")
add.plotsandlines(inset.cols=cols)
if(pngs){dev.off()}
```

### Suppl Fig - Beta-diversity

```{r}
ra <- median.betadiv*10000
range=c(0,max(values(ra), na.rm=TRUE))
## extract, zoom and shift inset maps
crop_and_shift(ra)

#### plot everything
if(pngs){png(paste0(pred.dir,"Circumpolar_betadiversity.png"), width=png.width, height=png.height, res=png.res)}
par(plot.par)
plot(ra, main="Betadiversity from median", col=cols.c.power, axes=FALSE, plg=plg.c, range=range, loc.main=loc.main)
text(x=l.x2, y=l.y2, "Local contribution to betadiversity")
text(x=l.x2, y=l.y2-500000, "x 10,000", cex=0.8)
# points(sp.samples, col="black", pch=19, cex=0.8)
add.plotsandlines(inset.cols=cols.c.power, plot.points=FALSE, range=range)
if(pngs){dev.off()}
```

```{r}
## extract, zoom and shift inset maps
crop_and_shift(top.percentiles$betadiversity)

#### plot everything
if(pngs){png(paste0(pred.dir,"Circumpolar_betadiversity_percentiles.png"), width=png.width, height=png.height, res=png.res)}
par(plot.par)
plot(top.percentiles$betadiversity, main="Betadiversity hotspots", col=cols, axes=FALSE, plg=list(x=l.x, y=l.y), loc.main=loc.main)
text(x=l.xt, y=l.yt, "percentiles")
add.plotsandlines()
if(pngs){dev.off()}
```



### Suppl Fig - Environmental Gaps

```{r Suppl FigX}
cols.gaps <- c("grey90","orange")

## extract, zoom and shift inset maps
crop_and_shift(gaps.hv.a)

#### plot everything
if(pngs){png(paste0(pred.dir,"Circumpolar_GapHypervolume_2km_All21SurveysCombined_AllVariablesExceptDepth2Canyons_Quantile095.png"), width=png.width, height=png.height, res=png.res)}
par(plot.par)
plot(gaps.hv.a, main="gaps based on sampled environmental conditions", col=cols.gaps, axes=FALSE, plg=list(x=l.x, y=l.y-100000), loc.main=loc.main, maxcell=12000000)
points(sp.samples, col="black", pch=19, cex=0.8)
add.plotsandlines(inset.cols=cols.gaps, plot.points=TRUE)
if(pngs){dev.off()}
```

```{r}
plot(gaps.hv.a, col=cols.gaps, xlim=c(-2800000,-2000000), ylim=c(800000,2500000))
points(sp.samples)
```

```{r}
plot(gaps.hv.a, col=cols.gaps, xlim=c(-400000,400000), ylim=c(-2150000,-1400000))
points(sp.samples)
```

### Suppl Fig - Richness-Uncertainty relationship

```{r}
ra <- top.percentiles$`high richness - low se`

## extract, zoom and shift inset maps
crop_and_shift(ra)

#### plot everything
if(pngs){png(paste0(pred.dir,"Circumpolar_Uncertainty_Richness_percentiles.png"), width=png.width, height=png.height, res=png.res)}
par(plot.par)
plot(ra, main="Top Richness & lowest SE", col=cols[1:3], axes=FALSE, plg=list(x=l.x, y=l.y), loc.main=loc.main)
text(x=l.xt, y=l.yt, "percentiles")
add.plotsandlines(inset.cols=cols[1:3])
if(pngs){dev.off()}

# ra <- low.richness.high.se
# 
# ## extract, zoom and shift inset maps
# crop_and_shift(ra)
# 
# #### plot everything
# if(pngs){png(paste0(pred.dir,"Circumpolar_Uncertainty_Richness_percentiles2.png"), width=png.width, height=png.height, res=png.res)}
# par(plot.par)
# plot(ra, main="Low Richness & highest SE", col=cols.c.power, axes=FALSE, plg=plg.c, range=range.ab, loc.main=loc.main)
# text(x=l.x2, y=l.y2, "%-cover")
# # points(sp.samples, col="black", pch=19, cex=0.8)
# add.plotsandlines(inset.cols=cols.c.power, plot.points=FALSE, range=range.ab)
# if(pngs){dev.off()}
```

### Suppl Fig - Abundance-Uncertainty relationship

```{r}
ra <- top.percentiles$`high abundance - low se`

## extract, zoom and shift inset maps
crop_and_shift(ra)

#### plot everything
if(pngs){png(paste0(pred.dir,"Circumpolar_Uncertainty_Abundance_percentiles.png"), width=png.width, height=png.height, res=png.res)}
par(plot.par)
plot(ra, main="Top Abundance & lowest SE", col=cols[1:3], axes=FALSE, plg=list(x=l.x, y=l.y), loc.main=loc.main)
text(x=l.xt, y=l.yt, "percentiles")
add.plotsandlines(inset.cols=cols[1:2])
if(pngs){dev.off()}

# ra <- low.abundance.high.se
# 
# ## extract, zoom and shift inset maps
# crop_and_shift(ra)
# 
# #### plot everything
# if(pngs){png(paste0(pred.dir,"Circumpolar_Uncertainty_Abundance_percentiles2.png"), width=png.width, height=png.height, res=png.res)}
# par(plot.par)
# plot(ra, main="Low Abundance & highest SE", col=cols.c.power, axes=FALSE, plg=plg.c, range=range.ab, loc.main=loc.main)
# text(x=l.x2, y=l.y2, "%-cover")
# # points(sp.samples, col="black", pch=19, cex=0.8)
# add.plotsandlines(inset.cols=cols.c.power, plot.points=FALSE, range=range.ab)
# if(pngs){dev.off()}
```

## Individual P/A maps {.tabset}

```{r, echo=FALSE, results='asis'}
## filenames
f.names <- paste0("Circumpolar_pa_median_",names(hmsc.maps.pa.env.median),".png")
if(pngs){
  for(i in 1:nlyr(hmsc.maps.pa.env.median)){
    ra <- hmsc.maps.pa.env.median[[i]]
    ## extract, zoom and shift inset maps
    crop_and_shift(ra)
     #### plot everything
    png(f.names[i], width=png.width, height=png.height, res=png.res)
    par(plot.par)
    plot(ra, main=names(ra), col=cols.c.power, axes=FALSE, plg=plg.c, range=range.pa, loc.main=loc.main)
    text(x=l.x2, y=l.y2, labels="Probability of presence")
    add.plotsandlines(inset.cols=cols.c.power, plot.points=FALSE, range=range.pa)
    dev.off()
  }
}
  
## Create tabs and embed images
for (i in 1:nlyr(hmsc.maps.pa.env.median)) {
  cat(paste0("### ", i, "\n"))
  cat(paste0("<img src='", f.names[i], "' alt='", gsub(".png", "", f.names[i]), "'>\n\n"))
}

# for(i in 1:10){
#   ra <- hmsc.maps.pa.env.median[[i]]
#   range <- c(0,1)#range(values(ra), na.rm=TRUE)
#   ## extract, zoom and shift inset maps
#   crop_and_shift(ra)
#    #### plot everything
#   par(mfrow=c(1,1), mar=c(3,0,1,0), oma=c(0,0,0,0))
#   plot(ra, main=names(ra), col=cols.c.power, axes=FALSE, plg=plg.c, range=range)
#   text(x=l.x2, y=l.y2, labels="Probability of presence")
#   add.plotsandlines(inset.cols=cols.c.power, plot.points=FALSE)
# }
```



## Environmental maps {.tabset}

### Depth

```{r}
#### hotspots on bathymetry
ra <- r.stack.subset$depth
range=range(values(ra), na.rm=TRUE)

## extract, zoom and shift inset maps
crop_and_shift(ra)#, top.biodiv.polygons)

if(pngs){png(paste0(pred.dir,"Circumpolar_environment_depth.png"), width=png.width, height=png.height, res=png.res)}
par(plot.par)
plot(ra, main="", col=cols.c, axes=FALSE, plg=plg.c, loc.main=loc.main)
# plot(top.biodiv.polygons, add=TRUE, col=cols.discr[1], border=NA)#, alpha=0.5)
text(x=l.x2, y=l.y2, "Depth (m)")
add.plotsandlines(inset.cols=cols.c.power, range=range)#, alpha=0.5)
if(pngs){dev.off()}
```

### Slope

```{r}
#### hotspots on bathymetry
ra <- r.stack.subset$slope
range=range(values(ra), na.rm=TRUE)

## extract, zoom and shift inset maps
crop_and_shift(ra)#, top.biodiv.polygons)

if(pngs){png(paste0(pred.dir,"Circumpolar_environment_slope.png"), width=png.width, height=png.height, res=png.res)}
par(plot.par)
plot(ra, main="", col=cols.c.power, axes=FALSE, plg=plg.c, loc.main=loc.main)
# plot(top.biodiv.polygons, add=TRUE, col=cols.discr[1], border=NA)#, alpha=0.5)
text(x=l.x2, y=l.y2, "Slope")
add.plotsandlines(inset.cols=cols.c.power, range=range)#, alpha=0.5)
if(pngs){dev.off()}
```

### TPI

```{r}
ra <- r.stack.subset$tpi
range=range(values(ra), na.rm=TRUE)

## Create a custom function to emphasize values near 0
emphasize_near_zero <- function(x) {
  sign(x) * sqrt(abs(x))
}
## Apply the custom function to your range of values
values <- seq(-60.5, 60.5, length.out = 100)
emphasized_values <- emphasize_near_zero(values)
## Create a color gradient function
color_gradient <- colorRampPalette(c("red", "white", "blue"))
## Generate a sequence of colors based on the emphasized values
cols.div <- color_gradient(100)[rank(emphasized_values)]

## extract, zoom and shift inset maps
crop_and_shift(ra)#, top.biodiv.polygons)

if(pngs){png(paste0(pred.dir,"Circumpolar_environment_tpi.png"), width=png.width, height=png.height, res=png.res)}
par(plot.par)
plot(ra, main="", col=cols.div, axes=FALSE, plg=plg.c, loc.main=loc.main, range=range)
# plot(top.biodiv.polygons, add=TRUE, col=cols.discr[1], border=NA)#, alpha=0.5)
text(x=l.x2, y=l.y2, "Topographic Position Index")
add.plotsandlines(inset.cols=cols.div, range=range)#, alpha=0.5)
if(pngs){dev.off()}
```

### Distance to canyons

```{r}
#### hotspots on bathymetry
ra <- r.stack.subset$distance2canyons/1000
range=range(values(ra), na.rm=TRUE)

## extract, zoom and shift inset maps
crop_and_shift(ra)#, top.biodiv.polygons)

if(pngs){png(paste0(pred.dir,"Circumpolar_environment_distance2canyons.png"), width=png.width, height=png.height, res=png.res)}
par(plot.par)
plot(ra, main="", col=cols.c, axes=FALSE, plg=plg.c, loc.main=loc.main)
# plot(top.biodiv.polygons, add=TRUE, col=cols.discr[1], border=NA)#, alpha=0.5)
text(x=l.x2, y=l.y2, "Distance to Canyons (km)")
add.plotsandlines(inset.cols=cols.c, range=range)#, alpha=0.5)
if(pngs){dev.off()}
```

### NPP (2002-2020 summer mean)

```{r}
ra <- r.stack.subset$npp_mean
range=range(values(ra), na.rm=TRUE)

## extract, zoom and shift inset maps
crop_and_shift(ra)#, top.biodiv.polygons)

if(pngs){png(paste0(pred.dir,"Circumpolar_environment_npp.png"), width=png.width, height=png.height, res=png.res)}
par(plot.par)
plot(ra, main="", col=cols.c, axes=FALSE, plg=plg.c, loc.main=loc.main)
# plot(top.biodiv.polygons, add=TRUE, col=cols.discr[1], border=NA)#, alpha=0.5)
text(x=l.x2, y=l.y2, "NPP (mg C/m^2/day)")
add.plotsandlines(inset.cols=cols.c, range=range)#, alpha=0.5)
if(pngs){dev.off()}
```

### Flux

```{r}
ra <- r.stack.subset$log.flux.mean
range=range(values(ra), na.rm=TRUE)

## extract, zoom and shift inset maps
crop_and_shift(ra)#, top.biodiv.polygons)

if(pngs){png(paste0(pred.dir,"Circumpolar_environment_flux.png"), width=png.width, height=png.height, res=png.res)}
par(plot.par)
plot(ra, main="", col=cols.c, axes=FALSE, plg=plg.c, loc.main=loc.main)
# plot(top.biodiv.polygons, add=TRUE, col=cols.discr[1], border=NA)#, alpha=0.5)
text(x=l.x2, y=l.y2, "log(seafloor food particle flux)")
add.plotsandlines(inset.cols=cols.c, range=range)#, alpha=0.5)
if(pngs){dev.off()}
```

### Sedimentation

```{r}
ra <- r.stack.subset$sed.mean
range=c(0,5000)

## extract, zoom and shift inset maps
crop_and_shift(ra)#, top.biodiv.polygons)

if(pngs){png(paste0(pred.dir,"Circumpolar_environment_sed.png"), width=png.width, height=png.height, res=png.res)}
par(plot.par)
plot(ra, main="", col=cols.c.power, axes=FALSE, plg=plg.c, loc.main=loc.main, range=range, fill_range=TRUE)
# plot(top.biodiv.polygons, add=TRUE, col=cols.discr[1], border=NA)#, alpha=0.5)
text(x=l.x2, y=l.y2, "particle sedimentation")
add.plotsandlines(inset.cols=cols.c.power, range=range, fill_range=TRUE)#, alpha=0.5)
if(pngs){dev.off()}
```

### Current speed mean

```{r}
ra <- r.stack.subset$seafloorcurrents_mean
range=range(values(ra), na.rm=TRUE)

## extract, zoom and shift inset maps
crop_and_shift(ra)#, top.biodiv.polygons)

if(pngs){png(paste0(pred.dir,"Circumpolar_environment_currmean.png"), width=png.width, height=png.height, res=png.res)}
par(plot.par)
plot(ra, main="", col=cols.c.power, axes=FALSE, plg=plg.c, loc.main=loc.main, range=range)
# plot(top.biodiv.polygons, add=TRUE, col=cols.discr[1], border=NA)#, alpha=0.5)
text(x=l.x2, y=l.y2, "m/s")
add.plotsandlines(inset.cols=cols.c.power, range=range)#, alpha=0.5)
if(pngs){dev.off()}
```

### Current speed residual

```{r}
ra <- r.stack.subset$seafloorcurrents_residual
range=range(values(ra), na.rm=TRUE)

## extract, zoom and shift inset maps
crop_and_shift(ra)#, top.biodiv.polygons)

if(pngs){png(paste0(pred.dir,"Circumpolar_environment_currresidual.png"), width=png.width, height=png.height, res=png.res)}
par(plot.par)
plot(ra, main="", col=cols.c.power, axes=FALSE, plg=plg.c, loc.main=loc.main, range=range)
# plot(top.biodiv.polygons, add=TRUE, col=cols.discr[1], border=NA)#, alpha=0.5)
text(x=l.x2, y=l.y2, "m/s")
add.plotsandlines(inset.cols=cols.c.power, range=range)#, alpha=0.5)
if(pngs){dev.off()}
```

### Temperature

```{r}
ra <- r.stack.subset$seafloortemperature
range=range(values(ra), na.rm=TRUE)

## extract, zoom and shift inset maps
crop_and_shift(ra)#, top.biodiv.polygons)

if(pngs){png(paste0(pred.dir,"Circumpolar_environment_temp.png"), width=png.width, height=png.height, res=png.res)}
par(plot.par)
plot(ra, main="", col=cols.c, axes=FALSE, plg=plg.c, loc.main=loc.main, range=range)
# plot(top.biodiv.polygons, add=TRUE, col=cols.discr[1], border=NA)#, alpha=0.5)
text(x=l.x2, y=l.y2, "degree C")
add.plotsandlines(inset.cols=cols.c, range=range)#, alpha=0.5)
if(pngs){dev.off()}
```

### Salinity

```{r}
ra <- r.stack.subset$seafloorsalinity
#range=range(values(ra), na.rm=TRUE)
range=c(33.8,35)
## extract, zoom and shift inset maps
crop_and_shift(ra)#, top.biodiv.polygons)

if(pngs){png(paste0(pred.dir,"Circumpolar_environment_sal.png"), width=png.width, height=png.height, res=png.res)}
par(plot.par)
plot(ra, main="", col=cols.c, axes=FALSE, plg=plg.c, loc.main=loc.main, range=range, fill_range=TRUE)
# plot(top.biodiv.polygons, add=TRUE, col=cols.discr[1], border=NA)#, alpha=0.5)
text(x=l.x2, y=l.y2, "")
add.plotsandlines(inset.cols=cols.c, range=range, fill_range=TRUE)#, alpha=0.5)
if(pngs){dev.off()}
```
