---
title: "display circumpolar biodiversity maps - file 2"
output:
  html_document:
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning=FALSE, message=FALSE, echo=FALSE,
                      dev='png', fig.height=10, fig.width=10, dpi=300)#,
#                      fig.path='display_predictive_maps_figures/') 


############################################################
# USER OPTIONS
############################################################
usr <- "JJ"

# Toggle writing figures to file
pngs <- FALSE
pdfs <- FALSE

# Spatial resolution
res <- "2km"

# Select the model folder to display in this report
# Examples:
#   "fam_cafe", "fam_cbpm", "fam_eppl", "fam_vpmg",
#   "npp_cafe", "npp_cbpm", "npp_eppl", "npp_vpmg"
model_id <- "npp_and_fam_cafe"

# Plot export settings
png.height <- 2400
png.width  <- 2400
png.res    <- 300
pdf.height <- 8
pdf.width  <- 8

############################################################
# LOAD SHARED PROJECT SETUP FILES
############################################################
# Look for 0_SourceFile.R and 0_Functions.R either in the current folder or one level above.
source_file <- if (file.exists("0_SourceFile.R")) {"0_SourceFile.R"
} else if (file.exists(file.path("..", "0_SourceFile.R"))) {file.path("..", "0_SourceFile.R")
} else {stop("Could not find 0_SourceFile.R in the current or parent directory.")}

functions_file <- if (file.exists("0_Functions.R")) {"0_Functions.R"
} else if (file.exists(file.path("..", "0_Functions.R"))) {file.path("..", "0_Functions.R")
} else {stop("Could not find 0_Functions.R in the current or parent directory.")}

source(source_file)
source(functions_file)


############################################################
# PACKAGES
############################################################
library(Hmsc)
library(terra)
library(SOmap)
library(viridis)
library(knitr)
library(RColorBrewer)
library(vioplot)
'%!in%' <- function(x, y) !('%in%'(x, y))

############################################################
# DIRECTORY DEFINITIONS
############################################################

# Environmental rasters
env.derived <- file.path(usr.dropbox.dir,
  "data_environmental/derived")

# Biological data
bio.dir <- file.path(usr.dropbox.dir, 
  "data_biological")

# Predictive-map root directory
pred.dir <- file.path(usr.dropbox.dir,
  "data_products/predictive_maps/circum_antarctic")

# Model-specific directory containing species maps and total abundance
pred.model.dir <- file.path(pred.dir, paste0("hmsc_with_", model_id))

# Subdirectories for current derived outputs
cont.dir   <- file.path(pred.model.dir, "continuous")
pct.dir    <- file.path(pred.model.dir, "percentiles")
hot.dir    <- file.path(pred.model.dir, "hotspots")
domain.dir <- file.path(pred.model.dir, "domains")
poly.dir   <- file.path(pred.model.dir, "polygons")

# Directory for any figure exports created by this report
export.dir <- file.path(usr.dropbox.dir,
  "data_products/publication_figures/JansenEtAl", model_id, "/")
dir.create(export.dir, recursive = TRUE, showWarnings = FALSE)

############################################################
# LOAD VARIABLES AND SETTINGS
############################################################
setup1_file <- if (file.exists("b2_setup1_JansenEtAl2026.R")) {"b2_setup1_JansenEtAl2026.R"
} else if (file.exists(file.path("biodiversity", "b2_setup1_JansenEtAl2026.R"))) {file.path("biodiversity", "b2_setup1_JansenEtAl2026.R")}

setup2_file <- if (file.exists("b2_setup2_JansenEtAl2026.R")) {"b2_setup2_JansenEtAl2026.R"
} else if (file.exists(file.path("biodiversity", "b2_setup2_JansenEtAl2026.R"))) {file.path("biodiversity", "b2_setup2_JansenEtAl2026.R")}

source(setup1_file)
source(setup2_file)
```

# Displaying predictive maps {.tabset}

## MPA relationships {.tabset}

```{r, echo=FALSE}
## Emma's files:
mpa.raw <- vect(paste0(sci.dir,"data_environmental/ccamlr_data_2025/MPAs_within_CCAMLR_Planning_Domains/MPAs_within_CCAMLR_Planning_Domains.shp"))
mpa.proposed <- vect(paste0(sci.dir,"data_environmental/ccamlr_data_2025/Proposed_MPAs/Proposed_MPAs.shp"))

## subset MPA to South Orkney and Ross Sea no-take
mpa <- mpa.raw[mpa.raw$ZoneType%in%"No-take"]
mpa <- mpa[mpa$MPA%in%c("South Orkney Islands Southern Shelf MPA", "Ross Sea Region MPA")]

#### East Antarctica
mpa.proposed.EA <- mpa.proposed[which(mpa.proposed$MPA%in%"East Antarctic Proposed MPA")]
#### AP
mpa.proposed.AP <- mpa.proposed[which(mpa.proposed$MPA%in%"Domain 1 Proposed MPA")]
#### Weddell Sea Phase 1
mpa.proposed.WS1 <- mpa.proposed[which(mpa.proposed$MPA%in%"Weddell Sea Proposed MPA Phase 1")]
#### Weddell Sea Phase 2
mpa.proposed.WS2 <- mpa.proposed[which(mpa.proposed$MPA%in%"Weddell Sea Proposed MPA Phase 2")]

#### and simplify to outer boundaries only
# Use the aggregate function to remove inner polygons
mpa.proposed.clean <- aggregate(mpa.proposed, dissolve = TRUE)
mpa.clean <- aggregate(mpa, dissolve = TRUE)
# # Use the union function to merge polygons that share a line
mpa.clean.merged <- union(mpa.clean)
```

### Percentiles

```{r, echo=FALSE}
ra <- top.percentiles$biodiversity
levels(ra)[[1]]$biodiversity[2:5] <- paste0("   ",levels(ra)[[1]]$biodiversity[2:5])

## extract, zoom and shift inset maps
crop_and_shift(ra)
#### plot everything
if(pngs){png(paste0(pred.dir,"Circumpolar_MPA_biodiversity_percentiles.png"), width=2400, height=2400, res=300)}
par(plot.par)
plot(ra, main="", col=cols, axes=FALSE, plg=list(x=l.x, y=l.y-100000), loc.main=loc.main)
text(x=l.xt, y=l.yt-100000, "percentiles")
plot(mpa.clean, add=TRUE, col="orange2", alpha=0.3)#, border=NA)
plot(mpa.proposed.clean, add=TRUE, col="red", alpha=0.1)#, border=NA)
#add.plotsandlines()
plot(coast.proj[values(coast.proj)=="land"], col="grey60", add=TRUE)
lines(coast.proj, col="grey40")
if(pngs){dev.off()}
```

### AES

```{r, echo=FALSE}
ra <- r2
range=range(values(ra), na.rm=TRUE)

if(pngs){png(paste0(pred.dir,"Circumpolar_MPA_biodiversity_AES.png"), width=2400, height=2400, res=300)}
par(plot.par)
plot(ra, main="", col=cols.grey, axes=FALSE, plg=plg.c, loc.main=loc.main)
plot(top.biodiv.polygons, add=TRUE, col=cols.discr[1], border=NA)#, alpha=0.5)
text(x=l.x2, y=l.y2, "Depth (m)")
plot(mpa.clean, add=TRUE, col="orange", alpha=0.3)
plot(mpa.proposed.clean, add=TRUE, col="red", alpha=0.1)
#add.plotsandlines()
plot(coast.proj[values(coast.proj)=="land"], col="grey60", add=TRUE)
lines(coast.proj, col="grey40")
if(pngs){dev.off()}
```

### Frequencies of hotspot occurrence inside/outside of MPAs

```{r, echo=FALSE}
#### frequencies of the full raster, should be as expected
freq_raster <- freq(top.percentiles$biodiversity)
tot.cells <- sum(freq_raster$count)
freq_raster$prop <- round(freq_raster$count/tot.cells,2)

#### calculate for existing mpa, proposed mpa, and unprotected areas
current.mpa.biodiv <- mask(top.percentiles$biodiversity, mpa)
freq_current <- freq(current.mpa.biodiv)
# freq_current$summed <- NA
# freq_current$summed[1] <- freq_current$count
# for(i in 2:5){freq_current$summed[i] <- sum(freq_current$summed[i-1], freq_current$count[i])}
# freq_current$summed.prop <- round(freq_current$summed/sum(freq_current$count),4)
# freq_current$summed.prop.of.total <- round(freq_current$summed/tot.cells,4)
freq_current$prop <- round(freq_current$count/sum(freq_current$count),4)
freq_current$prop.of.total <- round(freq_current$count/tot.cells,4)

proposed.mpa.biodiv <- mask(top.percentiles$biodiversity, mpa.proposed)
freq_proposed <- freq(proposed.mpa.biodiv)
freq_proposed$prop <- round(freq_proposed$count/sum(freq_proposed$count),4)
freq_proposed$prop.of.total <- round(freq_proposed$count/tot.cells,4)

proposed.EA.mpa.biodiv <- mask(top.percentiles$biodiversity, mpa.proposed.EA)
freq_proposed_EA <- freq(proposed.EA.mpa.biodiv)
freq_proposed$prop.EA <- round(freq_proposed_EA$count/sum(freq_proposed_EA$count),4)
freq_proposed$prop.EA.of.total <- round(freq_proposed_EA$count/tot.cells,4)

outside.mpa.biodiv1 <- mask(top.percentiles$biodiversity, mpa.proposed, inverse=TRUE)
outside.mpa.biodiv <- mask(outside.mpa.biodiv1, mpa, inverse=TRUE)
freq_outside <- freq(outside.mpa.biodiv)
freq_outside$prop <- round(freq_outside$count/sum(freq_outside$count),4)
freq_outside$prop.of.total <- round(freq_outside$count/tot.cells,4)

# ## masking needs to exclude cells with center-points outside the line
# current.mpa.biodiv2 <- mask(top.biodiv, mpa, touches=FALSE)
# freq_current2 <- freq(current.mpa.biodiv2)
# inside.mpa.biodiv2 <- mask(top.biodiv, mpa.proposed, touches=FALSE)
# freq_within2 <- freq(inside.mpa.biodiv2)
# outside.mpa.biodiv2 <- mask(top.biodiv, mpa.proposed, inverse=TRUE, touches=FALSE)
# freq_outside2 <- freq(outside.mpa.biodiv2)
```

##### Cell-counts raw: per managment zone (75th are 15% of cells, 90th are 5% of cells, 95th are 4% of cells)

```{r}
freq_raw.count1 <- cbind(freq_raster[2:3], freq_current$count, freq_proposed$count, freq_outside$count, freq_proposed_EA$count)
freq_raw.count.sums <- c(NA ,colSums(freq_raw.count1[,2:6]))
freq_raw.count <- rbind(freq_raw.count.sums, freq_raw.count1)
names(freq_raw.count) <- c("percentile","all","current","proposed","outside","proposed.EA")
freq_raw.count[1,1] <- "Total"
freq_raw.count
```

##### Cell-counts: per management zone

```{r}
freq_count <- rbind(freq_raw.count)
freq_count[3,2:6] <- colSums(freq_count[3:6,2:6])
freq_count[4,2:6] <- colSums(freq_count[4:6,2:6])
freq_count[5,2:6] <- colSums(freq_count[5:6,2:6])
freq_count
```

```{r}
print("% of hotspots currently protected")
freq_count[4,3]/freq_count[4,2]

print("% of hotspots protected when including proposed MPAs")
sum(freq_count[4,3:4])/freq_count[4,2]

print("% of hotspots protected from proposed East Antarctic MPA")
sum(freq_count[4,6])/freq_count[4,2]

## how much is shallower than 550m, and hotspots in there
depth550 <- r2
depth550[values(depth550)<(-550)] <- NA
depth.biodiv <- mask(top.percentiles$biodiversity, depth550)
##
freq_depth <- freq(depth.biodiv)
print("% of hotspots shallower than 550m")
(sum(freq_depth[3:5,3]))/freq_count[4,2]
## which hotspots are in mpas
#current.mpa.biodiv
## hotspots in either shallow/mpa or both
sel.depth.mpa <- which(depth.biodiv[]%in%2:4|current.mpa.biodiv[]%in%2:4)
current.mpa.or.shallow <- top.percentiles$biodiversity
current.mpa.or.shallow[] <- NA
current.mpa.or.shallow[sel.depth.mpa] <- 1
freq.current.mpa.or.shallow <- freq(current.mpa.or.shallow)

print("% of hotspots in MPAs or shallower than 550m")
(freq.current.mpa.or.shallow[3])/freq_count[4,2]

```


##### Proportions: per management zone

```{r}
freq_prop <- cbind(freq_count$percentile,round(freq_count[,2:6]/freq_count[,2],4))
names(freq_prop) <- names(freq_raw.count)
freq_prop
```

##### Proportions: per managment zone relative to total shelf area

```{r}
freq_prop.shelf <- cbind(freq_count$percentile,round(freq_count[,2:6]/freq_count[1,2],5))
names(freq_prop.shelf) <- names(freq_raw.count)
freq_prop.shelf
```

##### Proportions: per managment zone relative to managment zone area

```{r}
divide_by_first_row <- function(column) {column[2:6] / column[1]}
freq_prop.zone <- cbind(freq_count$percentile[2:6],round(as.data.frame(lapply(freq_count[,2:5], divide_by_first_row)),4))
freq_prop.zone
```

##### km\^2 area: per management zone (in 1000 km\^2)

```{r}
## as absolute values in km^2 (*1000)
freq_area.1000km <- freq_count
freq_area.1000km[,2] <- round((freq_count[,2]/4)/1000,3)
freq_area.1000km[,3] <- round((freq_count[,3]/4)/1000,3)
freq_area.1000km[,4] <- round((freq_count[,4]/4)/1000,3)
freq_area.1000km[,5] <- round((freq_count[,5]/4)/1000,3)
(freq_area.1000km)
```

```{r}
dat <- as.data.frame(freq_prop.zone[,3:5])
names(dat) <- c("Existing MPA","Proposed MPA","Not protected")
```

```{r, fig.width=7, fig.height=5}
par(mar=c(5,4,4,7),oma=c(0,0,0,10))
plot(1:ncol(dat), dat[1,], type = "n", xaxt = "n", xlab = "Management Zones", ylab = "Proportion of area", main = "", ylim=c(0,0.35))
axis(1, at = 1:ncol(dat), labels = names(dat))
# add percentile lines
abline(0.01,0, lty=2)
abline(0.05,0, lty=2)
abline(0.10,0, lty=2)
abline(0.25,0, lty=2)
# Plot each row as a line
for (i in 1:nrow(dat)) {
  lines(1:ncol(dat), dat[i,], col = cols[i], lwd = 3, add=TRUE)
}
# Add a legend
legend("topright", legend = c("75th", "90th", "95th", "99th"), col = cols[2:5], lwd = 3, xpd = TRUE, inset = c(-0.4, 0), title="percentile", bty="n")
```

```{r, fig.width=7, fig.height=5}
mpa.mat <- as.matrix(t(freq_count[4,3:4]/freq_count[4,2]))
colnames(mpa.mat) <- ""
rownames(mpa.mat)[1] <- "existing"
transp.red <- adjustcolor("red", alpha.f = 0.1)
transp.orange <- adjustcolor("orange2", alpha.f = 0.3)
barplot(mpa.mat, 
        col = c(transp.orange, transp.red), # Colors for the stacked segments
        ylab = "proportion of hotspots covered",
        legend.text = rownames(mpa.mat), # Legend labels for the stacked segments
        args.legend = list(x = "topright", bty = "n"))
```


### Frequencies of hotspot percentiles inside/outside of MPAs

```{r, echo=FALSE}
## Define the inset plotting region (adjust the coordinates as needed)
inset_x <- 0.49
inset_y <- 0.43
inset_width <- 0.1
inset_height <- 0.3
```

```{r, echo=FALSE}
#### prep:
ra <- top.percentiles$biodiversity
levels(ra)[[1]]$biodiversity[2:5] <- paste0("   ",levels(ra)[[1]]$biodiversity[2:5])

## extract, zoom and shift inset maps
crop_and_shift(ra)

#### plot everything
if(pngs){png(paste0(pred.dir,"Circumpolar_MPA_biodiversity_coverage.png"), width=2400, height=2400, res=300)}
if(pdfs){pdf("Circumpolar_MPA_biodiversity_coverage.pdf", width=8, height=8)}
par(plot.par)
plot(ra, main="", col=cols, axes=FALSE, plg=list(x=l.x-100000, y=l.y-100000), loc.main=loc.main)
text(x=l.xt-100000, y=l.yt-100000, "percentiles")
plot(mpa.clean, add=TRUE, col="orange2", alpha=0.3)#, border=NA)
plot(mpa.proposed.clean, add=TRUE, col="red", alpha=0.1)#, border=NA)
#add.plotsandlines()
plot(coast.proj[values(coast.proj)=="land"], col="grey60", add=TRUE)
lines(coast.proj, col="grey40")
text(x=-200000, y=650000, "% of hotspots covered", srt=90, cex=0.8)
text(x=1000000, y=150000, "MPAs existing   ", cex=0.8)
text(x=1000000, y=1000000, "MPAs proposed", cex=0.8)

## Save current plotting parameters
old_par <- par(no.readonly = TRUE)

## Define the new plotting region
par(fig = c(inset_x, inset_x + inset_width, inset_y, inset_y + inset_height), new = TRUE)

## plot inset graphic
par(mar=c(2,2,0,1))
barplot(mpa.mat*100, col="white", cex.axis=0.8)
barplot(mpa.mat*100, col=c(transp.orange, transp.red), add=TRUE, axes=FALSE)

# plot(1:ncol(dat), dat[1,], type = "n", xaxt = "n", yaxt = "n", xlab = "", ylab = "", main = "", ylim=c(0,0.15))
# rect(par("usr")[1], par("usr")[3], par("usr")[2], par("usr")[4], col = "white", border = NA)
# axis(1, at = 1:ncol(dat), labels = names(dat), cex.axis=0.7, padj=-1)
# axis(2, at=c(0,0.05,0.1,0.15,0.2,0.25,0.3), labels=c(0,0.05,0.1,0.15,0.2,0.25,0.3), cex.axis=0.7, padj=1)
# mtext("Proportion of area covered", side=2, line=1.5, cex=0.8)
# ## add percentile lines
# abline(0.01,0, lty=2)
# abline(0.05,0, lty=2)
# abline(0.10,0, lty=2)
# abline(0.25,0, lty=2)
# ## Plot each row as a line
# for (i in 1:nrow(dat)) {
#   lines(1:ncol(dat), dat[i,], col = cols[i], lwd = 3, add=TRUE)
# }
## Add a legend
#legend("topright", legend = c("75th", "90th", "95th", "99th"), col = cols[2:5], lwd = 3, xpd = TRUE, inset = c(-0.4, 0), title="percentile", bty="n")
if(pngs|pdfs){dev.off()}
```

```{r, echo=FALSE, eval=FALSE}
#### prep:
ra <- r2
range=range(values(ra), na.rm=TRUE)
## extract, zoom and shift inset maps
crop_and_shift(ra)

#### plot everything
par(plot.par)
plot(ra, main="", col=cols.grey, axes=FALSE, plg=plg.c, loc.main=loc.main)
text(x=l.x2, y=l.y2, "Depth (m)")
plot(top.biodiv.polygons, add=TRUE, col=cols.discr[1], border=NA)#, alpha=0.5)
plot(mpa.clean, add=TRUE, col="orange2", alpha=0.3)#, border=NA)
plot(mpa.proposed.clean, add=TRUE, col="red", alpha=0.1)#, border=NA)
#add.plotsandlines()
plot(coast.proj[values(coast.proj)=="land"], col="grey60", add=TRUE)
lines(coast.proj, col="grey40")

## Save current plotting parameters
old_par <- par(no.readonly = TRUE)

## Define the new plotting region
par(fig = c(inset_x, inset_x + inset_width, inset_y, inset_y + inset_height), new = TRUE)

## plot inset graphic
par(mar=c(2,2,0,1))
plot(1:ncol(dat), dat[1,], type = "n", xaxt = "n", yaxt = "n", xlab = "", ylab = "", main = "", ylim=c(0,0.15))
rect(par("usr")[1], par("usr")[3], par("usr")[2], par("usr")[4], col = "white", border = NA)
axis(1, at = 1:ncol(dat), labels = names(dat), cex.axis=0.7, padj=-1)
axis(2, at=c(0,0.05,0.1,0.15,0.2,0.25,0.3), labels=c(0,0.05,0.1,0.15,0.2,0.25,0.3), cex.axis=0.7, padj=1)
mtext("Proportion of area covered", side=2, line=1.5, cex=0.8)
## add percentile lines
abline(0.01,0, lty=2)
abline(0.05,0, lty=2)
abline(0.10,0, lty=2)
abline(0.25,0, lty=2)
## Plot each row as a line
for (i in 1:nrow(dat)) {
  lines(1:ncol(dat), dat[i,], col = cols[i], lwd = 3, add=TRUE)
}
```


### Other proportions

#### Hotspot abundances vs other areas

```{r}
## number of cells in top 10%
(cell.count.top10 <- sum(freq_raw.count[4:6,2]))
(cell.count.bottom90 <- sum(freq_raw.count[2:3,2]))

## abundance inside/outside the top 10%
abund.inside <- mask(hmsc.maps.totalabundance.env$median_corrected, top.biodiv.polygons)/540
abund.outside <- mask(hmsc.maps.totalabundance.env$median_corrected, top.biodiv.polygons, inverse=TRUE)/540

## number of cells in top 10% & bottom 90%, and equivalent area (each cell is 4km^2)
(mean.inside <- mean(values(abund.inside, na.rm=TRUE)))
(tot.area.top10 <- cell.count.top10*4)
(tot.area.top10.covered <- tot.area.top10*mean.inside)

(mean.outside <- mean(values(abund.outside, na.rm=TRUE)))
(tot.area.bottom90 <- cell.count.bottom90*4)
(tot.area.bottom90.covered <- tot.area.bottom90*mean.outside)

## proportion of cells that are shallower than 550m and therefore already protected
depth.inside <- mask(r2, top.biodiv.polygons)
depth.outside <- mask(r2, top.biodiv.polygons, inverse=TRUE)

length(which(values(depth.inside, na.rm=TRUE)>=-550))/cell.count.top10
```









## CCAMLR Planning Domains {.tabset}

```{r}
domain.dat.proj <- vect(paste0(sci.dir,"data_environmental/ccamlr_data_2025/geographical_data/mpapd/CCAMLR_MPAPD_EPSG6932.shp"))
crs(domain.dat.proj) <- "EPSG:6932"
domain.dat <- project(domain.dat.proj, hmsc.maps.richness.env$median)
cex=0.9
```

#### Comparing differences in each Planning Domain: circumpolar AES vs regional AES vs other areas

```{r}
## load environmental values for hotspots in each planning domain
load("C:/Users/jjansen/OneDrive - University of Tasmania/science/data_biological/BiodiversityHotspots_EnvironmentalSetting_Domains.Rdata")
```

### Compiled

```{r}
ra <- top.percentiles.domains

## the percentile descriptions create problems here, so deleting them and keeping the values 0-4 only:
levels_list <- levels(ra)
#Function to remove descriptions and keep only values
clean_levels <- function(levels_df) {
  levels_df$value <- as.numeric(levels_df$value)
  levels_df$value2 <- as.numeric(levels_df$value)  # Add a placeholder column
  return(levels_df[, c("value", "value2")])
}
#Apply the function to each level
cleaned_levels <- lapply(levels_list, clean_levels)
#Set the cleaned levels back to the SpatRaster object
levels(ra) <- cleaned_levels

## combine domains
ra.rich <- sum(ra[[1:6]], na.rm=TRUE)
ra.abund <- sum(ra[[7:12]], na.rm=TRUE)
ra.biodiv <- sum(ra[[13:18]], na.rm=TRUE)

## small overlap between 1-2, 4-5, 5-6, 6-1
sel1 <- which(!is.na(sum(ra[[1:2]])[]))
sel4 <- which(!is.na(sum(ra[[4:5]])[]))
sel5 <- which(!is.na(sum(ra[[5:6]])[]))
sel6 <- which(!is.na(sum(ra[[c(6,1)]])[]))
ra.rich[sel1] <- ra[[1]][sel1]
ra.rich[sel4] <- ra[[4]][sel4]
ra.rich[sel5] <- ra[[5]][sel5]
ra.rich[sel6] <- ra[[6]][sel6]
ra.abund[sel1] <- ra[[7]][sel1]
ra.abund[sel4] <- ra[[10]][sel4]
ra.abund[sel5] <- ra[[11]][sel5]
ra.abund[sel6] <- ra[[12]][sel6]
ra.biodiv[sel1] <- ra[[13]][sel1]
ra.biodiv[sel4] <- ra[[16]][sel4]
ra.biodiv[sel5] <- ra[[17]][sel5]
ra.biodiv[sel6] <- ra[[18]][sel6]
```

```{r, echo=FALSE}
ra <- ra.biodiv
levels(ra) <- as.factor(c("0-74th","75-89th", "90-94th", "95-98th", "99th"))
ra <- extend(ra, ext(-3400000,4300000,-3400000,4600000), fill=NA)

par(plot.par)
plot(coast.proj[values(coast.proj)=="land"], col=land.col, axes=FALSE, xlim=c(-3400000,4000000), ylim=c(-3400000,4600000))
plot(ra, main="", col=cols, plg=list(x=-300000, y=500000), loc.main=loc.main, add=TRUE) #, axes=FALSE
plot(domain.dat[-which(domain.dat$Location%in%c("Kerguelen Plateau", "North Scotia Arc", "del Cano - Crozet"))], add=TRUE) #, axes=FALSE
lines(coast.proj, col=coast.col)
text(x=0, y=600000, "percentiles")
text(x=-2800000, y=600000, "Domain 1")
text(x=-1400000, y=2000000, "Domain 3")
text(x=100000, y=2800000, "Domain 4")
text(x=2800000, y=-1600000, "Domain 7")
text(x=-200000, y=-2600000, "Domain 8")
text(x=-2500000, y=-1000000, "Domain 9")










#### plot everything
#if(pngs){png(paste0(pred.dir,"Circumpolar_MPA_domains_biodiversity_percentiles.png"), width=2400, height=2400, res=300)}
par(plot.par)
plot(ra, main="", col=cols, axes=FALSE, plg=list(x=l.x, y=l.y), loc.main=loc.main)
plot(domain.dat[-which(domain.dat$Location%in%c("Kerguelen Plateau", "North Scotia Arc", "del Cano - Crozet"))], axes=FALSE, xlim=ext(ra)[1:2], ylim=ext(ra)[3:4], add=TRUE)
text(x=0, y=0, "percentiles")

## add coastline
coast.col="grey40"
land.col="grey60"
plot(coast.proj[values(coast.proj)=="land"], col=land.col, add=TRUE)
lines(coast.proj, col=coast.col)

# plot(mpa.clean, add=TRUE, col="orange2", alpha=0.3)#, border=NA)
# plot(mpa.proposed.clean, add=TRUE, col="red", alpha=0.1)#, border=NA)
# add.plotsandlines()
# plot(coast.proj[values(coast.proj)=="land"], col="grey60", add=TRUE)
# lines(coast.proj, col="grey40")
#if(pngs){dev.off()}
```

```{r, echo=FALSE}
# Variable names and labels for violin plots
var_names <- c("depth", "curr", "temp", "flux")
y_labels <- c(
  "Depth (m)",
  "Mean seafloor current speed (m/s)",
  "Seafloor temperature (\u00B0C)",
  "log(seafloor particle flux)"
)
```

```{r, echo=FALSE}
ra <- ra.rich

## extract, zoom and shift inset maps
crop_and_shift(ra)
#### plot everything
if(pngs){png(paste0(pred.dir,"Circumpolar_MPA_domains_richness_percentiles.png"), width=2400, height=2400, res=300)}
par(plot.par)
plot(domain.dat[-which(domain.dat$Location%in%c("Kerguelen Plateau", "North Scotia Arc", "del Cano - Crozet"))], axes=FALSE, xlim=ext(ra)[1:2], ylim=ext(ra)[3:4])
plot(ra, main="", col=cols, axes=FALSE, plg=list(x=l.x, y=l.y), loc.main=loc.main, add=TRUE)
text(x=l.xt, y=l.yt, "percentiles")
# plot(mpa.clean, add=TRUE, col="orange2", alpha=0.3)#, border=NA)
# plot(mpa.proposed.clean, add=TRUE, col="red", alpha=0.1)#, border=NA)
add.plotsandlines()
# plot(coast.proj[values(coast.proj)=="land"], col="grey60", add=TRUE)
# lines(coast.proj, col="grey40")
if(pngs){dev.off()}
```

### Domain 1

```{r, fig.height=10, fig.width=10}
ra <- top.percentiles.domains$`biodiversity domain 1`
masked.ra <- mask(top.percentiles$biodiversity,ra)

xlim=c(-2750000,-1700000)
ylim=c(150000,2500000)
mar <- c(8,1,2,0)
inset=c(0,-0.19)
# l.x.dom=-2350000
# l.y.dom=1600000

if(pngs){png(paste0(pred.dir,"Circumpolar_Domain1_biodiversity_percentiles.png"), width=2400, height=3000, res=300)}
par(mfrow=c(1,2))
plot(masked.ra, main="Circumpolar Biodiversity Hotspots",
     xlim=xlim, ylim=ylim, col=cols, legend=FALSE, mar=mar)#, axes=FALSE)
legend("bottom", inset=inset, legend=levels(masked.ra)[[1]][,2], fill=cols, xpd=NA, cex=cex)#, bty="n")
# plot(coast.proj[values(coast.proj)=="land"], col=land.col, add=TRUE)
lines(coast.proj, col=coast.col)
plot(ra, main="Domain 1 Biodiversity Hotspots", 
     xlim=xlim, ylim=ylim, col=cols, legend=FALSE, mar=mar)#, axes=FALSE)
legend("bottom", inset=inset, legend=levels(ra)[[1]][,2], fill=cols, xpd=NA, cex=cex)
# text(x=l.xt, y=l.yt, "percentiles")
# add.plotsandlines()
lines(coast.proj, col=coast.col)
if(pngs){dev.off()}
```

```{r, fig.height=4, fig.width=10}
# Domain to plot
dom <- "dom1"

# Set up plotting area
par(mfrow = c(1, 4), mgp = c(1, 1, 0))
# Loop through variables
for (i in seq_along(var_names)) {
  # Combine all values to set y-axis limits
  all_vals <- c(
    vals.domtop[[dom]][[i]],
    vals.domtopout[[dom]][[i]],
    vals.circtop[[dom]][[i]],
    vals.circtopout[[dom]][[i]]
  )
  # Create empty plot
  plot(1, type = "n", xlim = c(0.5, 2.5), ylim = range(all_vals, na.rm = TRUE),
       xaxt = "n", xlab = "", ylab = "", main = "")
  # Add half violins
  vioplot(vals.domtop[[dom]][[i]],    at = 1, side = "left",  col = col.in, add = TRUE)
  vioplot(vals.domtopout[[dom]][[i]], at = 1, side = "right", col = col.out, add = TRUE)
  vioplot(vals.circtop[[dom]][[i]],   at = 2, side = "left",  col = col.in, add = TRUE)
  vioplot(vals.circtopout[[dom]][[i]],at = 2, side = "right", col = col.out, add = TRUE)
  # Add axis and title
  axis(1, at = 1:2, labels = c("Domain", "Circumpolar"))
  title(ylab = y_labels[i], line = 3)
}
```

```{r, fig.height=4, fig.width=10}
# Domain to plot
dom <- "dom1"

# Set up plotting area
par(mfrow = c(1, 4), mgp = c(1, 1, 0))
# Loop through variables
for (i in seq_along(var_names)) {
  # Create empty plot space
  plot(1, type = "n", xlim = c(0.5, 3.5), ylim = range(
    vals.domtop[[dom]][[i]],
    vals.domtopout[[dom]][[i]],
    vals.circtop[[dom]][[i]],
    na.rm = TRUE
  ), xaxt = "n", xlab = "", ylab = "", main = "")
  # Add half violins
  vioplot(vals.domtop[[dom]][[i]], at = 1, side = "left", col = col.in, add = TRUE)
  vioplot(vals.domtopout[[dom]][[i]], at = 2, side = "left", col = col.out, add = TRUE)
  vioplot(vals.circtop[[dom]][[i]], at = 3, side = "left", col = col.in, add = TRUE)
  # Add axis and title
  axis(1, at = 1:3, labels = c("Dom-in", "Dom-out", "Circ-in"))
  title(ylab = y_labels[i], line = 3)
}
```

### Domain 3

```{r, fig.height=10, fig.width=10}
ra <- top.percentiles.domains$`biodiversity domain 3`
masked.ra <- mask(top.percentiles$biodiversity,ra)

xlim=c(-2500000,-600000)
ylim=c(750000,1850000)
inset=c(0,0)
# l.x.dom=-1900000
# l.y.dom=1700000

if(pngs){png(paste0(pred.dir,"Circumpolar_Domain3_biodiversity_percentiles.png"), width=2400, height=3000, res=300)}
par(mfrow=c(2,1))
plot(masked.ra, main="Circumpolar Biodiversity Hotspots",
     xlim=xlim, ylim=ylim, col=cols, legend=FALSE)
legend("top", inset=inset, legend=levels(masked.ra)[[1]][,2], fill=cols, cex=cex, bty="n")
lines(coast.proj, col=coast.col)
plot(ra, main="Domain 3 Biodiversity Hotspots",
     xlim=xlim, ylim=ylim, col=cols, legend=FALSE)
legend("top", inset=inset, legend=levels(ra)[[1]][,2], fill=cols, cex=cex, bty="n")
lines(coast.proj, col=coast.col)
# text(x=l.xt, y=l.yt, "percentiles")
# add.plotsandlines()
if(pngs){dev.off()}
```

```{r, fig.height=4, fig.width=10}
# Domain to plot
dom <- "dom3"

# Set up plotting area
par(mfrow = c(1, 4), mgp = c(1, 1, 0))
# Loop through variables
for (i in 1:11) {
  # Combine all values to set y-axis limits
  all_vals <- c(
    vals.domtop[[dom]][[i]],
    vals.domtopout[[dom]][[i]],
    vals.circtop[[dom]][[i]],
    vals.circtopout[[dom]][[i]]
  )
  # Create empty plot
  plot(1, type="n", xlim=c(0.5, 2.5), ylim=range(all_vals, na.rm=TRUE), xaxt="n", xlab="", ylab="", main="")
  
  # Add half violins
  vioplot(vals.domtop[[dom]][[i]],    at = 1, side = "left",  col = col.in, add = TRUE)
  vioplot(vals.domtopout[[dom]][[i]], at = 1, side = "right", col = col.out, add = TRUE)
  vioplot(vals.circtop[[dom]][[i]],   at = 2, side = "left",  col = col.in, add = TRUE)
  vioplot(vals.circtopout[[dom]][[i]],at = 2, side = "right", col = col.out, add = TRUE)
  # Add axis and title
  axis(1, at = 1:2, labels = c("Domain", "Circumpolar"))
  title(ylab = y_labels[i], line = 3)
}
```

### Domain 4

```{r, fig.height=10, fig.width=10}
ra <- top.percentiles.domains$`biodiversity domain 4`
masked.ra <- mask(top.percentiles$biodiversity,ra)

xlim=c(-700000,1200000)
ylim=c(1700000,2650000)
inset=c(0,0.02)
mar=c(3,2,2,1)
# l.x.dom=-100000
# l.y.dom=2000000

if(pngs){png(paste0(pred.dir,"Circumpolar_Domain4_biodiversity_percentiles.png"), width=2400, height=3000, res=300)}
par(mfrow=c(2,1))
plot(masked.ra, main="Circumpolar Biodiversity Hotspots",
     xlim=xlim, ylim=ylim, col=cols[1:4], legend=FALSE, mar=mar)
lines(coast.proj, col=coast.col)
legend("bottom", inset=inset, legend=levels(masked.ra)[[1]][,2], fill=cols, cex=cex, bty="n")
plot(ra, main="Domain 4 Biodiversity Hotspots",
     xlim=xlim, ylim=ylim, col=cols, legend=FALSE, mar=mar)
lines(coast.proj, col=coast.col)
legend("bottom", inset=inset, legend=levels(ra)[[1]][,2], fill=cols, cex=cex, bty="n")
if(pngs){dev.off()}
```

```{r, fig.height=4, fig.width=10}
# Domain to plot
dom <- "dom4"

# Set up plotting area
par(mfrow = c(1, 4), mgp = c(1, 1, 0))
# Loop through variables
for (i in seq_along(var_names)) {
  # Combine all values to set y-axis limits
  all_vals <- c(
    vals.domtop[[dom]][[i]],
    vals.domtopout[[dom]][[i]],
    vals.circtop[[dom]][[i]],
    vals.circtopout[[dom]][[i]]
  )
  # Create empty plot
  plot(1, type = "n", xlim = c(0.5, 2.5), ylim = range(all_vals, na.rm = TRUE),
       xaxt = "n", xlab = "", ylab = "", main = "")
  
  # Add half violins
  vioplot(vals.domtop[[dom]][[i]],    at = 1, side = "left",  col = col.in, add = TRUE)
  vioplot(vals.domtopout[[dom]][[i]], at = 1, side = "right", col = col.out, add = TRUE)
  vioplot(vals.circtop[[dom]][[i]],   at = 2, side = "left",  col = col.in, add = TRUE)
  vioplot(vals.circtopout[[dom]][[i]],at = 2, side = "right", col = col.out, add = TRUE)
  # Add axis and title
  axis(1, at = 1:2, labels = c("Domain", "Circumpolar"))
  title(ylab = y_labels[i], line = 3)
}
```

### Domain 7

```{r, fig.height=12, fig.width=10}
ra <- top.percentiles.domains$`biodiversity domain 7`
masked.ra <- mask(top.percentiles$biodiversity,ra)

xlim=c(1100000,3000000)
ylim=c(-2350000,2350000)
mar <- c(8,1,2,0)
inset=c(0,-0.12)
# l.x.dom=1300000
# l.y.dom=0

if(pngs){png(paste0(pred.dir,"Circumpolar_Domain7_biodiversity_percentiles.png"), width=2400, height=3600, res=300)}
par(mfrow=c(1,2))
plot(masked.ra, col=cols, xlim=xlim, ylim=ylim, main="Circumpolar Biodiversity Hotspots", legend=FALSE, mar=mar)#, axes=FALSE)
legend("bottom", inset=inset, legend=levels(masked.ra)[[1]][,2], fill=cols, cex=cex, xpd=NA)
lines(coast.proj, col=coast.col)
plot(ra,col=cols, xlim=xlim, ylim=ylim, main="Domain 7 Biodiversity Hotspots", legend=FALSE, mar=mar)#, axes=FALSE)
legend("bottom", inset=inset, legend=levels(ra)[[1]][,2], fill=cols, cex=cex, xpd=TRUE)
lines(coast.proj, col=coast.col)
# text(x=l.xt, y=l.yt, "percentiles")
# add.plotsandlines()
if(pngs){dev.off()}
```

```{r, fig.height=4, fig.width=10}
# Domain to plot
dom <- "dom7"

# Set up plotting area
par(mfrow = c(1, 4), mgp = c(1, 1, 0))
# Loop through variables
for (i in seq_along(var_names)) {
  # Combine all values to set y-axis limits
  all_vals <- c(
    vals.domtop[[dom]][[i]],
    vals.domtopout[[dom]][[i]],
    vals.circtop[[dom]][[i]],
    vals.circtopout[[dom]][[i]]
  )
  # Create empty plot
  plot(1, type = "n", xlim = c(0.5, 2.5), ylim = range(all_vals, na.rm = TRUE),
       xaxt = "n", xlab = "", ylab = "", main = "")
  
  # Add half violins
  vioplot(vals.domtop[[dom]][[i]],    at = 1, side = "left",  col = col.in, add = TRUE)
  vioplot(vals.domtopout[[dom]][[i]], at = 1, side = "right", col = col.out, add = TRUE)
  vioplot(vals.circtop[[dom]][[i]],   at = 2, side = "left",  col = col.in, add = TRUE)
  vioplot(vals.circtopout[[dom]][[i]],at = 2, side = "right", col = col.out, add = TRUE)
  # Add axis and title
  axis(1, at = 1:2, labels = c("Domain", "Circumpolar"))
  title(ylab = y_labels[i], line = 3)
}
```

### Domain 8

```{r, fig.height=10, fig.width=6}
ra <- top.percentiles.domains$`biodiversity domain 8`
masked.ra <- mask(top.percentiles$biodiversity,ra)

xlim=c(-800000,1350000)
ylim=c(-2650000,-1150000)
mar <- c(8,1,2,0)
inset=c(0,-0.35)
# l.x.dom=475000
# l.y.dom=-1130000

if(pngs){png(paste0(pred.dir,"Circumpolar_Domain8_biodiversity_percentiles.png"), width=3600, height=1800, res=300)}
par(mfrow=c(1,2))
plot(masked.ra, col=cols, xlim=xlim, ylim=ylim, main="Circumpolar Biodiversity Hotspots", legend=FALSE, mar=mar)#, axes=FALSE)
legend("bottom", inset=inset, legend=levels(masked.ra)[[1]][,2], fill=cols, cex=cex, xpd=NA)
lines(coast.proj, col=coast.col)
plot(ra, col=cols, xlim=xlim, ylim=ylim, main="Domain 8  Biodiversity Hotspots", legend=FALSE, mar=mar)#, axes=FALSE)
legend("bottom", inset=inset, legend=levels(ra)[[1]][,2], fill=cols, cex=cex, xpd=NA)
lines(coast.proj, col=coast.col)
# text(x=l.xt, y=l.yt, "percentiles")
# add.plotsandlines()
if(pngs){dev.off()}
```

```{r, fig.height=4, fig.width=10}
# Domain to plot
dom <- "dom8"

# Set up plotting area
par(mfrow = c(1, 4), mgp = c(1, 1, 0))
# Loop through variables
for (i in seq_along(var_names)) {
  # Combine all values to set y-axis limits
  all_vals <- c(
    vals.domtop[[dom]][[i]],
    vals.domtopout[[dom]][[i]],
    vals.circtop[[dom]][[i]],
    vals.circtopout[[dom]][[i]]
  )
  # Create empty plot
  plot(1, type = "n", xlim = c(0.5, 2.5), ylim = range(all_vals, na.rm = TRUE),
       xaxt = "n", xlab = "", ylab = "", main = "")
  
  # Add half violins
  vioplot(vals.domtop[[dom]][[i]],    at = 1, side = "left",  col = col.in, add = TRUE)
  vioplot(vals.domtopout[[dom]][[i]], at = 1, side = "right", col = col.out, add = TRUE)
  vioplot(vals.circtop[[dom]][[i]],   at = 2, side = "left",  col = col.in, add = TRUE)
  vioplot(vals.circtopout[[dom]][[i]],at = 2, side = "right", col = col.out, add = TRUE)
  # Add axis and title
  axis(1, at = 1:2, labels = c("Domain", "Circumpolar"))
  title(ylab = y_labels[i], line = 3)
}
```

### Domain 9

```{r, fig.height=6, fig.width=10}
ra <- top.percentiles.domains$`biodiversity domain 9`
masked.ra <- mask(top.percentiles$biodiversity,ra)

xlim=c(-2300000,-700000)
ylim=c(-1410000,250000)
mar <- c(8,1,2,0)
inset=c(0,-0.35)
# l.x.dom=-1800000
# l.y.dom=100000
# grat <-  graticule(seq(-150, -90, by=30), proj = crs(ra))#, xlim = xl, ylim = yl)
# plot(grat, axes=FALSE, xlim=xlim, ylim=ylim)
# plot(coast.proj[values(coast.proj)=="land"], col="white", add=TRUE)

if(pngs){png(paste0(pred.dir,"Circumpolar_Domain9_biodiversity_percentiles.png"), width=2400, height=1800, res=300)}
par(mfrow=c(1,2))
plot(masked.ra, main="Circumpolar Biodiversity Hotspots",
     col=cols, xlim=xlim, ylim=ylim, legend=FALSE, mar=mar)#, axes=FALSE)
legend("bottom", inset=inset, legend=levels(masked.ra)[[1]][,2], fill=cols, cex=cex, xpd=NA)
lines(coast.proj, col=coast.col)
plot(ra, main="Domain 9 Biodiversity Hotspots", 
     xlim=xlim, ylim=ylim, col=cols, legend=FALSE, mar=mar)#, axes=FALSE)
legend("bottom", inset=inset, legend=levels(ra)[[1]][,2], fill=cols, cex=cex, xpd=NA)
lines(coast.proj, col=coast.col)
# text(x=l.xt, y=l.yt, "percentiles")
# add.plotsandlines()
if(pngs){dev.off()}
```

```{r, fig.height=4, fig.width=10}
# Domain to plot
dom <- "dom9"

# Set up plotting area
par(mfrow = c(1, 4), mgp = c(1, 1, 0))
# Loop through variables
for (i in seq_along(var_names)) {
  # Combine all values to set y-axis limits
  all_vals <- c(
    vals.domtop[[dom]][[i]],
    vals.domtopout[[dom]][[i]],
    vals.circtop[[dom]][[i]],
    vals.circtopout[[dom]][[i]]
  )
  # Create empty plot
  plot(1, type = "n", xlim = c(0.5, 2.5), ylim = range(all_vals, na.rm = TRUE),
       xaxt = "n", xlab = "", ylab = "", main = "")
  
  # Add half violins
  vioplot(vals.domtop[[dom]][[i]],    at = 1, side = "left",  col = col.in, add = TRUE)
  vioplot(vals.domtopout[[dom]][[i]], at = 1, side = "right", col = col.out, add = TRUE)
  vioplot(vals.circtop[[dom]][[i]],   at = 2, side = "left",  col = col.in, add = TRUE)
  vioplot(vals.circtopout[[dom]][[i]],at = 2, side = "right", col = col.out, add = TRUE)
  # Add axis and title
  axis(1, at = 1:2, labels = c("Domain", "Circumpolar"))
  title(ylab = y_labels[i], line = 3)
}
```
