---
title: "display biodiversity maps"
output: html_document
---

```{r}
## fitting an hmsc using Otsos book, course scripts and : https://besjournals.onlinelibrary.wiley.com/action/downloadSupplement?doi=10.1111%2F2041-210X.13345&file=mee313345-sup-0002-AppendixS2.pdf
library(Hmsc)
library(terra)
library(viridis)
'%!in%' <- function(x,y)!('%in%'(x,y))

##############################################################################################################
#res <- "500m"
res <- "2km"

##############################################################################################################
user = "Jan"
#user = "charley"
#user="nicole"
if (user == "Jan") {
  sci.dir <-      "C:/Users/jjansen/Desktop/science/"
  env.derived <-  paste0(sci.dir,"data_environmental/derived/")
  #bio.dir <-      paste0(sci.dir,"data_biological/")
  ## remote repository (DOESN'T WORK YET):
  # env.dir <- "https://data.imas.utas.edu.au/data_transfer/admin/files/EnvironmentalData/"
  ## common paths (after "sci.dir")
  tools.dir <-    paste0(sci.dir,"SouthernOceanBiodiversityMapping/Useful_Functions_Tools/")
  ARC_Data.dir <- paste0(sci.dir,"SouthernOceanBiodiversityMapping/ARC_Data/")
} 
if (user == "charley") {
  sci.dir <- "C:/Users/cgros/code/IMAS/"
  ARC_Data.dir <- paste0(sci.dir,"ARC_Data/")
  env.derived <-  "C:/Users/cgros/data/SO_env_layers/derived/"
  tools.dir <-    paste0(sci.dir,"Useful_Functions_Tools/")
}
if (user == "nicole") {
  sci.dir <-    "C:/Users/hillna/OneDrive - University of Tasmania/UTAS_work/Projects/Benthic Diversity ARC/"
  ARC_Data.dir <- paste0(sci.dir,"Analysis/ARC_Data/")
  env.derived <-  paste0(sci.dir,"data_environmental/derived/")
  tools.dir <-    paste0(sci.dir,"Analysis/Useful_Functions_Tools/")
}

r.stack <- rast(paste0(env.derived,"Circumpolar_EnvData_",res,"_shelf_mask_unscaled_variables.tif"))
r2 <- r.stack$depth
```

```{r}
modeltype = 1
model = 1
thin = 10  ## a value of 10 means every 10th iteration is kept (the higher the less correlated the samples are but the longer it takes)
samples = 800 ## how many total samples we want
transient = ceiling(0.5*samples*thin)
adaptNf = rep(ceiling(0.4*samples*thin),1)
nChains = 4

#########################################
## first define the character string to save the files:
base.str <- paste0("/pvol/biodiversity_prediction/",res,
                    "_model_", as.character(model), "_",
                    c("pa","abundance")[modeltype], 
                    "_chains_",as.character(nChains),
                    "_thin_", ... = as.character(thin),
                    "_samples_", as.character(samples),"_")
```

```{r}
################################
## read in and compare predictions:
ra.dir <- paste0(sci.dir,"data_biological/circumpolar_prediction_outputs/")
ra.list <- list.files(ra.dir)
ra.list <- ra.list[grep("800_",ra.list)]

hmsc.maps <- rast(paste0(ra.dir,ra.list))

hmsc.maps.mean <- subset(hmsc.maps,seq(1,nlyr(hmsc.maps),by=2))
hmsc.maps.sd <- subset(hmsc.maps,seq(2,nlyr(hmsc.maps),by=2))

hmsc.maps.richness <- sum(hmsc.maps.mean)
```

## p/a maps {.tabset}

### p/a - overview

```{r, fig.width=10, fig.height=10}
plot(hmsc.maps.richness, col=viridis(99))
```

### p/a - individual species

```{r, fig.width=13, fig.height=7}
par(mfrow=c(1,2), oma=c(2,0,2,3))
mar <- c(1,1,1,0)
for(i in seq(1,166, by=2)){
  range <- range(minmax(c(hmsc.maps[[i]],hmsc.maps[[i+1]])))
  plot(hmsc.maps[[i]], range=range, main=names(hmsc.maps)[i], mar=mar, legend=FALSE, col=viridis(99))
  plot(hmsc.maps[[i+1]], range=range, main=names(hmsc.maps)[i+1], mar=mar, col=viridis(99))
}
```


### p/a uncertainty hotspots

requires more thinking about how to detect the hotspots  
essentially need to look at where a standardised prediction error is the smallest, because larger predictions will have larger errors... And for p/a data small and large probabilities will have small errors, so can't simply divide occurrence by the error...  

```{r, eval=FALSE, echo=FALSE, fig.width=13, fig.height=7}
par(mfrow=c(1,2), oma=c(2,0,2,3))
mar <- c(1,1,1,0)
uncertainty_map <- hmsc.maps[[19]]/hmsc.maps[[20]]
plot(hmsc.maps[[19]], main=names(hmsc.maps)[19], mar=mar, legend=FALSE, col=viridis(99))
plot(uncertainty_map, main="proportional uncertainty", mar=mar, col=viridis(99))

par(mfrow=c(1,1), oma=c(2,0,2,3))
mar <- c(1,1,1,0)
uncertainty_map <- hmsc.maps[[19]]/(hmsc.maps[[20]]^2)
plot(uncertainty_map, main="proportional uncertainty", mar=mar, col=viridis(99))


#top 10% occurrence, top 1% certainty
#top10% occurrence, bottom 1% certainty

unc.map2 <- uncertainty_map
unc.map2[which(!is.na(uncertainty_map[]))] <- 0
sel.uncert.top1 <- selectHighest(uncertainty_map, 50000)
unc.map2[sel.uncert.top1] <- 1

prob.map2 <- hmsc.maps[[19]]
prob.map2[which(!is.na(hmsc.maps[[19]][]))] <- 0
sel.prob.top1 <- selectHighest(hmsc.maps[[19]], 50000)
prob.map2[sel.prob.top1] <- 1

par(mfrow=c(1,1))
plot(unc.map2, main="top 5% certainty")
plot(prob.map2, main="top 5% probability of presence")
```

```{r, eval=FALSE, echo=FALSE, fig.width=13, fig.height=7}
m <- hmsc.maps[[47]]
s <- hmsc.maps[[48]]
par(mfrow=c(1,2), oma=c(2,0,2,3))
mar <- c(1,1,1,0)
uncertainty_map <- m/s
plot(m, main=names(m), mar=mar, legend=FALSE, col=viridis(99))
plot(uncertainty_map, main="proportional uncertainty", mar=mar, col=viridis(99))

par(mfrow=c(1,1), oma=c(2,0,2,3))
mar <- c(1,1,1,0)
uncertainty_map <- m/(s^2)
plot(uncertainty_map, main="proportional uncertainty", mar=mar, col=viridis(99))


#top 10% occurrence, top 1% certainty
#top10% occurrence, bottom 1% certainty

# unc.map2 <- uncertainty_map
# unc.map2[which(!is.na(uncertainty_map[]))] <- 0
# sel.uncert.top1 <- selectHighest(uncertainty_map, 50000)
# unc.map2[sel.uncert.top1] <- 1
unc.map2 <- s
unc.map2[which(!is.na(s[]))] <- 0
sel.uncert.top1 <- selectHighest(s, 200000, low=TRUE)
unc.map2[sel.uncert.top1] <- 1

prob.map2 <- m
prob.map2[which(!is.na(m[]))] <- 0
sel.prob.top1 <- selectHighest(m, 200000)
prob.map2[sel.prob.top1] <- 1

par(mfrow=c(1,1))
plot(unc.map2, main="top 10% certainty")
plot(prob.map2, main="top 10% probability of presence")
plot(unc.map2+prob.map2)
```

### p/a - regional snapshots

#### Antarctic Peninsula

```{r, fig.height=8, fig.width=10}
maxcell=10000000
xlim=c(-2800000,-1800000)
ylim=c(1350000,2350000)
plot(hmsc.maps.richness, xlim=xlim, ylim=ylim, maxcell=maxcell, col=viridis(99))
```

#### Ross Sea

```{r, fig.height=8, fig.width=10}
maxcell=10000000
xlim=c(-800000,500000)
ylim=c(-2350000,-1250000)
plot(hmsc.maps.richness, xlim=xlim, ylim=ylim, maxcell=maxcell, col=viridis(99))
```

### p/a - East Antarctica - individual species

```{r}
source(paste0(tools.dir,"SOmap_functions_JJ.R"))
```

```{r}
library(raster)
JJ_SOmap(env.var.raster=raster(hmsc.maps[[19]]), ramp.col = viridis(99), border=FALSE, graticules = TRUE, label = "probability of presence") #, leg.steps=leg.steps, lcex=lcex
```




```{r, fig.width=7, fig.height=10}
maxcell=10000000
xlim=c(0,2500000)
ylim=c(-2400000,-1000000)
par(mfrow=c(2,1), bg=NA)
for(i in 19){
  range <- range(minmax(c(hmsc.maps[[i]],hmsc.maps[[i+1]])))
  plot(hmsc.maps[[i]], xlim=xlim, ylim=ylim, maxcell=maxcell, range=range, main=names(hmsc.maps)[i], mar=mar, legend=FALSE, col=viridis(99))
  plot(hmsc.maps[[i+1]], xlim=xlim, ylim=ylim, maxcell=maxcell, range=range, main=names(hmsc.maps)[i+1], mar=mar, col=viridis(99))
}
```

```{r, fig.width=7, fig.height=10}
maxcell=10000000
xlim=c(1000000,3000000)
ylim=c(-2400000,1600000)
png(paste0(sci.dir,"Circumpolar_BHAnt_zoom.png"), width=2400, height=2400, res=300)
par(mfrow=c(1,2),bg=NA)
for(i in 19){
  range <- range(minmax(c(hmsc.maps[[i]],hmsc.maps[[i+1]])))
  plot(hmsc.maps[[i]], xlim=xlim, ylim=ylim, maxcell=maxcell, range=range, main=names(hmsc.maps)[i], mar=mar, legend=FALSE, col=viridis(99), axes=FALSE)
  plot(hmsc.maps[[i+1]], xlim=xlim, ylim=ylim, maxcell=maxcell, range=range, main=names(hmsc.maps)[i+1], mar=mar, col=viridis(99), axes=FALSE)
}
dev.off()
```
```{r, fig.width=7, fig.height=10}
maxcell=10000000
xlim=c(1000000,3000000)
ylim=c(-2400000,1600000)
png(paste0(sci.dir,"Circumpolar_richness_zoom.png"), width=2400, height=2400, res=300)
par(mfrow=c(1,2),bg=NA)
for(i in 19){
  range <- range(minmax(c(hmsc.maps[[i]],hmsc.maps[[i+1]])))
  plot(hmsc.maps.richness, xlim=xlim, ylim=ylim, maxcell=maxcell, mar=mar, col=viridis(99), axes=FALSE)
}
dev.off()
```
